### 2018-02-04 paper.tex
Updating OPTN environment subsection

\documentclass[12pt]{article}

\usepackage{amssymb,amsmath,amsfonts,eurosym,geometry,ulem,graphicx,caption,color,setspace,sectsty,comment,footmisc,caption,pdflscape,subfigure,array,hyperref}
\usepackage{cite}
\usepackage{natbib}
\usepackage{etoolbox}
\usepackage{float}
\usepackage{algpseudocode}
\usepackage{booktabs}

\AtBeginEnvironment{algorithm}{\setstretch{1}}

\normalem

%\geometry{left=1.0in,right=1.0in,top=1.0in,bottom=1.0in}

\begin{document}

\begin{titlepage}
\title{Improving Dynamic Kidney Exchange}
\author{Vitor Hadad\thanks{Boston College}}
\date{December 2017}
\maketitle
\begin{abstract}

\bigskip
\end{abstract}
\setcounter{page}{0}
\thispagestyle{empty}
\end{titlepage}
\pagebreak

%\doublespacing

\section{Introduction} \label{sec:introduction}

\paragraph{Main idea} We propose a new dynamic kidney matching algorithm that takes as input a pool of patient-donor pairs and computes, for each cycle that is currently available for matching, a probability that it should be chosen to be cleared today, as well as another probability indicating whether we should not match any more cycles. In order to get these probabilities, our algorithm first produces a prior distribution estimated by observing past choices made by an infeasible optimal algorithm, and then updates these priors with simulations.

Our algorithm faces two main challenges. The first challenge is the input data representation, because the relevant information consists not only of the characteristics of each donor-patient pair, but also the underlying structure of their compatibility graph. For example, we need it to be the case that the same cycle will have different probabilities of being matched depending on what who are all the other patient-pairs in the graph, what characteristics they have, and how compatible their are to each other. Our solution to this obstacle involves using mapping the raw input data and the compatibility graph to points on the euclidean space, and then using these as regressors.

The second challenge is output data representation. As we said above, one of the components of the output is a joint probability distribution over each available cycle, indicating the likelihood that the cycle will be matched today. However, the number of available cycles will be different at each period, so we need an algorithm that is able to output probability distribution over a vector of variable length. In this paper, we experiment with a variety of neural network architectures for this task, and we note that recurrent neural networks provide the best performance.

\paragraph{Main results} We test our algorithm in a variety of data-generating processes that we call \emph{environments}. Each environment was inspired by previous work on dynamic kidney matching, and has different rules governing which patient-donor pairs are compatible or what are their relevant observable characteristics. In each environment, we compare the performance of our algorithm, as measured by the number of matched pairs within a long time period, to two benchmarks. The first one is an infeasible optimal algorithm (\texttt{OPT}) that knows exactly which patients will be come to the pool and how long their sojourn will be, and is therefore able to compute the maximal matching cardinality. Our second benchmark is \texttt{Myopic}, an algorithm that greedily matches and clears the maximal matching at each period, and is used a proxy to the algorithms that are being used in many real life kidney exchanges. Our results show that we were indeed able to improve upon \texttt{Myopic}, proving that indeed there might be potential gains to be made from taking dynamic issues into serious consideration.

Our simulations show that there is a sizable performance gap between \texttt{Myopic} and \texttt{OPT}: for the parameter combinations we have tested, \texttt{Myopic} may match only about 75\% of the pairs that \texttt{Myopic} matches. In addition, this performance gap widens as we increase the number of maximum cycle length from 2 to 3.

At the same time, we remark that even after our improvements, the gap between \texttt{Myopic} and \texttt{OPT} resists. Therefore, in this paper we also invite other researchers to consider alternative algorithms for dynamic kidney exchange, or produce theoretical upper bounds for them. We hope that the new methods proposed in this paper serve as further inspiration for the field.

\section{Background and literature review} \label{sec:literature}

Since its proposal in the medical literature over three decades ago by \cite{rapaport1986case}, kidney exchange has been the subject of a great deal of attention in the academic literature. Throughout the past twenty years, developments in microeconomic theory and operations research and have informed, reshaped and improved how living donor transplants are conducted in many countries. More recently, computer scientists have also started to focus on the production of better and faster algorithms for kidney exchange.

Since this paper sits in the intersection between matching theory and computational economics, and uses methods drawn from computer science and machine learning. Let us take a brief look at these fields in turn.

\subsection{Kidney Exchange} \label{subsec:kidney}

\paragraph{Definitions}

>HLA compatibility can also be defined in terms of
mismatch acceptability—mismatches in allosensitized
transplant candidates that result in a negative crossmatch
[74]. In this context, unacceptable mismatches are antigens
reacting with antibody detectable in the patient
sera, whereas acceptable mismatches are those with no
detectable antibody

\paragraph{Early years} In economics, the literature on kidney exchange is started by \cite{roth2004kidney}, who provided a kidney exchange mechanism inspired by the \emph{housing} problem studied in the literature of mechanism design as in \cite{shapley1974cores} and later \cite{abdulkadirouglu1999house}. While their mechanism had significant virtues as being strategy-proof and Pareto-efficient, it also relied on large number of exchanges being conducted simultaneously, and drew criticism for making assumptions about heterogeneous preferences over kidneys. Responding to these assessments, a subsequent work by \cite{roth2005pairwise} focused on logistically practical mechanisms that used only 2-way exchanges and also had patients indifferent over all compatible kidneys. In a later paper, \cite{roth2007efficient} demonstrated via simulations that allowing for 3-way exchanges in addition to 2-way exchanges could increase the cardinality of matched pairs by a great deal, but also showed that larger cycles would only bring about modest improvements.

\paragraph{Recent advances in static matching problems} In the second half of the oughties, both in academic literature and medical practice the focus began to shift from exchanges via cycles to exchanges emanating from \emph{non-directed donors} (NDD) -- altruistic donors who have no patient attached to them and are willing to donate their organs to anyone who would need it. Such exchanges yield a \emph{chain} of transplants that begins with the NDD and may either terminate with the last pair donating to a waitlist recipient in a kidney registry, or not terminate at all and have the last pair's donor await an opportunity to become a future living donor. The former kind was dubbed \emph{domino paired donation} by \cite{montgomery2006domino}, and all its transplants occur simultaneously. In the latter kind, transplants occur may staggeredly over several months, hence it was named \emph{non-simultaneous, extended, altruistic donor chain} by \cite{rees2009nonsimultaneous}.\footnote{On occasion, NEAD chains have also been called \emph{Never-ending altruistic donor chains}, as relayed by \cite[p.~235-6]{roth2015gets}.}

A significant amount of attention was devoted in the next years to these NEAD chains, including notably \cite{ashlagi2011nonsimultaneous} and \cite{ashlagi2012need}, whose results show that NEAD chains benefit highly sensitized patients in sparse pools of moderate size, because they decrease the need for simultaneous double-coincidence of wants in the exchange market. In fact, in recent years \cite{anderson2015finding} reported that ``NEAD chains are responsible for the majority of successful kidney transplants conducted via kidney exchange at both the [Allied for Paired Donation] and within other major exchange programs".

\paragraph{Dynamic kidney exchange} Meanwhile, a small subliterature started focusing on the problem of \emph{dynamic kidney exchange} that concerns this paper. In a seminal paper, \cite{unver2010dynamic} posited a continuous-time model where pairs arrive according to a Poisson arrival rate, enter the pool whenever their donor is incompatible with them, and from thereon suffer a constant cost of waiting Ünver was able to derive dynamic mechanism that depends produces optimal n-way exchanges by leveraging additional simplifying assumptions, including that pairs do not leave the pool unless they are matched, and that pairs within the pool are only blood-type incompatible. These assumptions render all pairs homogeneous, which dramatically decreases the dimension of the state space and allows for an easily interpretable solution.

\cite{ashlagi2013kidney} note that the graphs of real-life pools tend to be sparse and filled with pairs that are either easy or extremely hard to match. In order to model this particular feature, they postulate a sparse heterogeneous random graph model containing only two such types, and propose a greedy algorithm that waiting until there are a certain number of each type, and then matches as many as pairs as possible. In a different vein, \cite{akbarpour2017thickness} introduces a simple model of stochastically arrival and departure. In their setting, they show that greedy algorithms that cleverly explout the time to match can can perform close to infeasible optimal benchmarks, even if these algorithms are ignorant about the global structure of the graph.

In the computer science and operations research literature, \cite{abraham2007clearing} and more recently \cite{anderson2015finding}, \cite{dickerson2016position} and \cite{dickerson2017small} have focused on producing scalable combinatorial programming algorithms that can take on tatic matching problems with large graphs and allowing for both chains and cycles of moderately large length. However, a series of papers starting with \cite{awasthi2009online} and followed by \cite{dickerson2012dynamic} and \cite{dickerson2015futurematch} have dealt with the dynamic kidney problem by \emph{weighted myopia}. Their idea is to prevent wasteful matchings (e.g., an O-donor to an AB-patient) by artificially introducing negative weights to graph components containing them. In \cite{dickerson2015futurematch}, these optimal weights are computed from simulations involving historical data.

\paragraph{Miscellaneous} Kidney exchange is a rich problem that can be studied from several different angles. For example, it is a special case of barter exchange, [[[Barter]]].  [[[Hospital incentives]]] [[[Medical literature]]]

\subsection{Machine learning and Artificial Intelligence}

\subsubsection{Neural Networks}

Neural networks are created by composing simpler parametric building blocks such as linear functions and simple nonlinear ones\cite{goodfellow2016deep}. Although they have existed for more than a half-century, in recent years they have received an enormous amount of attention in the statistical, computer science and optimization literatures. In this work, we will use \emph{multilayer feedforward networks}, \emph{recurrent neural networks} (with and without an extra property named \emph{attention}) and \emph{graph convolutional networks}. While a comprehensive survey on the motivation and use-cases of all these types of neural networks is outside the scope of this work, in the next paragraphs we will give an abridged overview of their definition and their immediate application to our problem.

\paragraph{Feedforward Neural Networks} Also called \emph{multilayer perceptrons}, these functions process each input

\paragraph{Recurrent Neural Networks and Sequence-to-sequence models}
\paragraph{Embedding non-euclidean domains} \label{par:embedding}
\subsubsection{Search algorithms}

\section{Models, Environments and Data} \label{sec:data}

A \textbf{dynamic kidney exchange problem} is a discrete-time Markov Decision Process (MDP) described by the tuple $(S, A, P, R)$, where

\begin{itemize}
  \item $S = (V, A, X)$ is a directed graph representing the current \emph{state}: the set $V$ is the finite set of patient-donor pairs and their observable characteristics; and an adjacency matrix $A$ whose entry $(i,j)$ is positive whenever the donor of vertex $i$ can donate to the patient of vertex $j$. The set $X$ is a set of 
  \item $Act$ is the finite set of available \emph{actions} when the current state is $S$. In our applications, this is the set of 2- or 3-cycles available to be cleared.
  \item $P$ is the \emph{transition probability} over the next states given current state and taken action.
  \item $R$ is the \emph{reward function} indicating how desirable it is to take an action $a \in Act$ when the state is $S$.
\end{itemize}

The space where a dynamic kidney exchange problem tuple lives is called a \textbf{kidney exchange environment}. Informally, an environment is the set of rules or configurations that govern, for example, whether we can match 2-cycles or 3-cycles per period, or which characteristic are relevant and available for the decision agent to observe. Before we describe each environment in detail, here are some common assumptions.

\paragraph{Poisson entry, Geometric death} At time point, the number of new incoming pairs is drawn from the $Poisson(r)$ distribution, where $r \in \mathbb{N}$ denotes the \emph{entry rate}, and equals the expected number of entrants per period. Upon entrance, each pair independently draws the length of their sojourn from the $Geometric(d)$ distribution. The parameter $d \in \mathbb{R}$ is the \emph{death rate}, and its reciprocal $\frac{1}{d}$ is the expected sojourn length. We note that, due to the memoryless property\footnote{If $X \sim Geometric(p)$, then $P(X > t+s | X > s) = P(X > t)$} of the Geometric distribution, the amount of time a pair has waited in the pool gives us no information about how much time they have until their death.

\paragraph{Observables} In every environment, at least two characteristics are relevant and observable: the major blood type group (A, B, O or AB); and how long the pair has been in the pool.

\paragraph{0-1 Preferences} In reality, some exchanges are more or less desirable than others, either for ethical concerns or because of predicted health benefit. For example, it is common for pediatric patients and previous organ donors receive higher priority, as do exchanges involving patients with no HLA mismatch. However, in this work we abstract from these concerns and consider every exchange to be equally desirable, so long as it is available. This is the same assumption as used in \cite{roth2005pairwise}.

\noindent Now, let's see each of the different environments in turn.

\subsection{ABO Environment}

The \emph{ABO environment} is the barest: each pair in the pool is solely characterized by their entry time, the length of their sojourn, and the ABO blood types of its patient and donor. Compatibility between two pairs is also decided only on blood-type compatibility. However, we make an assumption previously used in \cite{unver2010dynamic} and allow for incompatibility between a donor and their own patient. The reason for this additional assumption is that, if there were truly no tissue type compatibilities, we would never see pairs of type (AB,$\cdot$), ($\cdot$, O), or (A,A), (O,O), (B,B), and (AB,AB), since their donors would be automatically compatible with their patients and they would never participate in an exchange. Let's explain the effect of this assumption in more detail.

The distribution of blood types in the US population is roughly O:48\%, A:36\%, B:11\%, AB:4\%. If we initially independently draw two people from this distribution and they form, say, an (A,B) pair (which happens with probability $0.36 \cdot 0.12 \approx 0.0432$), then we allow them to the pool immediately, since their B-donor is blood-type incompatible with their A-patient. However, if they happen to form an (A,O) pair (with probability $0.36 \cdot 0.48 \approx 0.172$), then they should only enter the pool if the donor and patient are tissue-type incompatible, otherwise the O donor would immediately donate their kidney to the patient and they would not need to enter the exchange pool at all. So we assume the probability of a positive crossmatch is $p_c = 0.11$ as in \cite{zenios2001primum}, then the probability that two people form an (A,O) and enter the pool becomes in fact $0.36 \cdot 0.48  \cdot 0.11 \approx 0.019$. By computing the probabilities of each pair in this manner and then normalizing so that they add up to one, we arrive at the numbers displayed on Table \ref{tab:abo_blood_type}.

\begin{table}[H]
\centering
\begin{tabular}{lr}
\toprule
Pair blood type &         Probability \\
\midrule
(O, O) &  0.058689 \\
(O, A) &  0.373803 \\
(O, B) &  0.158257 \\
(O, AB) &  0.042669 \\
(A, O) &  0.041119 \\
(A, A) &  0.028809 \\
(A, B) &  0.110888 \\
(A, AB) &  0.029899 \\
(B, O) &  0.017410 \\
(B, A) &  0.110888 \\
(B, B) &  0.005160 \\
(B, AB) &  0.012660 \\
(AB, O) &  0.004690 \\
(AB, A) &  0.003290 \\
(AB, B) &  0.001390 \\
(AB, AB) &  0.000380 \\
\bottomrule
\end{tabular}
\caption{Blood type probabilities in the ABO environment}
\label{tab:abo_blood_type}
\end{table}

\subsection{RSU Environment}

The \emph{RSU} environment was inspired largely by the simulation setup in \cite{roth2007efficient}. In these works, in addition to the blood type, a pair is also endowed with a panel reactive antibody (PRA) level that governs the probability of a crossmatch with a random donor. The lower the PRA, the higher the number of potentially compatible pairs.

The simulation process is as follows. Initially, we draw a pair in the same manner as in the ABO environment. Next, we draw if the patient is a female (with probability around 41\%), and if so we also draw whether her donor is her husband (spouses comprise about 49\% of donors). Finally, we draw a PRA level for the patient (Low: 70.1\%, Medium: 20\%, High: 9.9\%). This PRA level determines the probability that they can receive a kidney from any donor, including their own: patients with low PRA have a 5\% probability of positive crossmatch with a random donor; patients with medium PRA have a 45\% chance, and patients with high PRA have a 90\% chance of a crossmatch. If a patient is bloody or tissue-type incompatible with their own donor, they enter the pool. In addition, if the patient is female and her husband is the donor, the probability of positive crossmatch for low, medium and high PRA patients goes up to 28.75\%, 58.75\% and 92.25\%. This last adjustment reflects the fact that women tend to produce antibodies against their husbands' antigens during pregnancy.

Once in the pool, the pair immediately forms directed edges with the existing pairs, again following the patient PRA distribution. The resulting random graph is akin to a Erd\"{o}s-R\'{e}nyi $G(n,p)$ random graph, except that the edge-forming probability is heterogeneous across different pair types.

\subsection{OPTN Environment}

Here we use historical data collected by the United Network for Organ Sharing (UNOS) data provided in the Standard Research and Analysis (STAR) dataset. The STAR dataset contains information from all patients that were ever registered to the kidney waiting list in the United States for the past three decades, as well as from all living donors that actually went to transplant.

\paragraph{Compatibility} In this environment, a donor and a patient are deemed \emph{compatible} if they are blood-type compatible and the donor's HLA profile exhibits no antigens that are unacceptable for the patient in any of the HLA-A, HLA-BW, HLA-C, HLA-DR, HLA-DPB, HLA-DQ, and HLA-DQA loci.

\paragraph{Simulation} We first created an artificial data set by randomly drawing patients from the historical waiting list, drawing donors from the historical living donor list, and checking for compatibility as explained in the previous paragraph. Compatible pairs were discarded. We iterated in this manner to construct a dataset of about one million incompatible pairs.

When simulation this environment, at each period pairs are drawn from this artificial data set. Once a pair enters the pool, it immediately forms directed edges with existing compatible pairs.

\paragraph{Observable characteristics} The agent observes

% Here we must emphasize that these are \emph{not} the patients that belong to the Kidney Paired Donation Pilot Program, and pairs that belong to kidney exchanges may have different characteristics from patients who were only waiting for a living donor (for example, \cite{ashlagi2013kidney} states that patients are much more likely to be highly sensitized). However, we believe that, as exchanges get larger, the distribution of pair characteristics should grow similar to their historical counterpart in the STAR dataset.

[[[Historical donor and patient cPRA]]].

\subsection{Benchmarks}

In this work, we focus on maximizing the cardinality of matchings over $T$ periods.

\subsubsection{Myopic}

The \emph{myopic}\footnote{In the computer science literature, algorithms that optimize locally at every step are usually called \emph{greedy}. However, in the kidney exchange literature this name is already reserved to an algorithm that matches incoming pairs at random, without necessarily optimizing the cardinality of matchings at every period.} algorithm is one that finds the maximal number of matchings at every period, ignoring that it might be useful to keep certain pairs in the pool for later use. As a result, the graph at the end of every period has no cycles shorter than the maximal allowed cycle length, for they would have been already cleared.

This algorithm also does not take into consideration any of the observables in the data. In particular, it makes no use of the fact that some pairs may be harder to match in the future, and therefore it might be advantageous to match them today if the opportunity arises. Conversely, it will not be less likely to carry out matchings involving easy-to-match pairs that could be postponed to the future at low expected cost.

\begin{align}
   &\max_{x} \sum_{c \in C} w_{c} x_{c} \\
      s.t. &\sum_{v \in c'} x_{c'} = 1 \qquad \forall v \in V\\
        &x_c \in \{ 0, 1\}  \qquad \forall c \in C
\end{align}

\subsubsection{Infeasible Optimal (OPT)}

The infeasible optimal algorithm (henceforth OPT) does away with the uncertainty arising from the temporal structure of the problem: it knows exactly which pairs will come into the pool, as well as their arrival, and the duration of their sojourns. In effect, this algorithm simply solves the static integer programming problem described at the beginning of section \ref{sec:literature:subsec:kidney}, except that the set of cycles has the additional constraint that a cycle cannot exist between two vertices if their sojourns fail to overlap.

% \subsubsection{Empirical comparison between myopic and OPT}
% 
% We simulated each environment for 3000 periods for each given environment and for each combination of entry rate, death rate and maximum available cycle length. Next, we analyzed the performance of OPT and myopic in each setup, as measured by the average number of pairs they were able to match per period. In order to avoid mixing problems, we discarded the first 1000 periods as burn-in.\footnote{While we refrain from a formal analysis of mixing time here, our thousand-period burn-in may be fairly conservative, as we empirically observe that for all our parameter combinations the size of the graph quick grows from zero to a stable level in (sometimes much) fewer than one hundred periods.}
% 
% The results are seen on Figure \ref{???}, where we show the performance ratio between the two algorithms. It is perhaps surprising that myopic's performance is often very close to OPT. In particular, when entry rates are relatively large or death rates are relatively low, the pool becomes thick enough that  opportunities for a reasonably good matching abound, so even a simple heuristic like myopic is able to do well. A bit less remarkably, high death rates also allow myopic to thrive, since dynamic considerations play a much smaller role when pairs leave the pool more often.
% 
% A more interesting remark is that in all of our three environments, when 3-cycles are allowed myopic does relative worse to OPT than when only 2-cycles are allowed. 
% 
% \subsection{Methods}
% 
% In this work, we propose a new kidney exchange algorithm that combines simulation-based search with a policy parametrized by neural networks. In the next paragraphs, we will explain each of these components in detail. However, first let us provide some motivation.
% 
% We begin by noting that we have a promising setup, as we are in possession of: good estimates of the data-generating process for each environment; an algorithm that is able to compute optimal actions, even if only in hindsight (\texttt{OPT}); one clear benchmark that we must beat (\texttt{Myopic}). At first glance, then, one might hypothesize that it would be possible to find reasonably good solutions using only simulations: simply simulate future states to a long horizon ahead, and choose today the action that on average leads to higher future rewards -- or the one that is most often chosen by \texttt{OPT}, or the one that most frequently leads to beating \texttt{Myopic}. Unfortunately, this turns out to be overly optimistic: approaches solely based on simulations do not perform very well, as shown in Table \ref{simulations_only}. We believe that the reason for this poor performance is the high branching factor of the problem -- in other words, there is high uncertainty about what the next period will look like.
% 
% 
% Reducing the problem:
% 
% 1. Our "actions" are simply a number of binary flags indicating who should be matched today and who should not
% 2. We can leverage the optimal algorithm to reduce the MDP to a simpler period-by-period MAB problem
% 3. Use 
% 
% % In lieu of just simulations, we propose to see each period's problem as a Bayesian multi-armed bandit problem (MAB). ... estimate... then use estimates as pseudo-counts... then sample from posterior.
% 
%

\subsubsection{Estimation} \label{subsubsec:estimation}

As explained in section \ref{subsubsec:embedding}, one of our main challenges is that we must deal not only with information about each patient-donor pair, but also

%\insert{algo1.tex}

% Using Thompson sampling at each episode might sound like a greedy approach to a problem where current actions might have effects on future rewards, but here we *are* using future rewards since we can simulate the future \emph{and} also assume that we will be act optimally within a time horizon.

\section{Results} \label{sec:results}

\subsection{Short-run gains}

\section{Discussions and Extensions} \label{sec:discussion}

HLA mistmatch -- doctors are willing

\section{Conclusion} \label{sec:conclusion}

\section{Acknowledgements}

This work was supported in part by Health Resources and Services Administration contract 234-2005-37011C. The content is the responsibility of the authors alone and does not necessarily reflect the views or policies of the Department of Health and Human Services, nor does mention of trade names, commercial products, or organizations imply endorsement by the U.S. Government.

\clearpage

\bibliographystyle{/Users/vitorhadad/Documents/kidney/matching/paper/te.bst}
\bibliography{references}

\clearpage

%\onehalfspacing

\section*{Tables} \label{sec:tab}
\addcontentsline{toc}{section}{Tables}

\clearpage

\section*{Figures} \label{sec:fig}
\addcontentsline{toc}{section}{Figures}

\clearpage

\section*{Appendix A. Placeholder} \label{sec:appendixa}
\addcontentsline{toc}{section}{Appendix A}

\end{document}

### 2018-02-04 paper.tex
Began algorithms section, drew three figures.

@@
 
-%\geometry{left=1.0in,right=1.0in,top=1.0in,bottom=1.0in}
+\geometry{left=1.5in,right=1.5in,top=1.0in,bottom=1.0in}
 
@@
 The STAR dataset contains information from all patients that were ever registered to the kidney waiting list in the United States for the past three decades, as well as from all living donors that actually went to transplant.
+We excluded individuals with missing HLA profile information,
+we had access to a detailed historical dataset containing 117813 patients and 9337 donors.
 
-\paragraph{Compatibility} In this environment, a donor and a patient are deemed \emph{compatible} if they are blood-type compatible and the donor's HLA profile exhibits no antigens that are unacceptable for the patient in any of the HLA-A, HLA-BW, HLA-C, HLA-DR, HLA-DPB, HLA-DQ, and HLA-DQA loci.
-
-\paragraph{Simulation} We first created an artificial data set by randomly drawing patients from the historical waiting list, drawing donors from the historical living donor list, and checking for compatibility as explained in the previous paragraph.
+The data set about was used in two ways.
+First, we created an artificial data set by randomly drawing patients from the historical waiting list, drawing donors from the historical living donor list, and checking for blood- and tissue-type compatibility as explained below.
 Compatible pairs were discarded.
@@
 
-When simulation this environment, at each period pairs are drawn from this artificial data set. Once a pair enters the pool, it immediately forms directed edges with existing compatible pairs.
+Second, we  used the historical data set to calculate two additional variables, a \emph{patient cPRA} and  a \emph{donor cPRA}.\footnote{We thank Itai Ashlagi for the suggestion of a donor cPRA.} As explained before, the patient cumulative panel reactive antibody, or cPRA, represents a measure of patient tissue-type incompatibility with a random donor, the donor cPRA is a novel measure of the opposite direction -- how frequently a donor is tissue-type incompatible with a random patient.
+We computed both measures empirically in our data set, using the definiton of compatibility below.
 
-\paragraph{Observable characteristics} The agent observes
+\begin{figure}
+\centering
+\includegraphics[width=1textwidth]{/Users/vitorhadad/Documents/kidney/matching/figures/computed_cpra.pdf}.
+\caption{Computed patient and donor cPRA.}
+\label{fig:cpra}
+\end{figure}
 
-% Here we must emphasize that these are \emph{not} the patients that belong to the Kidney Paired Donation Pilot Program, and pairs that belong to kidney exchanges may have different characteristics from patients who were only waiting for a living donor (for example, \cite{ashlagi2013kidney} states that patients are much more likely to be highly sensitized).
-However, we believe that, as exchanges get larger, the distribution of pair characteristics should grow similar to their historical counterpart in the STAR dataset.
+\paragraph{Compatibility} A donor and a patient are deemed \emph{compatible} if they are blood-type compatible and tissue-type compatible.
+The latter is true when the donor's HLA profile exhibits no antigens that are unacceptable for the patient in any of the A, B, Bw, C, DR, DPB, DQ, and DQA loci.
 
-[[[Historical donor and patient cPRA]]].
+\paragraph{Simulation} During a simulation, pairs are drawn from this artificial data set. Once a pair enters the pool, it immediately forms directed edges with existing compatible pairs.
+
+\paragraph{Observable characteristics} For each pair, the agent observes the entry time, the waiting time, and 375 dummies corresponding to blood type and HLA profile of patient and donor.
 
@@
 
-In this work, we focus on maximizing the cardinality of matchings over $T$ periods.
+In this work, we focus on maximizing the average number of matched pairs over $T$ periods, where $T$ is a large number relative to the rates of patient entry and death.
 
@@
 
-The \emph{myopic}\footnote{In the computer science literature, algorithms that optimize locally at every step are usually called \emph{greedy}.
-However, in the kidney exchange literature this name is already reserved to an algorithm that matches incoming pairs at random, without necessarily optimizing the cardinality of matchings at every period.} algorithm is one that finds the maximal number of matchings at every period, ignoring that it might be useful to keep certain pairs in the pool for later use.
-As a result, the graph at the end of every period has no cycles shorter than the maximal allowed cycle length, for they would have been already cleared.
-
-This algorithm also does not take into consideration any of the observables in the data.
-In particular, it makes no use of the fact that some pairs may be harder to match in the future, and therefore it might be advantageous to match them today if the opportunity arises.
-Conversely, it will not be less likely to carry out matchings involving easy-to-match pairs that could be postponed to the future at low expected cost.
-
-\begin{align}
-   &\max_{x} \sum_{c \in C} w_{c} x_{c} \\
-      s.t. &\sum_{v \in c'} x_{c'} = 1 \qquad \forall v \in V\\
-        &x_c \in \{ 0, 1\}  \qquad \forall c \in C
-\end{align}
+The \emph{myopic} algorithm finds the maximal number of matchings at every period, and clears it immediately.
+It disregards all observable characteristics of each pair, and in particular it disregards that some pairs might be useful to keep certain pairs may be easier or harder to match.
+In doing so, it may forgo the opportunity of matching a hard-to-match patient today, or postpone an easy-to-match pair for later.
 
@@
 The infeasible optimal algorithm (henceforth OPT) does away with the uncertainty arising from the temporal structure of the problem: it knows exactly which pairs will come into the pool, as well as their arrival, and the duration of their sojourns.
-In effect, this algorithm simply solves the static integer programming problem described at the beginning of section \ref{sec:literature:subsec:kidney}, except that the set of cycles has the additional constraint that a cycle cannot exist between two vertices if their sojourns fail to overlap.
+This algorithm is essentially a large static kidney matching problem, except that the set of available exchanges has the additional constraint that a cycle cannot exist between two vertices if their sojourns fail to overlap.
 
@@
 %
+
+\subsection{Algorithms}
+
+We divide our algorithms into three classes: \emph{direct prediction} methods, \emph{multi-armed bandit} methods and \emph{reinforcement learning} methods. They are explained next.
+
+\subsubsection{Direct prediction} \label{subsubsec:estimation}
+
+Dynamic and static algorithms can split their work: the former can determine which pairs should be matched at each period, while the latter decides how they should be matched among themselves.
+The \emph{direct prediction} exploits this insight, essentially reducing the problem to a classification task: at each period, they aim to produce a 0-1 label for each node indicating whether is should be matched in this period (1) or left for later (0).
+Selected nodes are then passed to a static solver that finds the maximal matching among them.
+Once these nodes are cleared, time evolves to the next period.
+This is illustrated in Figure \ref{fig:direct_prediction}.
+
+\begin{figure}
+\centering
+\hspace*{-1cm}
+\includegraphics[width=1.1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/direct_prediction.pdf}.
+\caption{Direct prediction methods.}
+\label{fig:direct_prediction}
+\end{figure}
+
+In order to produce data to feed into our ``classifier", each environment was simulated and solved using OPT, with each pair's observable characteristics as regressors, and a binary variable indicating whether they were matched or not.
+Each resulting artifical dataset of approximately 2 million observations was then fed to a series of predictive algorithms, including penalized logistic regression\cite{wu2009genome}, support vector machines with radial basis function kernels\cite{cortes1995support}, random forest classifiers\cite{breiman2001random}, gradient tree boosting classifiers\cite{friedman2001greedy} and deep feedforward networks\cite{goodfellow2016deep}.
+The performance of these algorithms as classification methods is compared in Table \ref{tab:direct_prediction}.
+
+[[[The poor performance?? could be attributable to not having represented the data properly etc. Non-euclidean spaces big challenge etc]]]
+
+ In order to represent information about the graph and how the nodes fit within it, we augment the original dataset using one or both of the following.
+First, vertex properties for each node such as the number of in and out-edges, average neighbor degree, some centrality measures (betweenness, in-degree, out-degree, harmonic, closeness) and more exotic indices such as the core number and pagerank of each node\cite{newman2010networks}.
+Second, using the \emph{node2vec} graph embedding algorithm of \cite{grover2016node2vec}, a continuous feature representation algorithm inspired by the skip-gram model.[[[More on this?]]].
+These algorithms will produce auxiliary matrices $E_t^{1}$ and $E_t^2$ that are concatenated to the original data regressor matrix $X_t$, and the entire augmented data set can again be fed to the predictive algorithms above.
+Their performance is compared on the bottom half of Table \ref{tab:direct_prediction}.
+
+ \subsection{Multi-armed bandits}
+
+ \begin{figure}
+ \centering
+ \hspace*{-1cm}
+ \includegraphics[width=1.1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/mab.pdf}.
+ \caption{Multi-armed bandit methods.}
+ \label{fig:mab}
+ \end{figure}
+
+ \begin{figure}
+ \centering
+ \includegraphics[width=1.1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/mab_step.pdf}.
+ \caption{Pseudo-reward computation in one step of a multi-armed bandit algorithm.}
+ \label{fig:mab_step}
+ \end{figure}
+
 % \subsection{Methods}
@@
 %
-
-\subsubsection{Estimation} \label{subsubsec:estimation}
-
-As explained in section \ref{subsubsec:embedding}, one of our main challenges is that we must deal not only with information about each patient-donor pair, but also
 

### 2018-02-05 paper.tex
First half of subsection on multi-armed bandit explanation

@@
 
-\usepackage{amssymb,amsmath,amsfonts,eurosym,geometry,ulem,graphicx,caption,color,setspace,sectsty,comment,footmisc,caption,pdflscape,subfigure,array,hyperref}
+\usepackage{amssymb,amsmath,amsfonts,eurosym,geometry,ulem,graphicx,caption,color,setspace,sectsty,comment,footmisc,pdflscape,subfigure,array,hyperref}
 \usepackage{cite}
@@
 \usepackage{booktabs}
+\usepackage[font=footnotesize]{caption}
 
@@
 
-% \subsubsection{Empirical comparison between myopic and OPT}
-% 
-% We simulated each environment for 3000 periods for each given environment and for each combination of entry rate, death rate and maximum available cycle length.
-Next, we analyzed the performance of OPT and myopic in each setup, as measured by the average number of pairs they were able to match per period.
-In order to avoid mixing problems, we discarded the first 1000 periods as burn-in.\footnote{While we refrain from a formal analysis of mixing time here, our thousand-period burn-in may be fairly conservative, as we empirically observe that for all our parameter combinations the size of the graph quick grows from zero to a stable level in (sometimes much) fewer than one hundred periods.}
-% 
-% The results are seen on Figure \ref{???}, where we show the performance ratio between the two algorithms.
-It is perhaps surprising that myopic's performance is often very close to OPT.
-In particular, when entry rates are relatively large or death rates are relatively low, the pool becomes thick enough that  opportunities for a reasonably good matching abound, so even a simple heuristic like myopic is able to do well.
-A bit less remarkably, high death rates also allow myopic to thrive, since dynamic considerations play a much smaller role when pairs leave the pool more often.
-% 
-% A more interesting remark is that in all of our three environments, when 3-cycles are allowed myopic does relative worse to OPT than when only 2-cycles are allowed. 
-%
+\subsubsection{Empirical comparison between myopic and OPT}
+
+Figure \ref{}
+
+ \begin{figure}
+ \centering
+ \hspace*{-4cm}
+ \includegraphics[width=1.5\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/greedy_opt_comparison_mcl2.pdf}
+ \caption{\texttt{Myopic} matches roughly the same number of pairs per period as \texttt{OPT} when the death rate is high (moving rightwards on the graphs), or when entry rate is high (moving downwards on the graphs).}
+ \label{fig:mab}
+ \end{figure}
 
@@
 
-\subsubsection{Direct prediction} \label{subsubsec:estimation}
+\subsubsection{Direct prediction} \label{subsubsec:direct_prediction}
 
@@
  In order to represent information about the graph and how the nodes fit within it, we augment the original dataset using one or both of the following.
-First, vertex properties for each node such as the number of in and out-edges, average neighbor degree, some centrality measures (betweenness, in-degree, out-degree, harmonic, closeness) and more exotic indices such as the core number and pagerank of each node\cite{newman2010networks}.
-Second, using the \emph{node2vec} graph embedding algorithm of \cite{grover2016node2vec}, a continuous feature representation algorithm inspired by the skip-gram model.[[[More on this?]]].
+First, with vertex properties for each node such as the number of in and out-edges, average neighbor degree, some centrality measures (betweenness, in-degree, out-degree, harmonic, closeness) and more exotic indices such as the core number and pagerank of each node\cite{newman2010networks}.
+Second, we augment the data using the \emph{node2vec} graph embedding algorithm of \cite{grover2016node2vec}, a continuous feature representation algorithm inspired by the skip-gram model.[[[More on this?]]].
 These algorithms will produce auxiliary matrices $E_t^{1}$ and $E_t^2$ that are concatenated to the original data regressor matrix $X_t$, and the entire augmented data set can again be fed to the predictive algorithms above.
@@
 
+We have access to the data-generating process (\emph{environments}) and an algorithm that is able to find the best matching for any data in hindsight (\texttt{OPT}).
+Therefore, in principle we could estimate the best choice of cycle to clear by simulating several periods ahead in the future, running \texttt{OPT}, and measuring the performance of our choice under any desired criterion.
+
+The approach outlined above turns out to be naive and incomplete, but it essentially contains the insight under which we will be working in this section.
+It is naive because simulations are computationally expensive, and in practice we cannot repeat them enough times get reliable estimates of the average performance for each cycle choice, especially in large graphs.
+It is also incomplete because it does not specify exactly what is the best information we should extract from \texttt{OPT} results.
+In order to solve the first problem, we leverage theory and algorithms from the multi-armed bandit (MAB) literature.
+For the second, we propose a secondary objective based on what we call \emph{pseudo-rewards}.
+
+Our procedure is illustrated in Figures \ref{fig:mab} and \ref{fig:mab_step}.
+At the beginning of the period $t$, the agent receives a set of cycles $C$ that are available to be cleared.
+If $C$ is empty, nothing happens and we move to the next period.
+Otherwise, the agent then picks a cycle $c \in C$ and simulates the future, including new entries and deaths, up to a horizon $h$.
+Next, \texttt{OPT} is run twice, once normally, and once with the additional constraint that $c$ be removed today.
+The size of the resulting matching in these two scenarios is compared.
+Naturally, the constrained version of \texttt{OPT} cannot achieve anything better than its unconstrained counterpart, but it might get to be equal.
+If it is, the agent receives a \emph{pseudo-reward} of one, otherwise it receives zero.
+This process is repeated: at each iteration $\ell$, a cycle $c_{\ell}$ is chosen and its pseudo-reward $r_{c, \ell}$ is revealed.
+When a preset computational budget of $L(|C|)$ iterations is hit, the agent then analyses the whole history of cycle choices and pseudo-rewards $H_{t} = \{ (c_{\ell}, r_{c\ell} ) \}_{\ell=1}^{L}$, and decides whether to match one of the cycles or move on to the next period.
+If a cycle $c$ is chosen it is immediately cleared, however the environment does not evolve to the next period yet.
+Instead, the history $H_t$ is discarded the procedure is repeated again with a reduced set of actions $C' \subset C$ that produces a new history $H_t'$ and so on, until either there are no more available choices or the agent decides to allow the environment to move on to time $t+1$.
+When that at last happens, entries and deaths are revealed, the agent receives a new set of cycles, and the process begin anew.
+All past information is ignored.
+
+Two important details were left out of the explanation above. First, how does the agent chooses the next cycle to test? Second, how does it decide which cycle to choose (or no cycle at all)?
+
+Let $c^{*}$ be a cycle that maximizes expected pseudo-rewards during one round of the algorithm:
+$$c^{*} \in \arg\max_{c} E[r_{c\ell}] $$
+
+ A \emph{multi-armed bandit} algorithm is a procedure to minimize \emph{regret}
+
  \begin{figure}
@@
  \hspace*{-1cm}
- \includegraphics[width=1.1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/mab.pdf}.
+ \includegraphics[width=1.1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/mab.pdf}
  \caption{Multi-armed bandit methods.}
@@
  \centering
- \includegraphics[width=1.1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/mab_step.pdf}.
+ \includegraphics[width=1.1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/mab_step.pdf}
  \caption{Pseudo-reward computation in one step of a multi-armed bandit algorithm.}

### 2018-02-08 paper.tex
New tables

@@
 
-\usepackage{amssymb,amsmath,amsfonts,eurosym,geometry,ulem,graphicx,caption,color,setspace,sectsty,comment,footmisc,pdflscape,subfigure,array,hyperref}
+\usepackage{amssymb,amsmath,amsfonts,eurosym,geometry,ulem,caption,color,setspace,sectsty,comment,footmisc,pdflscape,subfigure,array,hyperref}
+\usepackage{graphicx}
 \usepackage{cite}
@@
 
-%\doublespacing
+\doublespacing
 
@@
 
-\begin{table}[H]
-\centering
-\begin{tabular}{lr}
-\toprule
-Pair blood type &         Probability \\
-\midrule
-(O, O) &  0.058689 \\
-(O, A) &  0.373803 \\
-(O, B) &  0.158257 \\
-(O, AB) &  0.042669 \\
-(A, O) &  0.041119 \\
-(A, A) &  0.028809 \\
-(A, B) &  0.110888 \\
-(A, AB) &  0.029899 \\
-(B, O) &  0.017410 \\
-(B, A) &  0.110888 \\
-(B, B) &  0.005160 \\
-(B, AB) &  0.012660 \\
-(AB, O) &  0.004690 \\
-(AB, A) &  0.003290 \\
-(AB, B) &  0.001390 \\
-(AB, AB) &  0.000380 \\
-\bottomrule
-\end{tabular}
-\caption{Blood type probabilities in the ABO environment}
-\label{tab:abo_blood_type}
-\end{table}
-
 \subsection{RSU Environment}
@@
 \centering
-\includegraphics[width=1textwidth]{/Users/vitorhadad/Documents/kidney/matching/figures/computed_cpra.pdf}.
+\includegraphics[width=\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/computed_cpra.pdf}
 \caption{Computed patient and donor cPRA.}
@@
 
-Figure \ref{}
+Our benchmark
 
- \begin{figure}
+ \begin{figure}[H]
  \centering
@@
  \caption{\texttt{Myopic} matches roughly the same number of pairs per period as \texttt{OPT} when the death rate is high (moving rightwards on the graphs), or when entry rate is high (moving downwards on the graphs).}
- \label{fig:mab}
+ \label{fig:greedy_opt_comparison_mcl2}
  \end{figure}
 
-\subsection{Algorithms}
+\subsection{Methods}
 
-We divide our algorithms into three classes: \emph{direct prediction} methods, \emph{multi-armed bandit} methods and \emph{reinforcement learning} methods. They are explained next.
+We divide our algorithms into three classes: \emph{direct prediction} methods, \emph{multi-armed bandit} methods. As we will explain shortly,
 
@@
 Dynamic and static algorithms can split their work: the former can determine which pairs should be matched at each period, while the latter decides how they should be matched among themselves.
-The \emph{direct prediction} exploits this insight, essentially reducing the problem to a classification task: at each period, they aim to produce a 0-1 label for each node indicating whether is should be matched in this period (1) or left for later (0).
+The \emph{direct prediction} exploits this insight, essentially reducing the problem to a classification task: at each period, we aim to produce a 0-1 label for each node indicating whether is should be matched in this period (1) or left for later (0).
 Selected nodes are then passed to a static solver that finds the maximal matching among them.
@@
 
-\begin{figure}
+\begin{figure}[H]
 \centering
@@
 
-In order to produce data to feed into our ``classifier", each environment was simulated and solved using OPT, with each pair's observable characteristics as regressors, and a binary variable indicating whether they were matched or not.
-Each resulting artifical dataset of approximately 2 million observations was then fed to a series of predictive algorithms, including penalized logistic regression\cite{wu2009genome}, support vector machines with radial basis function kernels\cite{cortes1995support}, random forest classifiers\cite{breiman2001random}, gradient tree boosting classifiers\cite{friedman2001greedy} and deep feedforward networks\cite{goodfellow2016deep}.
+In order to produce data to feed into our ``classifier", we simulated and solved a 1000-period run environment was simulated and solved using OPT using with each each node's observable characteristics as regressors, and a binary variable indicating whether they were matched or not as the regressand.
+The resulting artificial dataset of approximately 2 million observations was then fed to a series of predictive algorithms, including penalized logistic regression\citep{wu2009genome}, support vector machines with radial basis function kernels\citep{cortes1995support}, random forest classifiers\citep{breiman2001random}, gradient tree boosting classifiers\citep{friedman2001greedy} and deep feedforward networks\citep{goodfellow2016deep}.
 The performance of these algorithms as classification methods is compared in Table \ref{tab:direct_prediction}.
@@
 Let $c^{*}$ be a cycle that maximizes expected pseudo-rewards during one round of the algorithm:
-$$c^{*} \in \arg\max_{c} E[r_{c\ell}] $$
+$$c_{\ell}^{*} \in \arg\max_{c} E[r_{c,\ell}] $$
 
- A \emph{multi-armed bandit} algorithm is a procedure to minimize \emph{regret}
+ Also, let \emph{regret} be defined as the difference between expected reward of the optimal choice $c^*$ and its own selected choice $c_{\ell}$.
+
+ $$\Delta_\ell := E[r_{c^{*},\ell}] - E[r_{c,\ell}]$$
+
+ A \emph{multi-armed bandit} (MAB) algorithm is a procedure to minimize the cumulative regret over $L$ rounds.
+A good MAB algorithm will act so as to balance exploration (trying out different choices to get high-quality estimates of their rewards) and exploitation (using out better choices more often to increase total rewards).
+
+ The literature on bandit algorithms is extensive and spans at least seventy years of research \citep{lattimore2018bandits}. Here we will focus on binary
 
@@
  \includegraphics[width=1.1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/mab_step.pdf}
- \caption{Pseudo-reward computation in one step of a multi-armed bandit algorithm.}
+ \caption{Pseudo-reward computation that takes place during one round of a multi-armed bandit algorithm.}
  \label{fig:mab_step}
@@
 
-% \subsection{Methods}
-% 
-% In this work, we propose a new kidney exchange algorithm that combines simulation-based search with a policy parametrized by neural networks.
-In the next paragraphs, we will explain each of these components in detail.
-However, first let us provide some motivation.
-% 
-% We begin by noting that we have a promising setup, as we are in possession of: good estimates of the data-generating process for each environment; an algorithm that is able to compute optimal actions, even if only in hindsight (\texttt{OPT}); one clear benchmark that we must beat (\texttt{Myopic}).
-At first glance, then, one might hypothesize that it would be possible to find reasonably good solutions using only simulations: simply simulate future states to a long horizon ahead, and choose today the action that on average leads to higher future rewards -- or the one that is most often chosen by \texttt{OPT}, or the one that most frequently leads to beating \texttt{Myopic}.
-Unfortunately, this turns out to be overly optimistic: approaches solely based on simulations do not perform very well, as shown in Table \ref{simulations_only}.
-We believe that the reason for this poor performance is the high branching factor of the problem -- in other words, there is high uncertainty about what the next period will look like.
-% 
-% 
-% Reducing the problem:
-% 
-% 1. Our "actions" are simply a number of binary flags indicating who should be matched today and who should not
-% 2. We can leverage the optimal algorithm to reduce the MDP to a simpler period-by-period MAB problem
-% 3. Use 
-% 
-% % In lieu of just simulations, we propose to see each period's problem as a Bayesian multi-armed bandit problem (MAB).
-...
-estimate...
-then use estimates as pseudo-counts...
-then sample from posterior.
-% 
-%
-
-%\insert{algo1.tex}
-
-% Using Thompson sampling at each episode might sound like a greedy approach to a problem where current actions might have effects on future rewards, but here we *are* using future rewards since we can simulate the future \emph{and} also assume that we will be act optimally within a time horizon.
-
 \section{Results} \label{sec:results}
 
-\subsection{Short-run gains}
+\subsection{Direct method}
+
+\subsection{Multi-armed bandits}
 
@@
 
-\bibliographystyle{/Users/vitorhadad/Documents/kidney/matching/paper/te.bst}
+\bibliographystyle{te.bst}
 \bibliography{references}
@@
 \addcontentsline{toc}{section}{Tables}
+
+\begin{table}[H]
+\centering
+\singlespacing
+\begin{tabular}{lr}
+\toprule
+Pair blood type &         Probability \\
+\midrule
+(O, O) &  0.058689 \\
+(O, A) &  0.373803 \\
+(O, B) &  0.158257 \\
+(O, AB) &  0.042669 \\
+(A, O) &  0.041119 \\
+(A, A) &  0.028809 \\
+(A, B) &  0.110888 \\
+(A, AB) &  0.029899 \\
+(B, O) &  0.017410 \\
+(B, A) &  0.110888 \\
+(B, B) &  0.005160 \\
+(B, AB) &  0.012660 \\
+(AB, O) &  0.004690 \\
+(AB, A) &  0.003290 \\
+(AB, B) &  0.001390 \\
+(AB, AB) &  0.000380 \\
+\bottomrule
+\end{tabular}
+\caption{Blood type probabilities in the ABO environment}
+\label{tab:abo_blood_type}
+\end{table}
+
+\begin{table}
+  \centering
+  \singlespacing
+  \hspace*{-3cm}
+  \input{tables/traditional_ml.tex}
+  \caption{\textbf{Traditional methods} \\ Performance of traditional econometric and classification methods when predicting whether a node will be matched (1) or not (0).
+  Recall is true positive rate $\frac{tp}{tp+fn}$;
+  Specificity is true negative rate $\frac{tp}{tn+fp}$;
+  Precision is $\frac{tp}{tp+fp}$; Accuracy is $\frac{tp+tn}{tp+tn+fp+fn}$. See section \ref{sec:???} for methodology and section \ref{sec:??} for discussion of results.
+  }
+  \label{tab:tradional_ml}
+\end{table}
 

### 2018-02-11 ec-18/kidney-ec-submission.tex
Added algorithms

\documentclass[format=acmsmall, review=false]{acmart}

\usepackage{acm-ec-18}

\usepackage{booktabs} % For formal tables

\usepackage[ruled]{algorithm2e} % For algorithms

\renewcommand{\algorithmcfname}{ALGORITHM}
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetAlCapHSkip{0pt}
\IncMargin{-\parindent}

\begin{document}
% Title portion. Note the short title for running heads 
\title[New strategies for improving kidney exchange]{New strategies for improving kidney exchange}  
\author{Submission 42}

% note that the abstract must come before \maketitle
\begin{abstract}
[[Kidney exchange is great, provides a way to increase number of lives saved]] [[Most algorithms static, don't think about future]] [[There might be ways to improve, but literature still small]] [[This work adds two contributions to the literature]] [[Direct prediction method, MAB-based methods]] [[The latter can improve the average number of kidneys exchange by 3\% over the myopic algorithm]]
\end{abstract}

% note: this command has been disabled to remove the ACM copyright block. Sorry...
%\thanks{This work is supported by the National Science Foundation,
%  under grant CNS-0435060, grant CCR-0325197 and grant EN-CS-0329609.}

\maketitle

\section{Introduction}

[[Kidney exchange]]

[[Will interchangeably use the words 'node', 'vertex' and 'pair']]

\paragraph{\textbf{Main idea}} We propose and compare two new approaches to improve the number of matching in kidney exchange, namely the \emph{direction estimation} method and the \emph{multi-armed-bandit} method. The former recasts the dynamic kidney exchange problem as a binary classification task: for each pair in the pool, we should predict a 0-1 label representing whether they should be matched today, or left for the future. The latter chooses exchanges that are less likely to produce ``regret" in the future (under a specific definition of regret that we elaborate on later).

Each approach faces different challenges. In the \emph{direct estimation} approach, one challenge is that an important part of our data -- the compatibility graph -- lives in a non-Euclidean space, and it is not obvious how to use it as an input to a classification algorithm. We attempt to apply a variety of new methods from the graphical signal processing literature to tackle this problem. An additional challenge is that traditional classification algorithms are not equipped to produce dependent labels, but in our setting whether or not we decide that a certain pair is to be matched today ought to be related to our decisions about everyone else in the pool. To overcome this we use recurrent neural networks and draw from the [[[[]]]]

In the \emph{multi-armed bandit} method, the main issue is that the enormous branching factor of our problem precludes an accurate estimation of regret for each possible exchange. Here, we develop a novel algorithm that iteratively chooses to match one additional pair or decides

\paragraph{\textbf{Main results}} We conduct simulations following three models from the kidney exchange literature, which we call \emph{environments}. The three environments differ by their rules regarding blood- and tissue-type compatibility.

We apply our methods to each of these environments and compare them to a myopic algorithm (representing how kidney exchanges are currently conducted), and to infeasible optimal algorithm. Comparison is made in terms of average number of pairs matched per period, and the average is taken over a long period of time (i.e., much longer a node's mean sojourn).

In all environments, we observe that the \emph{direct estimation} unfortunately does equally well or worse than the myopic algorithm. However, the \emph{multi-armed bandit} method is able to improve upon it by 0.5-5\% in terms of the number of average pairs matched per period. Moreover, we observe that there is still a moderately large performance gap between our algorithms and the infeasible optimal.

\subsection{Related literature} This paper pulls ideas from microeconomics and matching theory, and uses methods drawn from computer science, machine learning, and operations research. Let us take a brief look at these fields in turn.

\paragraph{Dynamic kidney exchange}

\paragraph{Computer science}

%%%%%%%%%
\section{Simulation environments}

\paragraph{Notation}

\subsection{Common features across environments}

\paragraph{Poisson entry, Geometric death} At time point, the number of new incoming pairs is drawn from the $Poisson(r)$ distribution, where $r \in \mathbb{N}$ denotes the \emph{entry rate}, and equals the expected number of entrants per period. Upon entrance, each pair independently draws the length of their sojourn from the $Geometric(d)$ distribution. The parameter $d \in \mathbb{R}$ is the \emph{death rate}, and its reciprocal $\frac{1}{d}$ is the expected sojourn length. We note that, due to the memoryless property\footnote{If $X \sim Geometric(p)$, then $P(X > t+s | X > s) = P(X > t)$} of the Geometric distribution, the amount of time a pair has waited in the pool gives us no information about how much time they have until their death.

\paragraph{Minimum set of observable characteristics} In every environment, at least two characteristics are relevant and observable: the major blood type groups (A, B, O or AB) of the patient and donors, and how long the pair has been in the pool. On the other hand, death times are never observed.

\paragraph{Binary preferences} In this work we abstract from these concerns and consider every exchange to be equally desirable, so long as it is available. This is the same assumption as used in \cite{roth2005pairwise},

\subsection{ABO Environment}

In the \emph{ABO environment}, a pair's observable characteristics are their current waiting time and dummies representing ABO blood types of its patient and donor. Compatibility between two distinct pairs is based only on blood-type compatibility. Blood types are drawn independently for patient and donor from the US population (roughly O:48\%, A:36\%, B:11\%, AB:4\%). In addition, we make an assumption previously used in \citet{unver2010dynamic} and allow for incompatibility between a donor and their own patient. The reason for this additional assumption is that, if there were truly no tissue type compatibilities, we would never observe pairs of type (AB,$\cdot$), ($\cdot$, O), or (A,A), (O,O), (B,B), and (AB,AB), since their donors would be automatically compatible with their patients and they would never participate in an exchange. Following \citet{zenios2001primum}, we assume that for such patients the probability that a donor and their patient are incompatible is $p_c = 0.11$. Arrival rates are adjusted accordingly, e.g., the arrival rate of $(A,O)$ pair is proportional to $0.48 \times 0.36 \times 0.11 \approx 0.019$.

\subsection{RSU Environment}

The \emph{RSU} environment drawn from the setup in \cite{roth2007efficient}. Observable covariates are dummies representing patient and donor ABO blood types, current waiting time, and a \emph{calculated panel reactive antibody (cPRA)} level that governs the probability of a crossmatch with a random donor.\footnote{The original \cite{roth2007efficient} paper called this simply \emph{PRA}. In real life there is an important distinction between the two measures, but for the purposes of a simulation model this distinction is immaterial. We keep the name \emph{cPRA} for consistency with OPTN environment.} The lower the cPRA, the higher the number of potentially compatible pairs.

The simulation process is as follows. First, we draw a pair in the same manner as in the ABO environment. Next, we draw if the patient is a female (with probability around 41\%), and if so we also draw whether her donor is her husband (spouses comprise about 49\% of donors). Finally, we draw a  cPRA level for the patient (Low: 70.1\%, Medium: 20\%, High: 9.9\%). This cPRA level determines the probability that they can receive a kidney from any donor, including their own: patients with low cPRA have a 5\% probability of positive crossmatch with a random donor; patients with medium cPRA have a 45\% chance, and patients with high cPRA have a 90\% chance of a crossmatch. If a patient is bloody or tissue-type incompatible with their own donor, they enter the pool. In addition, if the patient is female and her husband is the donor, the probability of positive crossmatch for low, medium and high cPRA patients goes up to 28.75\%, 58.75\% and 92.25\%. This last adjustment reflects the fact that women tend to produce antibodies against their husbands' antigens during pregnancy.

Once in the pool, the pair immediately forms directed edges with the existing pairs, again following the patient cPRA distribution. The resulting random graph is akin to a Erd\"{o}s-R\'{e}nyi $G(n,p)$ random graph where the probability of forming edges is heterogeneous across different pair types.

\subsection{OPTN Environment}

In the \emph{OPTN environment}, we use historical data collected by the United Network for Organ Sharing (UNOS) data provided in the Standard Research and Analysis (STAR) dataset. The STAR dataset contains information from all patients that were ever registered to the kidney waiting list in the United States for the past three decades, as well as from living donors that participated in an transplant. From this original dataset, we excluded patients that were registered for more than one organ (including kidney+pancreas), patients that were not waiting for their first kidney transplant, and patients and donors with incomplete HLA profile information. The resulting historical dataset still contained 117813 patients and 9337 living donors.

\paragraph{Patient and donor cPRA} We added two additional variables to the original dataset: a \emph{patient cPRA} and a \emph{donor cPRA}. While the patient cPRA is a measure of patient tissue-type incompatibility with a random donor \cite{cecka2010calculated}, our \meph{donor cPRA} is a novel measure of the opposite direction -- how frequently a donor is tissue-type incompatible with a random patient.\footnote{We thank Itai Ashlagi for the suggestion of a donor cPRA.} Both were computed empirically: for each patient our dataset, we checked the how many donors exhibited antigens that are unacceptable for the patient in any of the A, B, Bw, C, DR, DPB, DQ, and DQA loci, and assigned this positive crossmatch probability as their patient cPRA; for the donors, we worked in the opposite direction by calculating the frequency of patients who exhibited antibodies against their donor's antigens, and that became their donor cPRA. Figure \ref{fig:cpra} shows the distribution across the entire population.\footnote{We should remark that we found a very different patient cPRA distribution than the one in the OPTN dataset. This may have been because: in real life cPRA is computed using deceased donor data, whereas we use living donor data; we have used different HLA equivalence tables; OPTN uses a more sophisticated model based on population genetics.\cite{optn2013cpra}}

\begin{figure}
\centering
\includegraphics[width=\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/computed_cPRA.pdf}
\caption{Patient and donor cPRA computed from historical data.}
\label{fig:cpra}
\end{figure}

\paragraph{Artificial dataset} Next, we created an artificial dataset by randomly drawing patients and living donors from the historical dataset and checking for blood-type compatibility, and tissue-type compatibility as explained above. Compatible pairs were discarded. We iterated in this manner to construct a dataset of about one million incompatible pairs. Its columns are a set of dummies representing patient and donor blood times, a set of dummies for each allele in each HLA locus, and

At every simulation period, a random number of pairs are drawn from this dataset.

[[[[Algorhtm can still see original alleles. Patient and donor cPRA offer "long-run" measure of compatibility]]]]

%%%%%%%%
\section{Methods}

\subsection{Objective and benchmarks}

In this work, we focus on maximizing the average number of matched pairs over $T$ periods, where $T$ is a large number relative to the rates of patient entry and death. We are interested in how our new methods perform relative to the following two benchmarks.

\paragraph{\textsc{Myopic}}

This algorithm finds the maximal matchings at every period, and clears it immediately. It disregards all observable characteristics of each pair, and in particular it disregards that some pairs might be useful to keep certain pairs may be easier or harder to match. In doing so, it may forgo the opportunity of matching a hard-to-match patient today, or postpone an easy-to-match pair for later. In essence, this is an approximation to what kidney exchanges currently do.

\paragraph{Optimal (\textsc{OPT})} This infeasible algorithm (henceforth OPT) does away with the uncertainty arising from the temporal structure of the problem: it knows exactly which pairs will come into the pool, as well as their arrival, and the duration of their sojourns. This algorithm is essentially a large static kidney matching problem, except that the set of available exchanges has the additional constraint that a cycle cannot exist between two vertices if their sojourns fail to overlap.

\textbf{Remark} \textit{How much is there to be improved upon?} [[]]

\begin{figure}[H]
\centering
\hspace*{-3.5cm}
\includegraphics[width=1.4\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/greedy_opt_comparison_mcl2_2.pdf}
\caption{Ratio of \texttt{Myopic} has better chances of achieving performances similar to \texttt{OPT} when the death rate is high (moving rightwards on the graphs), or when entry rate is high (moving downwards on the graphs).}
\label{fig:greedy_opt_compariso}
\end{figure}

\section{Algorithms}

We present two new methods to determine which patients should be matched.

\subsection{Direct prediction} \label{subsec:direct_prediction}

Dynamic and static algorithms can work in tandem: the former can determine which pairs should be matched at each period, while the latter decides how they should be matched among themselves. The \emph{direct prediction} exploits this insight, essentially reducing the problem to a classification task: at each period, we aim to produce a binary label for each node indicating whether is should be matched in this period (1) or left for later (0). Selected nodes are then passed to a static solver that finds the maximal matching among them. Once these nodes are cleared, time evolves to the next period. This is illustrated in Figure \ref{fig:direct_prediction}.

\begin{figure}[H]
\centering
\hspace*{-1cm}
\includegraphics[width=1.1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/direct_prediction.pdf}.
\caption{Direct prediction methods.}
\label{fig:direct_prediction}
\end{figure}

\section{Results}

\subsection{Direct estimation policy}

In order to produce data to feed into our ``classifier", we simulated and solved a 1000-period run environment was simulated and solved using OPT using with each each node's observable characteristics as regressors, and a binary variable indicating whether they were matched or not as the regressand. The resulting artificial dataset of approximately 2 million observations was then fed to a series of predictive algorithms, including penalized logistic regression\citep{wu2009genome}, support vector machines with radial basis function kernels\citep{cortes1995support}, random forest classifiers\citep{breiman2001random}, gradient tree boosting classifiers\citep{friedman2001greedy} and deep feedforward networks\citep{goodfellow2016deep}. The performance of these algorithms as classification methods is compared in Table \ref{tab:direct_prediction} in the Supplementary Material section.

\subsection{MAB-based policy}

We have access to the data-generating process (\emph{environments}) and an algorithm that is able to find the best matching for any data in hindsight (\texttt{OPT}). Therefore, in principle we could estimate the best choice of cycle to clear by simulating several periods ahead in the future, running \texttt{OPT}, and measuring the performance of our choice under any desired criterion.

The approach outlined above turns out to be naive and incomplete, but it essentially contains the insight under which we will be working in this section. It is naive because simulations are computationally expensive, and in practice we cannot repeat them enough times get reliable estimates of the average performance for each cycle choice, especially in large graphs. It is also incomplete because it does not specify exactly what is the best information we should extract from \texttt{OPT} results. In order to solve the first problem, we leverage theory and algorithms from the multi-armed bandit (MAB) literature. For the second, we propose a secondary objective based on what we call \emph{pseudo-rewards}.

Our procedure is illustrated in Figures \ref{fig:mab} and \ref{fig:mab_step}. At the beginning of the period $t$, the agent receives a set of cycles $C$ that are available to be cleared. If $C$ is empty, nothing happens and we move to the next period. Otherwise, the agent then picks a cycle $c \in C$ and simulates the future, including new entries and deaths, up to a horizon $h$. Next, \texttt{OPT} is run twice, once normally, and once with the additional constraint that $c$ be removed today. The size of the resulting matching in these two scenarios is compared. Naturally, the constrained version of \texttt{OPT} cannot achieve anything better than its unconstrained counterpart, but it might get to be equal. If it is, the agent receives a \emph{pseudo-reward} of one, otherwise it receives zero. This process is repeated: at each iteration $\ell$, a cycle $c_{\ell}$ is chosen and its pseudo-reward $r_{c, \ell}$ is revealed. When a preset computational budget of $L(|C|)$ iterations is hit, the agent then analyses the whole history of cycle choices and pseudo-rewards $H_{t} = \{ (c_{\ell}, r_{c\ell} ) \}_{\ell=1}^{L}$, and decides whether to match one of the cycles or move on to the next period. If a cycle $c$ is chosen it is immediately cleared, however the environment does not evolve to the next period yet. Instead, the history $H_t$ is discarded the procedure is repeated again with a reduced set of actions $C' \subset C$ that produces a new history $H_t'$ and so on, until either there are no more available choices or the agent decides to allow the environment to move on to time $t+1$. When that at last happens, entries and deaths are revealed, the agent receives a new set of cycles, and the process begin anew. All past information is ignored.

Two important details were left out of the explanation above. First, how does the agent chooses the next cycle to test? Second, how does it decide which cycle to choose (or no cycle at all)?

Let $c^{*}$ be a cycle that maximizes expected pseudo-rewards during one round of the algorithm:
$$c_{\ell}^{*} \in \arg\max_{c} E[r_{c,\ell}] $$

 Also, let \emph{regret} be defined as the difference between expected reward of the optimal choice $c^*$ and its own selected choice $c_{\ell}$.

 $$\Delta_\ell := E[r_{c^{*},\ell}] - E[r_{c,\ell}]$$

 A \emph{multi-armed bandit} (MAB) algorithm is a procedure to minimize the cumulative regret over $L$ rounds. A good MAB algorithm will act so as to balance exploration (trying out different choices to get high-quality estimates of their rewards) and exploitation (using out better choices more often to increase total rewards).

 The literature on bandit algorithms is extensive and spans at least seventy years of research \citep{lattimore2018bandits}. Here we will focus on binary

 \begin{figure}
  \centering
  \hspace*{-1cm}
  \includegraphics[width=1.1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/mab.pdf}
  \caption{\textbf{Function} \textsc{choose\_cycle}}
  \label{fig:mab}
  \end{figure}

  \begin{figure}
  \centering
  \includegraphics[width=1.1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/mab_step.pdf}
  \caption{\textbf{Function} \textsc{get\_pseudo\_reward} }
  \label{fig:get_pseudo_reward}
  \end{figure}

\section{Conclusion}

\paragraph{Summary}

\paragraph{Extensions} In reality, some exchanges are more or less desirable than others, either for ethical concerns or because of predicted health benefit. For example, it is common for pediatric patients and previous organ donors receive higher priority, as do exchanges involving patients with no HLA mismatch.

\newcommand\mycommfont[1]{\footnotesize\upshape\textcolor{blue}{#1}}
\SetCommentSty{mycommfont}

% Algorithm
\begin{algorithm}[t]
	\SetAlgoLined
	\KwData{Cycle $c$; Horizon $h$; \\
          Lists pseudo-reward statistics $Avg$, $Std$; \\
          Environment simulator object \textsc{Env}; \\
          Oracle solver \textsc{OPT}; }
  \DontPrintSemicolon
  \SetKwFunction{FChoose}{\textsc{get\_pseudo\_reward}}
  \SetKwProg{Fn}{Function}{:}{}
  \Fn{\FChoose{\textsc{Env}, \ $c$, \ $Avg$, \ $Std$, \ $h$ }}{
    \tcp{Simulate up to horizon $h$ and find optimal matching} \\
    \textsc{Env.simulate}($h$)  \\
    $r_1$ \gets \textsc{OPT.solve}(\textsc{Env})  \\
    \tcp{Remove cycle $c$, find constrained optimal matching} \\
    \textsc{Env.remove}($c$)  \\
    $r_2$ \gets \textsc{OPT.solve}(\textsc{Env}) \\
    \tcp{Return 1 if rewards are equal, 0 otherwise} \\
    \KwRet $r_1 == r_2$ \;
  }
	\caption{Function \textsc{get\_pseudo\_reward}}
	\label{alg:get_pseudo_reward}
\end{algorithm}

\begin{algorithm}[t]

  \SetAlgoLined
  \DontPrintSemicolon

  \KwData{Horizon $h$; Number of iterations $L$; Threshold $thres$; \\
    Environment simulator object \textsc{Env};
    Multi-armed bandit algorithm object \textsc{MAB};
  }

  \SetKwFunction{FChoose}{\texttt{choose\_cycle}}
  \SetKwProg{Fn}{Function}{:}{}

  \Fn{\FChoose{\texttt{Env},\ \texttt{h} }}{
    \tcp{Initialize lists of current pseudo-reward statistics} \\
    $C$ \gets \textsc{Env.get\_available\_cycles()} \\
    $Avg$ \gets \textsc{zeros(length(C))} \\
    $Std$ \gets \textsc{zeros(length(C))} \\

    \tcp{Begin iterations} \\
    \For{$i\gets0$ \KwTo $L$}{
      \tcp{Bandit algorithm chooses next cycle to test given statistics} \\
      $c$ \gets \textsc{MAB.pull($C$, Avg, Std)}  \\
      \tcp{Compute pseudo reward for this cycle and update statistics} \\
      $r$ \gets \textsc{get\_pseudo\_reward}(\textsc{Env},\ $c$,\ $Avg$,\ $Std$,\ $h$) \\
      $Avg$ \gets \textsc{update\_running\_average($Avg$, $r$)} \\
      $Std$ \gets \textsc{update\_running\_std($Std$, $r$)} \\ 
    }  
    \tcp{Bandit algorithm chooses best cycle given statistics} \\
    $c\_best$ \gets \textsc{MAB.choose($C$, Avg, Std)} \\
    \tcp{Return best cycle, unless none of the pseudo-reward averages are above a certain threshold} \\
    \eIf{\textsc{All}($Avg \leq thres$)} {
      \KwRet \text{NULL} \\ 
      }{
      \KwRet $c$ \\
      }
    }
  }
	\caption{Function \texttt{choose\_cycle}}
	\label{alg:choose_cycle}
\end{algorithm}

\section{Conclusions}

% Appendix
\appendix
\section{Sections}

\section{Supplementary materials}

\begin{acks}

% The author would like to thank Utku \"{U}nver, Stefan Hoderlein, Arthur Lewbel, Mohammad Akbarpour, Itai Ashlagi, David Parkes, John Dickerson, Enkhmanlai Amarsaikhan and Sainbayar Sukhbaatar for helpful discussions.

The author would like to thank Itai Ashlagi for the idea of including a donor cPRA in the OPTN environment.

This work was supported in part by Health Resources and Services Administration contract 234-2005-37011C. The content is the responsibility of the authors alone and does not necessarily reflect the views or policies of the Department of Health and Human Services, nor does mention of trade names, commercial products, or organizations imply endorsement by the U.S. Government.

\end{acks}

% Bibliography
\bibliographystyle{ACM-Reference-Format}
\bibliography{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/references}

\end{document}

### 2018-02-11 ec-18/sample-ec-submission-simple.tex
Added algorithms

\documentclass[format=acmsmall, review=false]{acmart}

\usepackage{acm-ec-18}
\usepackage{booktabs} % For formal tables

\begin{document}
% Title portion. Note the short title for running heads 
\title[A VCG-Based Mechanism for Electronic Commerce]{A VCG-Based Mechanism for Electronic Commerce Applications}  
\author{Submission 42}

% note that the abstract must come before \maketitle
\begin{abstract}
In auction theory, a Vickrey--Clarke--Groves (VCG) auction is a type of sealed-bid auction of multiple items. Bidders submit bids that report their valuations for the items, without knowing the bids of the other people in the auction. The auction system assigns the items in a socially optimal manner: it charges each individual the harm they cause to other bidders. It also gives bidders an incentive to bid their true valuations, by ensuring that the optimal strategy for each bidder is to bid their true valuations of the items. (Thanks for this text, Wikipedia.) This paper shows that the same idea works for electronic commerce applications.
\end{abstract}

\maketitle

% Renew this after \maketitle if the default list of authors is too long for headers
%\renewcommand{\shortauthors}{W.\ Vickrey et.\ al.}

\section{Introduction}

From here on it's straightforward \LaTeX. You should know that the style uses Natbib author--year citations. 
That means citations that work like footnotes use ``cite'' \cite{Abril07} and citations that work like a part of speech
are cited as we do \citet{Abril07}.

\begin{acks}

This is where we put acknowledgements.
\end{acks}

% Bibliography
\bibliographystyle{ACM-Reference-Format}
\bibliography{sample-bibliography}

\end{document}

### 2018-02-11 ec-18/sample-ec-submission.tex
Added algorithms

\documentclass[format=acmsmall, review=false]{acmart}

\usepackage{acm-ec-18}

\usepackage{booktabs} % For formal tables

\usepackage[ruled]{algorithm2e} % For algorithms
\renewcommand{\algorithmcfname}{ALGORITHM}
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetAlCapHSkip{0pt}
\IncMargin{-\parindent}

\begin{document}
% Title portion. Note the short title for running heads 
\title[A VCG-Based Mechanism for Electronic Commerce]{A VCG-Based Mechanism for Electronic Commerce Applications}  
\author{Submission 42}

% note that the abstract must come before \maketitle
\begin{abstract}
In auction theory, a Vickrey--Clarke--Groves (VCG) auction is a type of sealed-bid auction of multiple items. Bidders submit bids that report their valuations for the items, without knowing the bids of the other people in the auction. The auction system assigns the items in a socially optimal manner: it charges each individual the harm they cause to other bidders. It also gives bidders an incentive to bid their true valuations, by ensuring that the optimal strategy for each bidder is to bid their true valuations of the items. (Thanks for this text, Wikipedia.) This paper shows that the same idea works for electronic commerce applications.
\end{abstract}

% note: this command has been disabled to remove the ACM copyright block. Sorry...
%\thanks{This work is supported by the National Science Foundation,
%  under grant CNS-0435060, grant CCR-0325197 and grant EN-CS-0329609.}

\maketitle

\section{Introduction}

OK, I can't be bothered to make the rest of this be about VCG. What follows is the ACM's demo article.

As a new technology, Wireless Sensor Networks (WSNs) has a wide
range of applications \cite{Culler-01, Bahl-02, Akyildiz-01}, including
environment monitoring, smart buildings, medical care, industrial and
military applications. Among them, a recent trend is to develop
commercial sensor networks that require pervasive sensing of both
environment and human beings, for example, assisted living
\cite{Akyildiz-02, Harvard-01,CROSSBOW} and smart homes
\cite{Harvard-01, Adya-01,CROSSBOW}.
% quote
\begin{quote}
	``For these applications, sensor devices are incorporated into human
	cloths \cite{Natarajan-01, Zhou-06, Bahl-02, Adya-01} for monitoring
	health related information like EKG readings, fall detection, and
	voice recognition''.
\end{quote}
While collecting all these multimedia information
\cite{Akyildiz-02} requires a high network throughput, off-the-shelf
sensor devices only provide very limited bandwidth in a single
channel: 19.2\,Kbps in MICA2 \cite{Bahl-02} and 250\,Kbps in MICAz.

In this article, we propose MMSN, abbreviation for Multifrequency
Media access control for wireless Sensor Networks. The main
contributions of this work can be summarized as follows.
% itemize
\begin{itemize}
	\item To the best of our knowledge, the MMSN protocol is the first
	multifrequency MAC protocol especially designed for WSNs, in which
	each device is equipped with a single radio transceiver and
	the MAC layer packet size is very small.
	\item Instead of using pairwise RTS/CTS frequency negotiation
	\cite{Adya-01, Culler-01, Tzamaloukas-01, Zhou-06},
	we propose lightweight frequency assignments, which are good choices
	for many deployed comparatively static WSNs.
	\item We develop new toggle transmission and snooping techniques to
	enable a single radio transceiver in a sensor device to achieve
	scalable performance, avoiding the nonscalable ``one
	control channel + multiple data channels'' design \cite{Natarajan-01}.
\end{itemize}

% Head 1
\section{MMSN Protocol}

% Head 2
\subsection{Frequency Assignment}

We propose a suboptimal distribution to be used by each node, which is
easy to compute and does not depend on the number of competing
nodes. A natural candidate is an increasing geometric sequence, in
which
% Numbered Equation
\begin{equation}
\label{eqn:01}
P(t)=\frac{b^{\frac{t+1}{T+1}}-b^{\frac{t}{T+1}}}{b-1},
\end{equation}
where $t=0,{\ldots}\,,T$, and $b$ is a number greater than $1$.

In our algorithm, we use the suboptimal approach for simplicity and
generality. We need to make the distribution of the selected back-off
time slice at each node conform to what is shown in
Equation~\eqref{eqn:01}. It is implemented as follows: First, a random
variable $\alpha$ with a uniform distribution within the interval $(0,
1)$ is generated on each node, then time slice $i$ is selected
according to the following equation:
% Unnumbered Equation
\[
i=\lfloor(T+1)\log_b[\alpha(b-1)+1]\rfloor.
\]
It can be easily proven that the distribution of $i$ conforms to Equation
(\ref{eqn:01}).

So protocols \cite{Bahl-02, Culler-01,Zhou-06,Adya-01,
	Tzamaloukas-01, Akyildiz-01} that use RTS/CTS
controls\footnote{RTS/CTS controls are required to be implemented by
	802.11-compliant devices. They can be used as an optional mechanism
	to avoid Hidden Terminal Problems in the 802.11 standard and
	protocols based on those similar to \citet{Akyildiz-01} and
	\citet{Adya-01}.} for frequency negotiation and reservation are not
suitable for WSN applications, even though they exhibit good
performance in general wireless ad hoc
networks.

% Head 3
\subsubsection{Exclusive Frequency Assignment}

In exclusive frequency assignment, nodes first exchange their IDs
among two communication hops so that each node knows its two-hop
neighbors' IDs. In the second broadcast, each node beacons all
neighbors' IDs it has collected during the first broadcast period.

% Head 4
\paragraph{Eavesdropping}

Even though the even selection scheme leads to even sharing of
available frequencies among any two-hop neighborhood, it involves a
number of two-hop broadcasts. To reduce the communication cost, we
propose a lightweight eavesdropping scheme.

\subsection{Basic Notations}

As Algorithm~\ref{alg:one} states, for each frequency
number, each node calculates a random number (${\textit{Rnd}}_{\alpha}$) for
itself and a random number (${\textit{Rnd}}_{\beta}$) for each of its two-hop
neighbors with the same pseudorandom number generator.

% Algorithm
\begin{algorithm}[t]
	\SetAlgoNoLine
	\KwIn{Node $\alpha$'s ID ($ID_{\alpha}$), and node $\alpha$'s
		neighbors' IDs within two communication hops.}
	\KwOut{The frequency number ($FreNum_{\alpha}$) node $\alpha$ gets assigned.}
	$index$ = 0; $FreNum_{\alpha}$ = -1\;
	\Repeat{$FreNum_{\alpha} > -1$}{
		$Rnd_{\alpha}$ = Random($ID_{\alpha}$, $index$)\;
		$Found$ = $TRUE$\;
		\For{each node $\beta$ in $\alpha$'s two communication hops
		}{
			$Rnd_{\beta}$ = Random($ID_{\beta}$, $index$)\;
			\If{($Rnd_{\alpha} < Rnd_{\beta}$) \text{or} ($Rnd_{\alpha}$ ==
				$Rnd_{\beta}$ \text{and} $ID_{\alpha} < ID_{\beta}$)\;
			}{
				$Found$ = $FALSE$; break\;
			}
		}
		\eIf{$Found$}{
			$FreNum_{\alpha}$ = $index$\;
		}{
			$index$ ++\;
		}
	}
	\caption{Frequency Number Computation}
	\label{alg:one}
\end{algorithm}

Bus masters are divided into two disjoint sets, $\mathcal{M}_{RT}$
and $\mathcal{M}_{NRT}$.
% description
\begin{description}
	\item[RT Masters]
	$\mathcal{M}_{RT}=\{ \vec{m}_{1},\dots,\vec{m}_{n}\}$ denotes the
	$n$ RT masters issuing real-time constrained requests. To model the
	current request issued by an $\vec{m}_{i}$ in $\mathcal{M}_{RT}$,
	three parameters---the recurrence time $(r_i)$, the service cycle
	$(c_i)$, and the relative deadline $(d_i)$---are used, with their
	relationships.
	\item[NRT Masters]
	$\mathcal{M}_{NRT}=\{ \vec{m}_{n+1},\dots,\vec{m}_{n+m}\}$ is a set
	of $m$ masters issuing nonreal-time constrained requests. In our
	model, each $\vec{m}_{j}$ in $\mathcal{M}_{NRT}$ needs only one
	parameter, the service cycle, to model the current request it
	issues.
\end{description}

Here, a question may arise, since each node has a global ID. Why
don't we just map nodes' IDs within two hops into a group of
frequency numbers and assign those numbers to all nodes within two
hops?

\section{Simulator}
\label{sec:sim}

If the model checker requests successors of a state which are not
created yet, the state space uses the simulator to create the
successors on-the-fly. To create successor states the simulator
conducts the following steps.
% enumerate
\begin{enumerate}
	\item Load state into microcontroller model.
	\item Determine assignments needed for resolving nondeterminism.
	\item For each assignment.
	\begin{enumerate}
		\item either call interrupt handler or simulate effect of next instruction, or
		\item evaluate truth values of atomic propositions.
	\end{enumerate}
	\item Return resulting states.
\end{enumerate}
Figure~\ref{fig:one} shows a typical microcontroller C program that
controls an automotive power window lift. The program is one of the
programs used in the case study described in Section~\ref{sec:sim}.
At first sight, the programs looks like an ANSI~C program. It
contains function calls, assignments, if clauses, and while loops.
% Figure
\begin{figure}
	\includegraphics{mouse}
	\caption{Code before preprocessing.}
	\label{fig:one}
\end{figure}

\subsection{Problem Formulation}

The objective of variable coalescence-based offset assignment is to find
both the coalescence scheme and the MWPC on the coalesced graph. We start
with a few definitions and lemmas for variable coalescence.

% Enunciations
\begin{definition}[Coalesced Node (C-Node)]A C-node is a set of
	live ranges (webs) in the AG or IG that are coalesced. Nodes within the same
	C-node cannot interfere with each other on the IG. Before any coalescing is
	done, each live range is a C-node by itself.
\end{definition}

\begin{definition}[C-AG (Coalesced Access Graph)]The C-AG is the access
	graph after node coalescence, which is composed of all C-nodes and C-edges.
\end{definition}

\begin{lemma}
	The C-MWPC problem is NP-complete.
\end{lemma}
\begin{proof} C-MWPC can be easily reduced to the MWPC problem assuming a
	coalescence graph without any edge or a fully connected interference graph.
	Therefore, each C-node is an uncoalesced live range after value separation
	and C-PC is equivalent to PC. A fully connected interference graph is made
	possible when all live ranges interfere with each other. Thus, the C-MWPC
	problem is NP-complete.
\end{proof}

\begin{lemma}[Lemma Subhead]The solution to the C-MWPC problem is no
	worse than the solution to the MWPC.
\end{lemma}
\begin{proof}
	Simply, any solution to the MWPC is also a solution to the
	C-MWPC. But some solutions to C-MWPC may not apply to the MWPC (if any
	coalescing were made).
\end{proof}

\section{Performance Evaluation}

During all the experiments, the Geographic Forwarding (GF)
\cite{Akyildiz-01} routing protocol is used. GF exploits geographic
information of nodes and conducts local data-forwarding to achieve
end-to-end routing. Our simulation is configured according to the
settings in Table~\ref{tab:one}. Each run lasts for 2 minutes and
repeated 100 times. For each data value we present in the results, we
also give its 90\% confidence interval.

% Table
\begin{table}%
	\caption{Simulation Configuration}
	\label{tab:one}
	\begin{minipage}{\columnwidth}
		\begin{center}
			\begin{tabular}{ll}
				\toprule
				TERRAIN\footnote{This is a table footnote. This is a
					table footnote. This is a table footnote.}   & (200m$\times$200m) Square\\
				Node Number     & 289\\
				Node Placement  & Uniform\\
				Application     & Many-to-Many/Gossip CBR Streams\\
				Payload Size    & 32 bytes\\
				Routing Layer   & GF\\
				MAC Layer       & CSMA/MMSN\\
				Radio Layer     & RADIO-ACCNOISE\\
				Radio Bandwidth & 250Kbps\\
				Radio Range     & 20m--45m\\
				\bottomrule
			\end{tabular}
		\end{center}
		\bigskip\centering
		\footnotesize\emph{Source:} This is a table
		sourcenote. This is a table sourcenote. This is a table
		sourcenote.

		\emph{Note:} This is a table footnote.
	\end{minipage}
\end{table}%

\section{Conclusions}

In this article, we develop the first multifrequency MAC protocol for
WSN applications in which each device adopts a
single radio transceiver. The different MAC design requirements for
WSNs and general wireless ad-hoc networks are
compared, and a complete WSN multifrequency MAC design (MMSN) is
put forth. During the MMSN design, we analyze and evaluate different
choices for frequency assignments and also discuss the nonuniform
back-off algorithms for the slotted media access design.

% Start of "Sample References" section

\section{Typical references in new ACM Reference Format}
A paginated journal article \cite{Abril07}, an enumerated
journal article \cite{Cohen07}, a reference to an entire issue \cite{JCohen96},
a monograph (whole book) \cite{Kosiur01}, a monograph/whole book in a series (see 2a in spec. document)
\cite{Harel79}, a divisible-book such as an anthology or compilation \cite{Editor00}
followed by the same example, however we only output the series if the volume number is given
\cite{Editor00a} (so Editor00a's series should NOT be present since it has no vol. no.),
a chapter in a divisible book \cite{Spector90}, a chapter in a divisible book
in a series \cite{Douglass98}, a multi-volume work as book \cite{Knuth97},
an article in a proceedings (of a conference, symposium, workshop for example)
(paginated proceedings article) \cite{Andler79}, a proceedings article
with all possible elements \cite{Smith10}, an example of an enumerated
proceedings article \cite{VanGundy07},
an informally published work \cite{Harel78}, a doctoral dissertation \cite{Clarkson85},
a master's thesis: \cite{anisi03}, an online document / world wide web
resource \cite{Thornburg01, Ablamowicz07, Poker06}, a video game (Case 1) \cite{Obama08} and (Case 2) \cite{Novak03}
and \cite{Lee05} and (Case 3) a patent \cite{JoeScientist001},
work accepted for publication \cite{rous08}, 'YYYYb'-test for prolific author
\cite{SaeediMEJ10} and \cite{SaeediJETC10}. Other cites might contain
'duplicate' DOI and URLs (some SIAM articles) \cite{Kirschmer:2010:AEI:1958016.1958018}.
Boris / Barbara Beeton: multi-volume works as books
\cite{MR781536} and \cite{MR781537}. If a citation is used as a part of speech, cite it as we do \citet{Abril07}.

% Appendix
\appendix
\section{Switching times}

In this appendix, we measure the channel switching time of Micaz
\cite{CROSSBOW} sensor devices.  In our experiments, one mote
alternatingly switches between Channels~11 and~12. Every time after
the node switches to a channel, it sends out a packet immediately and
then changes to a new channel as soon as the transmission is finished.
We measure the number of packets the test mote can send in 10 seconds,
denoted as $N_{1}$. In contrast, we also measure the same value of the
test mote without switching channels, denoted as $N_{2}$. We calculate
the channel-switching time $s$ as
\begin{displaymath}%
s=\frac{10}{N_{1}}-\frac{10}{N_{2}}.
\end{displaymath}%
By repeating the experiments 100 times, we get the average
channel-switching time of Micaz motes: 24.3\,$\mu$s.

\section{Supplementary materials}

	\subsection{This is an example of Appendix subsection head}

	Channel-switching time is measured as the time length it takes for
	motes to successfully switch from one channel to another. This
	parameter impacts the maximum network throughput, because motes
	cannot receive or send any packet during this period of time, and it
	also affects the efficiency of toggle snooping in MMSN, where motes
	need to sense through channels rapidly.

	By repeating experiments 100 times, we get the average
	channel-switching time of Micaz motes: 24.3 $\mu$s. We then conduct
	the same experiments with different Micaz motes, as well as
	experiments with the transmitter switching from Channel 11 to other
	channels. In both scenarios, the channel-switching time does not have
	obvious changes. (In our experiments, all values are in the range of
	23.6 $\mu$s to 24.9 $\mu$s.)

	\subsection{Appendix subsection head}

	The primary consumer of energy in WSNs is idle listening. The key to
	reduce idle listening is executing low duty-cycle on nodes. Two
	primary approaches are considered in controlling duty-cycles in the
	MAC layer.

\begin{acks}

	The authors would like to thank Dr. Maura Turolla of Telecom
	Italia for providing specifications about the application scenario.

	The work is supported by the \grantsponsor{GS501100001809}{National
		Natural Science Foundation of
		China}{http://dx.doi.org/10.13039/501100001809} under Grant
	No.:~\grantnum{GS501100001809}{61273304\_a}
	and~\grantnum[http://www.nnsf.cn/youngscientsts]{GS501100001809}{Young
		Scientsts' Support Program}.

\end{acks}

% Bibliography
\bibliographystyle{ACM-Reference-Format}
\bibliography{sample-bibliography}

\end{document}

### 2018-02-11 paper.tex
Added algorithms

@@
 \usepackage[font=footnotesize]{caption}
+\usepackage{lineno}
 
@@
 
+\linenumbers
+
 \section{Introduction} \label{sec:introduction}
 
-\paragraph{Main idea} We propose a new dynamic kidney matching algorithm that takes as input a pool of patient-donor pairs and computes, for each cycle that is currently available for matching, a probability that it should be chosen to be cleared today, as well as another probability indicating whether we should not match any more cycles.
+\paragraph{Main idea} We study new dynamic kidney matching algorithm that takes as input a pool of patient-donor pairs and computes, for each cycle that is currently available for matching, a probability that it should be chosen to be cleared today, as well as another probability indicating whether we should not match any more cycles.
 In order to get these probabilities, our algorithm first produces a prior distribution estimated by observing past choices made by an infeasible optimal algorithm, and then updates these priors with simulations.
@@
 However, if they happen to form an (A,O) pair (with probability $0.36 \cdot 0.48 \approx 0.172$), then they should only enter the pool if the donor and patient are tissue-type incompatible, otherwise the O donor would immediately donate their kidney to the patient and they would not need to enter the exchange pool at all.
-So we assume the probability of a positive crossmatch is $p_c = 0.11$ as in \cite{zenios2001primum}, then the probability that two people form an (A,O) and enter the pool becomes in fact $0.36 \cdot 0.48  \cdot 0.11 \approx 0.019$.
+So we assume the probability of a positive crossmatch is $p_c = 0.11$ as in \cite{zenios2001primum}, then the probability that two people form an (A,O) and enter the pool becomes in fact proportional to $0.36 \cdot 0.48  \cdot 0.11 \approx 0.019$.
 By computing the probabilities of each pair in this manner and then normalizing so that they add up to one, we arrive at the numbers displayed on Table \ref{tab:abo_blood_type}.

### 2018-02-11 ec-18/kidney-ec-submission.tex
Pseudo-reward intuition figure

@@
 \IncMargin{-\parindent}
+\newcommand\mycommfont[1]{\footnotesize\upshape\textcolor{blue}{#1}}
+\SetCommentSty{mycommfont}
 
@@
 \caption{Ratio of \texttt{Myopic} has better chances of achieving performances similar to \texttt{OPT} when the death rate is high (moving rightwards on the graphs), or when entry rate is high (moving downwards on the graphs).}
-\label{fig:greedy_opt_compariso}
+\label{fig:greedy_opt_comparison}
 \end{figure}
+
+\begin{algorithm}[htbp]
+	\SetAlgoLined
+  \DontPrintSemicolon
+
+  \KwData{Environment simulator object \textsc{Env}; \\
+          Integer programming solver object \textsc{Solver}; \\
+          Statistical method \textsc{Classifier} \\
+          Threshold $thres$; }
+
+  \SetKwFunction{FChoose}{\textsc{direct\_prediction}}
+  \SetKwProg{Fn}{Function}{:}{}
+  \Fn{\FChoose{\textsc{Env}, \ \textsc{Solver}, \ \textsc{Classifier}, \ $thres$}}{
+    \tcp{Retrieve data about nodes and (potentially) graph from current environment} \\
+    $X, E_1, E_2$ \gets \textsc{Env.get\_data}()  \\
+    \tcp{Pass data through classifier, predict matching probability for each pair} \\
+    $prob$ \gets \textsc{Classifier}($X$, \ $E_1$, $E_2$)  \\
+    \tcp{Get index of pairs whose probability is higher than threshold} \\
+    $index$ $\gets$ \textsc{which}($prob > thres$) \\
+    \tcp{Find maximal matching restricted to this subset} \\
+    chosen\_cycles $\gets$ \textsc{Solver.solve(Env, subset=$index$)}  \\  
+    \tcp{Return chosen cycles to be cleared} \\
+    \KwRet chosen\_cycles \;
+  }
+	\caption{Function \textsc{direct\_prediction}}
+	\label{alg:direct_prediction}
+\end{algorithm}
 
@@
 \centering
-\hspace*{-1cm}
-\includegraphics[width=1.1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/direct_prediction.pdf}.
-\caption{Direct prediction methods.}
+\includegraphics[width=\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/direct_prediction.pdf}.
+\caption{One period of simulation using direct prediction methods to choose cycles.}
 \label{fig:direct_prediction}
@@
   \centering
-  \hspace*{-1cm}
-  \includegraphics[width=1.1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/mab.pdf}
-  \caption{\textbf{Function} \textsc{choose\_cycle}}
+  %\hspace{-1cm}
+  \includegraphics[width=1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/mab.pdf}
+  \caption{One period of simulation using multi-armed bandit methods to choose cycles.}
   \label{fig:mab}
@@
   \centering
-  \includegraphics[width=1.1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/mab_step.pdf}
-  \caption{\textbf{Function} \textsc{get\_pseudo\_reward} }
+  %\hspace*{-0.5cm}
+  \includegraphics[width=1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/mab_step.pdf}
+  \caption{\textbf{Function} \textsc{get\_pseudo\_reward} \\ Multi-armed bandits evaluate if a cycle should be cleared by checking if there is a high chance that the cycle will be used in the future.
+Such cycles get a \emph{lower} reward, and are left for later.}
   \label{fig:get_pseudo_reward}
+  \end{figure}
+
+\paragraph{What are pseudo-rewards?} Figure \ref{fig:pseudo_reward_intuition}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=0.95\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/pseudo_reward_intuition.pdf}
+\caption{Pseudo-reward intuition}
+\label{fig:pseudo_reward_intuition}
 \end{figure}
@@
 
-\newcommand\mycommfont[1]{\footnotesize\upshape\textcolor{blue}{#1}}
-\SetCommentSty{mycommfont}
-
 % Algorithm
-\begin{algorithm}[t]
+\begin{algorithm}[htbp]
 	\SetAlgoLined
@@
 
-\begin{algorithm}[t]
+\begin{algorithm}[htbp]
 

### 2018-02-12 ec-18/kidney-ec-submission.tex
Missing literature review and results

@@
 \usepackage[ruled]{algorithm2e} % For algorithms
+
+\usepackage{ textcomp }
 
@@
 % Title portion. Note the short title for running heads 
-\title[New strategies for improving kidney exchange]{New strategies for improving kidney exchange}  
+\title[New strategies for dynamic kidney exchange]{New strategies for dynamic kidney exchange}  
 \author{Submission 42}
@@
 
-\paragraph{\textbf{Main idea}} We propose and compare two new approaches to improve the number of matching in kidney exchange, namely the \emph{direction estimation} method and the \emph{multi-armed-bandit} method.
-The former recasts the dynamic kidney exchange problem as a binary classification task: for each pair in the pool, we should predict a 0-1 label representing whether they should be matched today, or left for the future.
-The latter chooses exchanges that are less likely to produce ``regret" in the future (under a specific definition of regret that we elaborate on later).
+\paragraph{\textbf{Main idea}} We propose and empirically evaluate two novel approaches for increasing the cardinality of matched pairs in a discrete time dynamic model of kidney exchange.
+Our two methods are named the \emph{direction estimation} and \emph{multi-armed-bandit} methods.
+The former recasts the dynamic kidney exchange problem as a binary classification task: for each pair in the pool, we predict a binary label representing whether they should be matched today (one), or left for the future (zero), given their observable characteristics.
+The latter uses simulations to select exchanges that are less likely to decrease the size of an optimal future matching.
+We restrict our analyses to pairwise exchanges (i.e., only two-cycles are allowed), but the ideas introduced here can be easily extended to more complex exchanges, computational constraints permitting.
 
-Each approach faces different challenges.
-In the \emph{direct estimation} approach, one challenge is that an important part of our data -- the compatibility graph -- lives in a non-Euclidean space, and it is not obvious how to use it as an input to a classification algorithm.
+Each of our two approaches faces their own challenges.
+In the \emph{direct estimation} method, one challenge is that an important part of our data -- the compatibility graph -- lives in a non-Euclidean space, and it is not obvious how to use it as an input to a classification algorithm.
 We attempt to apply a variety of new methods from the graphical signal processing literature to tackle this problem.
-An additional challenge is that traditional classification algorithms are not equipped to produce dependent labels, but in our setting whether or not we decide that a certain pair is to be matched today ought to be related to our decisions about everyone else in the pool.
-To overcome this we use recurrent neural networks and draw from the [[[[]]]]
+[[[An additional challenge is that traditional classification algorithms are not equipped to produce dependent labels, but in our setting whether or not we decide that a certain pair is to be matched today ought to be related to our decisions about everyone else in the pool.]]]]
 
-In the \emph{multi-armed bandit} method, the main issue is that the enormous branching factor of our problem precludes an accurate estimation of regret for each possible exchange.
-Here, we develop a novel algorithm that iteratively chooses to match one additional pair or decides
+In the \emph{multi-armed bandit} method, we would like to predict how likely it is that a certain pair should be reserved for a future exchange, and match those less frequently.
+However, computationally costly simulations and the enormous branching factor of our problem precludes us from accurately estimating of this probability.
+In order to overcome this challenge, we employ multi-armed bandit (MAB) algorithms that optimally choose which exchange to collect statistics about under a relatively small computational budget.
 
-\paragraph{\textbf{Main results}} We conduct simulations following three models from the kidney exchange literature, which we call \emph{environments}.
-The three environments differ by their rules regarding blood- and tissue-type compatibility.
+\paragraph{\textbf{Main results}} Our methods are evaluated via simulations.
+The simulation setups, which we call \emph{environments}, are inspired by models used the kidney exchange literature, and differ by their rules regarding blood- and tissue-type compatibility.
 
-We apply our methods to each of these environments and compare them to a myopic algorithm (representing how kidney exchanges are currently conducted), and to infeasible optimal algorithm.
-Comparison is made in terms of average number of pairs matched per period, and the average is taken over a long period of time (i.e., much longer a node's mean sojourn).
+We apply our methods to each simulation setup and compare them to a \textsc{myopic} algorithm that clears the maximal matching at each period, and to an infeasible \textsc{OPT} algorithm that is able to observe future states of the world and act optimally.
+Our metric for comparison is the per-period average number of matched pairs over a long period of time.
 
 In all environments, we observe that the \emph{direct estimation} unfortunately does equally well or worse than the myopic algorithm.
-However, the \emph{multi-armed bandit} method is able to improve upon it by 0.5-5\% in terms of the number of average pairs matched per period.
-Moreover, we observe that there is still a moderately large performance gap between our algorithms and the infeasible optimal.
+However, the \emph{multi-armed bandit} method is able to improve substantially upon the \textsc{Myopic} benchmark, [[[suggesting that we could save more lives etc.
+]]]]
 
@@
 
-\paragraph{Dynamic kidney exchange}
+\paragraph{Matching theory} Kidney exchange as an economic problem was first studied in a series of papers by \citet{roth2004kidney, roth2005pairwise, roth2007efficient}, but the fist paper on dynamic kidney exchange was written a few years later by \citet{unver2010dynamic}, who derived an optimal matching mechanism in a simple continuous time model.
+Two of the simulation setups in our work follows a modified, discrete-time version of the models in these seminal papers.
 
-\paragraph{Computer science}
+In the computer science literature, \citet{awasthi2009online} and later \citet{dickerson2015futurematch} wrote computational alternatives to simple myopic kidney exchange.
+Our methods are similar to the \emph{weighted myopia} scheme seen in these papers.
+Other papers that influence ours are \citet{akbarpour2017thickness} and \citet{ashlagi2013kidney}.
+
+\paragraph{Computer science and machine learning}
 
@@
 
-\paragraph{Notation}
+\paragraph{Basic definitions and notation}
+
+An \emph{agent} is an algorithm that decides who should be matched.
 
@@
 
-\paragraph{Poisson entry, Geometric death} At time point, the number of new incoming pairs is drawn from the $Poisson(r)$ distribution, where $r \in \mathbb{N}$ denotes the \emph{entry rate}, and equals the expected number of entrants per period.
+\paragraph{Poisson entry, Geometric death} At each period, the number of new incoming pairs is drawn from the $Poisson(\lambda)$ distribution, where $\lambda \in \mathbb{N}$ denotes the \emph{entry rate}, and equals the expected number of entrants per period.
 Upon entrance, each pair independently draws the length of their sojourn from the $Geometric(d)$ distribution.
@@
 
-\paragraph{Minimum set of observable characteristics} In every environment, at least two characteristics are relevant and observable: the major blood type groups (A, B, O or AB) of the patient and donors, and how long the pair has been in the pool.
-On the other hand, death times are never observed.
-
 \paragraph{Binary preferences} In this work we abstract from these concerns and consider every exchange to be equally desirable, so long as it is available.
-This is the same assumption as used in \cite{roth2005pairwise},
+This is the same assumption as used in \cite{roth2005pairwise}.
 
@@
 
-In the \emph{ABO environment}, a pair's observable characteristics are their current waiting time and dummies representing ABO blood types of its patient and donor.
-Compatibility between two distinct pairs is based only on blood-type compatibility.
-Blood types are drawn independently for patient and donor from the US population (roughly O:48\%, A:36\%, B:11\%, AB:4\%).
-In addition, we make an assumption previously used in \citet{unver2010dynamic} and allow for incompatibility between a donor and their own patient.
-The reason for this additional assumption is that, if there were truly no tissue type compatibilities, we would never observe pairs of type (AB,$\cdot$), ($\cdot$, O), or (A,A), (O,O), (B,B), and (AB,AB), since their donors would be automatically compatible with their patients and they would never participate in an exchange.
+In the \emph{ABO environment}, compatibility between two distinct pairs is based only on blood-type compatibility.
+Blood types are drawn independently for patient and donor from the US population (roughly $O$:$49\%$, $A$:$36\%$, $B$:$11\%$, $AB$:$4\%$).
+In addition, we make an assumption previously used in \citet{unver2010dynamic} to allow for incompatibility between a donor and their own patient.
+The reason for this additional assumption is that, if there were truly no tissue type compatibilities, we would never observe pairs of type $(AB,\cdot)$, $(\cdot, O)$, or $(A,A)$, $(O,O)$, $(B,B)$, and $(AB,AB)$, since their donors would be automatically compatible with their patients and they would never participate in an exchange.
 Following \citet{zenios2001primum}, we assume that for such patients the probability that a donor and their patient are incompatible is $p_c = 0.11$.
@@
 
-The \emph{RSU} environment drawn from the setup in \cite{roth2007efficient}.
-Observable covariates are dummies representing patient and donor ABO blood types, current waiting time, and a \emph{calculated panel reactive antibody (cPRA)} level that governs the probability of a crossmatch with a random donor.\footnote{The original \cite{roth2007efficient} paper called this simply \emph{PRA}.
+The \emph{RSU} environment is named after the simulation model in \citet*{roth2007efficient}.
+Each pair is characterized by patient and donor ABO blood types, current waiting time, and a \emph{calculated panel reactive antibody (cPRA)} level that governs the probability of a crossmatch with a random donor.\footnote{The original \citet{roth2007efficient} paper called this simply \emph{PRA}.
 In real life there is an important distinction between the two measures, but for the purposes of a simulation model this distinction is immaterial.
-We keep the name \emph{cPRA} for consistency with OPTN environment.} The lower the cPRA, the higher the number of potentially compatible pairs.
+We keep the name \emph{cPRA} for consistency with OPTN environment later.} The lower the cPRA, the higher the number of potentially compatible pairs.
 
@@
 From this original dataset, we excluded patients that were registered for more than one organ (including kidney+pancreas), patients that were not waiting for their first kidney transplant, and patients and donors with incomplete HLA profile information.
-The resulting historical dataset still contained 117813 patients and 9337 living donors.
+The resulting dataset contained 117813 patients and 9337 living donors.
 
@@
 Figure \ref{fig:cpra} shows the distribution across the entire population.\footnote{We should remark that we found a very different patient cPRA distribution than the one in the OPTN dataset.
-This may have been because: in real life cPRA is computed using deceased donor data, whereas we use living donor data; we have used different HLA equivalence tables; OPTN uses a more sophisticated model based on population genetics.\cite{optn2013cpra}}
+This may have been because: in real life cPRA is computed using deceased donor data, while we used living donor data; we may have used different HLA equivalence tables; OPTN uses a more sophisticated model based on population genetics.\cite{optn2013cpra}}
 
@@
 \includegraphics[width=\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/computed_cPRA.pdf}
-\caption{Patient and donor cPRA computed from historical data.}
+\caption{\textbf{Patient and donor cPRA} \\
+    Computed from historical data.
+Our patient cPRA is the frequency of living donors that exhibit antigens that are unacceptable to the patient.
+We also define a donor cPRA by calculating the frequency of patients that have antibodies against the donors antigens.}
 \label{fig:cpra}
@@
 
-\textbf{Remark} \textit{How much is there to be improved upon?} [[]]
+\paragraph{\textbf{Remark}} \textit{How much is there to be improved upon?} In Figure \ref{fig:greedy_opt_comparison}, we compare \textsc{Myopic} and \textsc{OPT} for a grid of different entry and death rates.
+These results show that the performance gap between the two can be fairly large, in particular in sparser environments like \emph{RSU} and \emph{OPTN}.
+We also note that the gap is narrower when: the death rate is high, because if most pairs will die soon, dynamic considerations play a smaller role; and when the entry rate is very large, because in a thicker market pairs are able to encounter a suitable match more easily.
 
@@
 \includegraphics[width=1.4\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/greedy_opt_comparison_mcl2_2.pdf}
-\caption{Ratio of \texttt{Myopic} has better chances of achieving performances similar to \texttt{OPT} when the death rate is high (moving rightwards on the graphs), or when entry rate is high (moving downwards on the graphs).}
+\caption{\textbf{Comparing \textsc{Myopic} and \textsc{OPT}}  \\
+Ratio of average per-period matched pairs for different entry and death rates, over 3000 periods (darker hues are better).
+\texttt{Myopic} has better chances of achieving performances similar to \texttt{OPT} when the death rate is high (moving rightwards on the graphs), or when entry rate is high (moving downwards on the graphs).}
 \label{fig:greedy_opt_comparison}
+\end{figure}
+
+% ---------------------------------------------------------------------------
+% Algorithms
+% ----------------------------------------------------------------------
+
+\section{Algorithms}
+
+We now present two novel methods to determine which patients should be matched.
+
+\subsection{Direct prediction} \label{subsec:direct_prediction}
+
+Dynamic and static algorithms can work in tandem: the former can determine which pairs should be matched today, while the latter decides how they should be matched among themselves.
+The \emph{direct prediction}, exploits this insight, essentially reducing the problem to a classification task: at each period, we aim to produce a binary label for each node indicating whether is should be matched in this period (1) or left for later (0).
+Selected nodes are then passed to a static solver that finds the maximal matching among them.
+Once these nodes are cleared, time evolves to the next period.
+The procedure is formalized in Algorithm \ref{alg:direct_prediction}, and also illustrated in Figure \ref{fig:direct_prediction}.
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/direct_prediction.pdf}.
+\caption{\textbf{Direct prediction method} \\ One period of simulation using direct prediction methods to choose cycles. See Algorithm \ref{alg:direct_prediction} for details.}
+\label{fig:direct_prediction}
 \end{figure}
@@
   \Fn{\FChoose{\textsc{Env}, \ \textsc{Solver}, \ \textsc{Classifier}, \ $thres$}}{
-    \tcp{Retrieve data about nodes and (potentially) graph from current environment} \\
+    \tcp{Retrieve node (and potentially also graph) data from current environment} \\
     $X, E_1, E_2$ \gets \textsc{Env.get\_data}()  \\
-    \tcp{Pass data through classifier, predict matching probability for each pair} \\
+    \tcp{Classifier predicts matching probability for each pair using data} \\
     $prob$ \gets \textsc{Classifier}($X$, \ $E_1$, $E_2$)  \\
@@
 
-\section{Algorithms}
+\subsection{Direct prediction methods}
 
-We present two new methods to determine which patients should be matched.
+In order to produce data to feed into our ``classifier", we repeatedly created 1000-period simulation runs and solved using them OPT using with each each node's observable characteristics as regressors, and a binary variable indicating whether they were matched or not as the regressand.
+The resulting artificial dataset of approximately 2 million observations was then fed to a series of predictive algorithms, including penalized logistic regression \citep{wu2009genome}, support vector machines with radial basis function kernels \citep{cortes1995support}, random forest classifiers \citep{breiman2001random}, gradient tree boosting classifiers \citep{friedman2001greedy} and deep feedforward networks \citep{goodfellow2016deep}.
 
-\subsection{Direct prediction} \label{subsec:direct_prediction}
+Moreover, in order to represent information about the graph and how the nodes fit within it, we augment the original dataset using one or both of the following.
+First, with vertex properties for each node such as the number of in and out-edges, average neighbor degree, some centrality measures (betweenness, in-degree, out-degree, harmonic, closeness) and more exotic indices such as the core number and pagerank of each node\cite{newman2010networks}.
+Second, we augment the data using the \emph{node2vec} graph embedding algorithm of \cite{grover2016node2vec}, a continuous feature representation algorithm inspired by the skip-gram model.[[[More on this?]]].
+These algorithms will produce auxiliary matrices $E_t^{1}$ and $E_t^2$ that are concatenated to the original data regressor matrix $X_t$, and the entire augmented data set can again be fed to the predictive algorithms above.
+Their performance is compared on the bottom half of Table \ref{tab:direct_prediction}.
 
-Dynamic and static algorithms can work in tandem: the former can determine which pairs should be matched at each period, while the latter decides how they should be matched among themselves.
-The \emph{direct prediction} exploits this insight, essentially reducing the problem to a classification task: at each period, we aim to produce a binary label for each node indicating whether is should be matched in this period (1) or left for later (0).
-Selected nodes are then passed to a static solver that finds the maximal matching among them.
-Once these nodes are cleared, time evolves to the next period.
-This is illustrated in Figure \ref{fig:direct_prediction}.
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/direct_prediction.pdf}.
-\caption{One period of simulation using direct prediction methods to choose cycles.}
-\label{fig:direct_prediction}
-\end{figure}
-
-\section{Results}
-
-\subsection{Direct estimation policy}
-
-In order to produce data to feed into our ``classifier", we simulated and solved a 1000-period run environment was simulated and solved using OPT using with each each node's observable characteristics as regressors, and a binary variable indicating whether they were matched or not as the regressand.
-The resulting artificial dataset of approximately 2 million observations was then fed to a series of predictive algorithms, including penalized logistic regression\citep{wu2009genome}, support vector machines with radial basis function kernels\citep{cortes1995support}, random forest classifiers\citep{breiman2001random}, gradient tree boosting classifiers\citep{friedman2001greedy} and deep feedforward networks\citep{goodfellow2016deep}.
-The performance of these algorithms as classification methods is compared in Table \ref{tab:direct_prediction} in the Supplementary Material section.
-
-\subsection{MAB-based policy}
+\subsection{Multi-armed bandit methods}
 
@@
 
- \begin{figure}
-  \centering
-  %\hspace{-1cm}
-  \includegraphics[width=1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/mab.pdf}
-  \caption{One period of simulation using multi-armed bandit methods to choose cycles.}
-  \label{fig:mab}
-  \end{figure}
-
-  \begin{figure}
-  \centering
-  %\hspace*{-0.5cm}
-  \includegraphics[width=1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/mab_step.pdf}
-  \caption{\textbf{Function} \textsc{get\_pseudo\_reward} \\ Multi-armed bandits evaluate if a cycle should be cleared by checking if there is a high chance that the cycle will be used in the future.
-Such cycles get a \emph{lower} reward, and are left for later.}
-  \label{fig:get_pseudo_reward}
-  \end{figure}
-
-\paragraph{What are pseudo-rewards?} Figure \ref{fig:pseudo_reward_intuition}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=0.95\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/pseudo_reward_intuition.pdf}
-\caption{Pseudo-reward intuition}
-\label{fig:pseudo_reward_intuition}
-\end{figure}
-
-\section{Conclusion}
-
-\paragraph{Summary}
-
-\paragraph{Extensions} In reality, some exchanges are more or less desirable than others, either for ethical concerns or because of predicted health benefit.
-For example, it is common for pediatric patients and previous organ donors receive higher priority, as do exchanges involving patients with no HLA mismatch.
-
  % Algorithm
@@
            Environment simulator object \textsc{Env}; \\
-          Oracle solver \textsc{OPT}; }
+           Oracle solver object \textsc{OPT}; }
    \DontPrintSemicolon
@@
 
+ \begin{figure}[htbp]
+  \centering
+  %\hspace{-1cm}
+  \includegraphics[width=1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/mab.pdf}
+  \caption{\textbf{Multi-armed bandit methods} \\ 
+  One period of simulation using multi-armed bandit methods to choose cycles.}
+  \label{fig:mab}
+  \end{figure}
+
+  \begin{figure}[htbp]
+  \centering
+  %\hspace*{-0.5cm}
+  \includegraphics[width=1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/mab_step.pdf}
+  \caption{\textbf{Function} \textsc{get\_pseudo\_reward} \\ Multi-armed bandits evaluate if a cycle should be cleared by checking if there is a high chance that the cycle will be used in the future.
+Such cycles get a \emph{lower} reward, and are left for later.}
+  \label{fig:get_pseudo_reward}
+  \end{figure}
+
+\paragraph{What are pseudo-rewards?} As we match and clear out a cycle $c$ today, we forgo the opportunity of using any future cycles involving the nodes in $c$.
+However, because there may be multiple optimal matchings, sometimes the cycles that become unavailable due to the removal of the nodes in $c$ do not matter, and we can still find a matching of the same cardinality without them.
+In other words, we pay no price for removing these future cycles.
+
+The pseudo-reward associated with cycle $c$ is a random variable whose expectation is the probability that removing a cycle today will \emph{not} negatively impact the optimal matching size between $t$ and $t+h$.
+The higher this number, the more confident we are that clearing $c$ out today will not give us trouble in the future.
+A concrete example of this idea in shown in Figure \ref{fig:pseudo_reward_intuition}.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=0.95\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/pseudo_reward_intuition.pdf}
+\caption{\textbf{Intuition for pseudo-rewards} \\
+  In the example above, at time $t$ we have three available cycles to choose from: $(1,2)$, $(1,3)$ and $(3,4)$, and we are contemplating choosing cycle $(1,2)$ at $t$.
+If we  constrain ourselves to do so, several other future cycles become unavailable (in red).
+Nevertheless, in three out of our four simulations we were able to clear the same number of cycles in the constrained scenario, so the running pseudo-reward average is $\frac{3}{4}$.
+}
+\label{fig:pseudo_reward_intuition}
+\end{figure}
+
+Pseudo-rewards are a natural way for us to control which patients should be matched today.
+Pairs that are easy to match will likely belong to many cycles, so by removing them we will be incurring a large cost in terms of future cycles that will become unavailable.
+But that means that the pseudo-reward associated with cycles that involve them will be lower, making them less attractive.
+On the other hand, patients that are harder to match will not participate many future exchange, so the price we pay for matching them today is low, and their average pseudo-reward is high.
+
+% ---------------------------------------------------------------------------
+% Results
+% ---------------------------------------------------------------------------
+
+\section{Results}
+
+\subsection{Direction prediction methods}
+
+Table \ref{tab:traditional_ml_classifier} shows the performance results for direct prediction methods as classification methods, i.e., how accurately they are able to predict whether a node should be matched or not.
+Results imply the following.
+
+First, while overall accuracy can be relatively high at 70-80\%, precision (defined as the ratio between true positives and both true and false positives) is low even for simpler environments like $ABO$, indicating that the models are over-predicting matchings.
+Second, augmenting the data with information about the graph has no discernible effect under any of the performance criteria.
+This suggests that there might be gains from using other algorithms that make better use of information about the compatibility graph.
+
+Unfortunately, the algorithms' substandard performance as classifiers translates into poor performance as policy as well.
+
+\subsection{Multi-armed bandit methods}
+
+\section{Conclusion}
+
+\paragraph{Summary}
+
+\paragraph{Extensions}
+Another source of error is the fact that we are predicting a label for each node independently from the labels we have predicted for other nodes.
+
+ In reality, some exchanges are more or less desirable than others, either for ethical concerns or because of predicted health benefit.
+For example, it is common for pediatric patients and previous organ donors receive higher priority, as do exchanges involving patients with no HLA mismatch.
+
 \section{Conclusions}
@@
 \section{Supplementary materials}
+
+\begin{table}
+  \centering
+  \singlespacing
+  \hspace*{-1cm}
+  \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/tables/traditional_ml_classifier.tex}
+   \caption{\textbf{Traditional methods} \\ Performance of traditional econometric and classification methods when predicting whether a node will be matched (1) or not (0).
+\emph{Additional} refers to extra information about graph structure passed to the classifier: \emph{Graph Stats} are graph-theoretical measures of node characteristics such as centrality; \emph{Embedding} is a 10-dimensional real embedding of the node provided by node2vec \citep{grover2016node2vec}.
+Recall is true positive rate $\frac{tp}{tp+fn}$; Specificity is true negative rate $\frac{tp}{tn+fp}$; Precision is $\frac{tp}{tp+fp}$; Accuracy is $\frac{tp+tn}{tp+tn+fp+fn}$.
+See section \ref{sec:[[[]]]} for methodology and section \ref{sec:[[[]]]} for discussion of results.
+}
+  \label{tab:tradional_ml_classifier}
+\end{table}
 

### 2018-04-04 chapter1/chapter1.tex
etc

\chapter{Heterogenous Production Functions, Panel Data, and Productivity Dispersion}

\doublespacing

\section{Introduction}

[Primary motivations of panel data: controlling for unobserved heterogeneity]

[Correlated random coefficients]

[]
Random coefficients
 Consumer panels with heterogeneous consumers
 Production function parameters (A,B) vary across firms in
same industry

Random coefficients (A,B) can be correlated with inputs (X1,X2)
  Estimator allows arbitrary correlations of production function parameters (A1,B1) with inputs (X1,X2)
  Timing assumptions with random coefficients (A,B)
  Input decisions made in period t − 1 with knowledge of t − 1
production function
  Estimation, consistency: censoring for unconditional means
  India: log TFP dispersion increases in random coefficients vs OLS

\section{Literature review}

In its most general form, the basic linear panel model used in econometrics can be described as

\begin{align}
  Y_{it} = A_{it} + \sum_{k=1}^{K} X_{it}^{(k)}B_{it}^{(k)}
\end{align}

\noindent where $Y_{it}, X_{it}, \{A_{it}, B_{it}\}_{k=1}^{K}$ are random variables associated with individual $i$ at period $t$. The parameters $B_{it}^{(k)}$ represent the causal effect of an increase in their associated regressor, while $A_{it}$ is a disturbance term whose interpretation depends on the specific application. When they have a nondegerate probability distribution, they receive various names in the literature, such as ``\emph{unobserved component}, \emph{latent variable}, or \emph{unobserved heterogeneity}"\citep[~p.251]{wooldridge2010econometric}.

 but here we call them \emph{random coefficients}.

Without further restrictions, nothing can be said about the objects $A_{it}$ and $B_{it}$, or even its moments. In the linear panel data literature, common identifying assumptions include the following three.

\begin{enumerate}
  \item $A_{it} = a + \epsilon_{it}$ and $B_{it} = b$, where $a, b$ are constants and $E[\epsilon_{it}|X_{it}] = 0 \forall t$.
  \item $A_{it} = A_{i} + \epsilon_{it}$ and $B_{it} = b$, where $b$ is a constant and $E[\epsilon_{it}|X_{is}] = 0 \forall t, s$.
  \item $A_{it} = A_{i} + \epsilon_{it}$ and $B_{it} = b$, where $b$ is a constant and $E[\epsilon_{it}|X_{is}] \neq 0 \forall t, s$.
\end{enumerate}

The first one is the standard linear model that can be consistently estimated by simply running a standard linear regression of $Y_{it}$ on $X_{it}$. The second and third are called \emph{random} and \emph{fixed} panel data models, and have received a great deal of attention in the literature (see [???] for a review). Figure \ref{fig:panel_data_models} illustrates the difference between them.

\begin{figure}
  \includegraphics[width=1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter1/figures/panel_data_models.pdf}
  \caption[Linear panel data models]{\textbf{Linear panel data models} \\ Different assumptions leading to different panel data models. Each different marker type and color indicates a different individual. \emph{Pooled}: all observations share same intercept and slope; \emph{Random effects}: each observation possesses their own time-invariant slope, but all share the same slope; \emph{Fixed effects}: same as random effects, but intercept and regressors can be correlated (figure shows positive correlation); \emph{FHHPS}: our paper generalizes the model by 1) allowing for intercept and slope to be correlated with regressors, and 2) allowing both intercept and slope to drift according to a specific Markovian process.}
  \label{fig:panel_data_models}
\end{figure}

In this paper we focus on the more general case where we allow for more than one random coefficient, and also allow them to be correlated with the regressor. The literature dealing with this more general case is substantially smaller. \citet{hoderlein2010analyzing} analyzed a for any number of random coefficients, but required the stronger assumption that the coefficients are independent from the regressors.

> [Chamberlain (1982), Arellano and Bonhomme (2011), Graham
and Powell (2012), Evdokimov (2011)]

> Recent papers on heterogeneous production functions Kasahara, Shrimpf and Suzuki (2016), Balat, Brambilla and Sazaki (2016)
  Extend proxy variable approaches but require variables to be chosen in period t − 1 with only static considerations, as in FOCs. Ackerberg and Hahn (2016)
  Scalar unobservable enters into nonparametric production function

  Cobb-Douglas
  Yi,t =Ai,t +BKKi,t +BLLi,t
    Productivity dispersion: standard deviation of Ai,t across plants or firms in same industry
    Syverson (2011) surveys empirical findings on dispersion
    Typical finding: some plants produce more than twice as much
  output for same inputs
    Assuming  BK,BL  is the same across plants in a 4- or 2-digit industry classification may be too restrictive
    If BK,BL  varybyplanti andtimet andabove i,t i,t
  specification is estimated, much of this heterogeneity enters into Ai,t

  Estimate BK as expenditure share on capital input i,t
  Based on static profit maximizing FOC for Cobb-Douglas
  Treats input as having no adjustment costs, opposite of our
approach
  Based on strong conduct assumption: static profit maximization
  We do not impose profit maximization, important for studying unproductive firms
  We allow unobservables outside production function to affect input choice
  Firm-specific input prices
  Adjustment costs
  Product demand

\section{Model and assumptions}

In order to reduce clutter we will drop the $i$ subscript. For $1 \leq t \leq T$, define
\begin{align*}
Y_{t} &= A_{t} + B_{t}X_{t}
\end{align*}
%$Z_{t}$ (a $k \times 1$ vector)
%and $\beta_t$  is a $k \times 1$ vector of fixed coefficients, 
where $X_{t}$ is a scalar, and $A_{t}, B_{t}$ are scalar random coefficients that evolve according to an AR(1) Markov process.
\begin{align}
  A_{t} &= A_{t-1} + U_{t} \\
  B_{t} &= B_{t-1} + V_{t}  
\end{align}
% 
% Let $r$ be a positive integer and for $t \geq 2$ define $W_t = (X_{t-1}, X_{t})$. We will say that $(U_t, V_t)$ are \emph{r\textsuperscript{th}-moment independent} of $W_t$ if 
%   \begin{align}
%     \text{For }1 \leq j \leq r,
%     \quad     
%     E[U_{t}^{r-j} V_{t}^{j} | W_t] = E[U_{t}^{r-j} V_{t}^{j}]
%   \end{align}
% \end{definition}

We will make the following sufficient assumptions for identifiability. Some of these are known to be stronger than necessary, and we will mention how to relax them later on. For ease of notation, let us denote $W_t = [X_t, X_{t-1}]^{T}$

\begin{assumption}{} \label{a:timeindependence}
  Conditional on recent covariates $W_t$, contemporaneous shocks are independent from past shocks.
  \begin{align}
    \text{For }2 \leq t \leq T ,
    \quad     
    (U_t, V_t) \perp \{ U_s, V_s \}_{s=0}^{t-1} \  | \ W_t
  \end{align}
\end{assumption}

\begin{assumption}{} \label{a:sequentialindependence}
  Shocks are sequentially independent from all covariates
  % \begin{align}
  %   \text{For }1 \leq j \leq r,
  %   \quad     
  %   E[U_{t}^{r-j} V_{t}^{j} | W_t] = E[U_{t}^{r-j} V_{t}^{j}]
  % \end{align}
  \begin{align}
    \forall t, \qquad U_t, V_t \perp (X_{1}, \cdots, X_{T})
  \end{align}
\end{assumption}

\begin{assumption}{}  \label{a:support}
  The covariate associated with the random coefficient is continuously distributed and is supported on the entire real line for all periods. Furthermore, their joint distribution is non-degenerate.
  \begin{align}
    \text{supp}(X_1) = \cdots = \text{supp}(X_T) = \mathbb{R} \qquad \text{and} \qquad \forall s \neq t, \ \ P(X_s = X_t) = 0 
  \end{align}
\end{assumption}

\begin{assumption}{}  \label{a:fullrank}
  Nonsingularity of the covariate matrix.
  \begin{align}
    \text{For }2 \leq t \leq T ,
    \quad     
    E[W_tW_t^{T}] \text{is nonsingular}
  \end{align}
\end{assumption}

We should note that assumptions \ref{a:timeindependence} and \ref{a:sequentialindependence} allow for future regressors to be correlated with past shocks. In the context of firm profit maximization, this means that a firm is allowed to choose future inputs based on their previous productivity. In a different context, they would also allow for a dynamic model with $X_t = Y_{t-1}$. Crucially, we have not made any assumptions about contemporaneous independence between random coefficients and regressors, thus allowing for a unobserved heterogeneity of a general type.

Assumption \ref{a:support} is a simplifying ``no-stayer" assumption that will facilitate some of the asymptotic derivations below. Assumption \ref{a:fullrank} is a common technical assumption often used for generic identification in linear models.

\section{Identifying random coefficient moments}

In this section, we show how to identify the $r^{th}$ moments of the random coefficients $(A_{t}, B_{t})$. Broadly, our identification strategy will follow the next steps.

\begin{enumerate}
  \item Identify shock moments $E[U_{t}^{\ell}V_{t}^{r-\ell}]$

  \item Express the conditional moments of the regressand $E[Y_{1}^\ell Y_{2}^{r - \ell} | W_{2i}]$ as a function of conditional moments of random coefficients $E[A_{t}^{\ell}B_{t}^{r-\ell}|W_{2}]$ and shocks $E[U_{t}^{\ell}V_{t}^{r-\ell}]$

  \item Invert the above so as to represent conditional moments of random coefficients as a function of conditional moments of the regressand and shocks $E[U_{t}^{\ell}V_{t}^{r-\ell}|W_{2}]$

  \item Integrate out to get unconditional moments $E[A_{t}^{\ell}B_{t}^{r-\ell}]$
\end{enumerate}

\subsection{First period moments}

Let us start with the first moment ($r = 1$). In this case, we are required to have $T \geq 2$ consecutive waves of panel data.

We begin by simply reminding ourselves of the system of equations for these two periods.
\begin{align}
  \begin{cases}
    %Y_{1} = A_{1} + B_{1}X_{1} + Z_1^{T}\beta_1 \\ 
    %Y_{2} = A_{2} + B_{2}X_{2} + Z_2^{T}\beta_2
    Y_{1} = A_{1} + B_{1}X_{1} \\ 
    Y_{2} = A_{2} + B_{2}X_{2} 
  \end{cases}
\end{align}

Let's expand the coefficients according to our model equations \ref{}.
\begin{align}
  \begin{cases}
    %Y_{1} = A_{1} + B_{1}X_{1}  + Z_1^{T}\beta_1 \\ 
    %Y_{2} = A_{1} + U_{2} + B_{1}X_{2} + V_{2}X_{t} + Z_2^{T}\beta_2 \label{eq:expanded}
    Y_{1} = A_{1} + B_{1}X_{1} \\ 
    Y_{2} = A_{1} + U_{2} + B_{1}X_{2} + V_{2}X_{t} \label{eq:expanded}
  \end{cases}
\end{align}

\paragraph{Step 1} In order to identify shocks, we consider the expectation of the difference between the two periods, conditioning on some point $W_2 = w_2 = (x, x)$.
\begin{align}
  %E[Y_{2} - Y_{1} \ | \  W_{2} = w_{2}] = 
  %E[U_{2}] + E[V_{2}]x_2 + z_2^{T}\beta_2 - z_1^{T}\beta_{1}
  E[Y_{2} - Y_{1} \ | \  W_{2} = w_{2}] = 
  E[U_{2}] + E[V_{2}]x
\end{align}

\noindent The expression above contains two caveats. First, that thanks our full support assumption \ref{a:support}, we can choose any point on $\mathbb{R}^2$ to condition on, including a point where $X_1 = X_2 = x$ as we just did. Second, that Assumption \ref{a:timeindependence} allowed us to drop the conditioning for the shocks.

Now, note that the moments $E[U_2]$ and $E[V_2]$ on the right-hand side are solutions to this minimization problem
\begin{align}
%\min_{a, b, \beta_1, \beta_2}
  \min_{\theta_u, \theta_v}  
  %E[(Y_{2} - Y_{1} - a - bx_2 - z_2^{T}\beta_2 + z_1^{T}\beta_{1})^2 | W_2 = w_2]
  E[(Y_{2} - Y_{1} - \theta_u -  x\theta_v)^2 \  | \ W_2 = w_2]
\end{align}
\noindent and that the solution is unique, by virtue of the strict convexity of the quadratic function and the full rank assumption \ref{a:full_rank}. It follows that the shock moments $E[U_2]$ and $E[V_2]$ are generically identified, and so are $\beta_1$ and $\beta_2$.

\paragraph{Step 2} For this step, we simply have to take expectations of \ref{eq:expanded}, now conditioning on $W_2 = w_2 = (x_1, x_2)$. (Note we have dropped the constraint $X_1 = X_2$).
\begin{align}
  \begin{cases}
    % E[Y_{1}|W_{2}]
    %   = E[A_{1}|W_{2}] + E[B_{1}|W_{2}]x_{1} + z_1^{T}\beta_1 \\ 
    % E[Y_{2}|W_{2}]
    %   = E[A_{1}|W_{2}] + E[U_{2}] + E[B_{1}|W_{2}]x_{2} + E[V_{2}]x_{t} + z_2^{T}\beta_2
      E[Y_{1}|W_{2}]
        = E[A_{1}|W_{2}] + E[B_{1}|W_{2}]x_{1} \\ 
      E[Y_{2}|W_{2}]
        = E[A_{1}|W_{2}] + E[U_{2}] + E[B_{1} \ | \ W_{2}]x_{2} + E[V_{2}]x_{2} 
  \end{cases}
\end{align}

\paragraph{Step 3} Writing the previous equation in matrices after a little rearranging, we get:
\begin{align}
  \begin{bmatrix}
   %  E[Y_{1}|W_{2}] - z_1^{T}\beta_1 \\  
  %  E[Y_{2}|W_{2}] - E[U_{2}] -  E[V_{2}]x_{t} - z_2^{T}\beta_1
    E[Y_{1}|W_{2}] \\  
    E[Y_{2}|W_{2}] - E[U_{2}] -  E[V_{2}]x_{t} 
  \end{bmatrix}
  =
  \begin{bmatrix}
    1 & x_1 \\
    1 & x_2  
  \end{bmatrix}
  \begin{bmatrix}
    E[A_{1}|W_{2}] \\
    E[B_{1}|W_{2}]
  \end{bmatrix}
\end{align}

\noindent As long as $x_1 \neq x_2$, we can invert the first matrix on the right-hand side. 
\begin{align}
  \begin{bmatrix}
    E[A_{1}|W_{2}] \\
    E[B_{1}|W_{2}]
  \end{bmatrix}
  =
  \begin{bmatrix}
    1 & x_1 \\
    1 & x_2  
  \end{bmatrix}^{-1}
  \begin{bmatrix}
    % E[Y_{1}|W_{2}]  - z_1^{T}\beta_1  \\  
    % E[Y_{2}|W_{2}] - E[U_{2}] -  E[V_{2}]x_{t} - z_2^{T}\beta_1
    E[Y_{1}|W_{2}] \\  
    E[Y_{2}|W_{2}] - E[U_{2}] -  E[V_{2}]x_{2}  \label{eq:inversion}
  \end{bmatrix}
\end{align}

\noindent since all the quantities on the right-hand side are known, we have identified all the conditional first moment of our random coefficients outside the line $X_1 = X_2$. Since our ``no-stayer" assumption \ref{a:support} guarantees that a randomly drawn point from the joint distribution of $X_1$ and $X_2$ will not be on the diagonal, we have generic identification.

Conditional second-period moments are also automatically identified, since due to our \ref{a:sequentialindependence} assumption we have that shocks are mean-independent from contemporaneous regressors, which allows us to write
\begin{align}
  E[A_2 | W_2 = w_2] = E[A_1|W_2 = w_2] + E[U_2] \\
  E[B_2 | W_2 = w_2] = E[B_1|W_2 = w_2] + E[V_2] \\
\end{align}

\paragraph{Step 4} The final step is simply to integrate out over the distribution of $W_2$ to get unconditional moments.

\subsection{Second period moments} The identification procedure is analogous, but this time we use the square of our model equations.

\paragraph{Step 1} Identify shock second moments on the diagonal $W_2 = w_2 = (x, x)$ using the following equation
\begin{align}
  %E[Y_{2} - Y_{1} \ | \  W_{2} = w_{2}] = 
  %E[U_{2}] + E[V_{2}]x_2 + z_2^{T}\beta_2 - z_1^{T}\beta_{1}
  E[(Y_{2} - Y_{1})^2 \ | \  W_{2} = w_{2}] = 
  E[U_{2}^2] + E[V_{2}^2]x^2 + 2E[U_{2}V_{2}]x
\end{align}

\noindent The analogous minimization procedure is
\begin{align}
%\min_{a, b, \beta_1, \beta_2}
E[U_2^2], E[V_2^2], E[U_2V_2] =
  \min_{\theta_{uu}, \theta_{vv}, \theta_{uv}}  
  %E[(Y_{2} - Y_{1} - a - bx_2 - z_2^{T}\beta_2 + z_1^{T}\beta_{1})^2 | W_2 = w_2]
  E[(Y_{2} - Y_{1} - \theta_{uu} -  x^2\theta_{vv} - 2x\theta_{uv})^2 \  | \ W_2 = w_2]
\end{align}

\noindent point identification is once more possible because the solution to this minimization problem is unique.

\paragraph{Steps 2-3} Again drop the constraint $X_1 = X_2$, and conditioning on $W_2 = w_2 = (x_1, x_2)$ consider the expectation of the square of our model equations.

\begin{align}
\begin{bmatrix}
  E[A_1^2 | x_{1}, x_{2}] \\ 
  E[B_1^2 | x_{1}, x_{2}]  \\ 
  E[A_1 B_1 | x_{1}, x_{2}]
\end{bmatrix}
=
\begin{bmatrix}
  1 & x_{1}^2 & 2x_{1} \\ 
  1 & x_{2}^2 & 2x_{2} \\ 
  1 & x_{1}x_{2} & x_{1} + x_{2} \\ 
\end{bmatrix}^{-1}
\begin{bmatrix}
  E[Y_1^2 | x_{1}, x_{2}] \\
  E[Y_2^2 | x_{1}, x_{2}] - C_{1i} \\
  E[Y_1 Y_2 | x_{1}, x_{2}] - C_{2i} \\
\end{bmatrix}
\end{align}

\noindent where $C_{1i}$ and $C_{2i}$ involve only quantities that were already estimated in previous steps.
\begin{align}
C_{1i} &= E[U^2] + 2E[U_2 V_2] x_{2} +  E[V_2^2]x_{2}^2 \\
      &- 2\{
E[A_1 | x_{1}, x_{2}] +
E[B_1 | x_{1}, x_{2}]x_{2}
\}
\{
E[U_2] + E[V_2]x_{2}
\}
\\
C_{2i} &= 
\{ 
  E[A_1 | x_{1}, x_{2}] + E[B_1 | x_{1}, x_{2}]x_{1}
\}
\{
E[U_2] + E[V_2]x_{2}
\}
\end{align}

Note that the relevant matrix above is invertible whenever $x_1 \neq x_2$.

\paragraph{Step 4} Analogous to step 4 in the first moment case.

\subsection{Identifying further moments}

The argument above generalizes for arbitrary moments. In fact, if we are only after moments a finite number of moments of the distribution, then assumption \ref{a:sequentialindependence} as follows. Let $r$ be a positive integer and for $t \geq 2$ define $W_t = (X_{t-1}, X_{t})$. We will say that $(U_t, V_t)$ are \emph{r\textsuperscript{th}-moment independent} of $W_t$ if 
  \begin{align}
    \text{For }1 \leq j \leq r,
    \quad     
    E[U_{t}^{r-j} V_{t}^{j} | W_t] = E[U_{t}^{r-j} V_{t}^{j}]
  \end{align}
\end{definition}

We can replace assumption \ref{a:sequentialindependence} by the following weaker assumption.

\paragraph{Assumption 3.1'} \label{a:sequentialindependence}
  \textot{Shocks are r\textsuperscript{th}-moment independent from all covariates}
  \begin{align}
    \text{For }1 \leq j \leq r,
    \quad     
    E[U_{t}^{r-j} V_{t}^{j} | W_t] = E[U_{t}^{r-j} V_{t}^{j}]
  \end{align}

\subsection{Extension: fixed coefficients}

Our model can be augmented with an arbitrary number of regressors associated with coefficients that are individual-invariant -- but not necessarily time-invariant. The model equations become
\begin{align*}
Y_{t} &= A_{t} + B_{t}X_{t} + Z_{t}^T \beta_{t} \qquad \beta_{t}: \text{fixed}
\end{align*}

Identification of the fixed coefficients happens on Step 1, when the minimization problem becomes
\begin{align}
 E[U_2], E[V_2], \beta_1, \beta_2  = \min_{\theta_u, \theta_v, \theta_{b_1}, \theta_{b_2}}  
  E[(Y_{2} - Y_{1} - a - bx_2 + z_2^{T}\theta_{b_1} - z_1^{T}\theta_{b_2})^2 | W_2 = w_2]
\end{align}
and the argument follows for the same reasons that were previously explained. Once we have identified $\beta_1, \beta_2$, we remove them from future computations using the modified regressand $\tilde{Y_t} := Y_{t} - Z_{t}\beta_{t}$, and the rest of the process delineated above goes through without further change.

\subsection{Identifying the marginal distribution of random coefficients}

In this section we will prove that we can identify the marginal density of random coefficients $f_{A_tB_t}$. We begin by strengthening the

%--------------
% ESTIMATION
%-------------

\section{Estimation} \label{sec:estimation}

Our estimation procedure follows three steps:

\begin{enumerate}
  \item Use local polynomial regression to estimate shock moments and individual-invariant coefficients
  \item Use any nonparametric estimator to retrieve estimates of random coefficient moments conditional on observables
  \item Average out conditional random moments by averaging out conditional moments, taking care to avoid
  observations for which $X_1 \approx X_2$.
\end{enumerate}

Let us consider each one of these steps in detail.

\subsection*{Estimating shocks}

\paragraph{First moments} Regress $\Delta Y_2$ on $X_2$ locally at the diagonal $X_1 \approx X_2$:
\begin{align}
  \hspace*{-1cm}
  (\widehat{E}[U_2], \widehat{E}[V_2], \ \cdot \ ) = \arg \min_{\theta_{U} ,\theta_{V}, \gamma}
\sum_{i} K_h (\Delta X_{2i}) \cdot 
      \left( \Delta Y_{2i} - \theta_{U} - X_{2i}\theta_{V} - g_1(X_{2i}, \Delta X_{2i}; \gamma \right)^2
\end{align}

\noindent where $K_h(\cdot)$ is a standard kernel with asymptotically zero bandwidth:
\begin{align}
  h_{shocks}(n) = c_{shocks}n^{−\alpha_{shocks}} \qquad \text{where } \lim_{n\rightarrow \infty} h_{n} = 0
\end{align}
\noindent and $g_1$ is a $K$-order polynomial in $X_2$ and $\Delta X_2$ with coefficients $\gamma$. In theory its inclusion is optional, but we have observed that its presence substantially improves the quality of the estimates.

If our model included time-invariant coefficients, they would also be estimated in this step.

\paragraph{Second moments} We estimate the uncentered moments by proceeding similarly to above:
\begin{align}
  (\widehat{E}[U_2^2], \widehat{E}[U_2 V_2] &\widehat{E}[V_2^2], \ \cdot \ ) = \\  
  \arg \min_{\theta_{U_2} , \theta_{UV}, \theta_{V_2}, \gamma}
\sum_{i} K_h(\Delta X_{2i}) &\cdot 
      \left( \Delta Y_{2i}^2 - \theta_{U_2} - 2X_{2i}\theta_{UV}  - X_{2i}^2\theta_{V_2} - g_2(X_{2i}, \Delta X_{2i}; \gamma) \right)^2
\end{align}

Naturally, once we have uncentered second moments and first moments, we can compute estimates of centered moments using the usual formulas for centered moments, e.g. $\widehat{Var}[U_2] = \widehat{E}[U_2^2] - \widehat{E}[U_2]^2$.

\subsection*{Conditional random coefficient moments}

\paragraph{Conditional moments of $Y_1$ and $Y_2$} Here we begin by estimating $E[Y_1^{m} Y_2^{n}|X_1 = x_1, X_2 = x_2]$ for $m,n \in \{ (1,0), (0,1), (1,1), (2,0), (0,2)\}$. This can be done using any nonparametric estimator, but we choose the Nadaraya-Watson for simplicity.
\begin{align}
  \widehat{E}[Y_1|X_1 = X_1, X_2 = X_2] =
  \frac{\sum_i Y_{1i}K_h(X_{1i} - x_1) K_h(X_{2i} - x_2)}
  {\sum_i K_h(X_{1i} - x_1) K_h(X_{2i} - x_2)}
\end{align}

\noindent where $K_h$ is again a standard kernel endowed with asymptotically vanishing bandwidth.
\begin{align}
  h_{nw}(n) = c_{nw}\hat{\sigma} n^{-\alpha_{nw}}
\end{align}

\noindent where $\hat{\sigma}$ is an estimate of $Std(X_1) = Std(X_2)$. Note that for the theoretical MISE-minimizing bandwidth, one should set $\alpha_{nw} = \frac{1}{6}$.

\paragraph{Solving for first moments} Once in possession of all the previous estimates, we can solve for estimates of conditional random coefficient moments using the empirical analogs of equations \ref{eq:first_moment_inversion} and \ref{eq:second_moment_inversion}.

\subsection*{Unconditional random coefficient moments}

In this final step, we must exclude observations near the diagonal $X_1 ≈ X_2$, and then average the conditional moments associated with the remaining observations. According to our asymptotic results, included observations must satisfy the following conditions:
\begin{align}
  |\Delta X_{2i}| &> c_{cens1} \hat{\sigma}_{X_2} n^{-\alpha_{cens1}} \quad \text{ for first moments}\\
  |\Delta X_{2i}| &> c_{cens1} \hat{\sigma}_{X_2} n^{-\alpha_{cens2}} \quad \text{ for second moments}
\end{align}

As shown in section \ref{\asymptotic_results}, the optimal thresholds can be shown to be associated with exponents $\alpha_{cens1} = \frac{1}{4}$ and $\alpha_{cens2} = \frac{1}{8}$. However, theory gives us no guidance regarding the choice of constants $c_{cens1}$ and $c_{cens2}$, so we resort to experimentation in the next section.

% -------------
%  Simulations
% -------------

\section{Simulations} \label{sec:simulation}

The purpose of this sections is to provide a simulation study in a setting that is similar to our empirical application. We begin by generating the first period random coefficients $(A_1, B_1)$, their second-period shocks $(U_2, V_2)$,
and regressors for both periods $(X_1, X_2)$. They are drawn from a jointly Normal distribution:

\begin{align} \label{eq:simulation_model}
\begin{bmatrix}
  A_1\\
  B_1\\
  X_1\\
  X_2\\
  U_2\\
  V_2 
\end{bmatrix}
\sim
\mathcal{N}\left(
\begin{bmatrix}
2 \\ 
1 \\ 
0 \\ 
0 \\ 
.3 \\
.1 
\end{bmatrix}
,
\begin{bmatrix}
  9 & 0.95 & 1.5 & 1.5 & 0 & 0 \\
  0.95 & 0.4 & 0.32 & 0.32 & 0 & 0 \\
  1.5 & 0.32 & 1 & 0.5 & 0 & 0 \\
  1.5 & 0.32 & 0.50 & 1 & 0 & 0 \\ 
  0 & 0 & 0 & 0 & 1 & 0.16 \\
  0 & 0 & 0 & 0 & 0.16 & 0.1 
\end{bmatrix}
\right)
\end{align}

These number were chosen to roughly reflect the characteristics that we expect them to have in a real data
application. The implied correlation matrix has a very simple structure.

\begin{align}  \label{eq:simulation_correlation}
  \begin{bmatrix}
      1&  0.5& 0.5&  0.5&  0&  0 \\
      0.5&  1& 0.5&  0.5&  0&  0 \\
      0.5& 0.5&  1&  0.5&  0&  0 \\
      0.5& 0.5& 0.5&   1&  0&  0 \\
      0&   0&  0&   0&  1& 0.5 \\
      0&   0&  0&   0& 0.5&  1
  \end{bmatrix}  
\end{align}

The remaining variables were created from these in the obvious manner. In order to understand how our estimator behaves as the number of observation increases, we used $n \in \{500, 2000, 5000, 10000, 20000 \}$.

As seen in Section \ref{sec:estimation}, our method requires the careful tuning of many different parameters. Since our methods are not amenable to cross-validation, we simulated the model in section 1 over a large grid of parameter combinations, which were then compared according to their RMSE against the true value. The RMSE-minimizing parameter configuration for each number of observations is shown on Table \ref{tab:best_parameters}.

\begin{table}[htdp]
  \centering
  \renewcommand{\arraystretch}{1.6}
  \begin{tabular}{l|cccc}
    n & $c_{shocks}$ & $\alpha_{shocks}$ & $c_{nw}$ & $\alpha_{nw}$ \\  \hline
    500 & 2 & $\frac{1}{5}$ & $\frac{1}{2}$ & $\frac{1}{6}$ \\
    2000 & 3 & $\frac{1}{5}$ & $\frac{1}{2}$ & $\frac{1}{6}$ \\
    5000 & 4 & $\frac{1}{5}$ & $\frac{1}{2}$ & $\frac{1}{6}$ \\
    10000 & 4 & $\frac{1}{5}$ & $\frac{1}{2}$ & $\frac{1}{6}$ \\
    20000 & 3 & $\frac{1}{5}$ & $\frac{1}{2}$ & $\frac{1}{3}$ \\
  \end{tabular}
  \caption{RMSE-minimizing parameter configurations for our DGP.}
  \label{tab:best_parameters}
\end{table}

Note that the best parameters remain fairly stable as we increase the number of observations. The most notable variation is in $\alpha_{nw}$: intuitively, when the number observations is large enough, a much smaller bandwidth is preferred, and that is translated into a more aggressive choice of exponent. We will use these numbers to guide our choice of tuning parameters in our empirical application in the next section.

The performance of our estimators is illustrated on Figures \ref{fig:simulation_shocks} and \ref{fig:simulation_random_coefficients} and on the accompanying tables on the Appendix. We observe that indeed the distribution tends to collapse to its true value as the number of observations increases.

\subsection{Bootstrap coverage}

Tables \ref{tab:bs_coverage_shocks}-\ref{tab:bs_coverage_random_coefficients} show the coverage of bootstrap confidence intervals for all estimated parameters. To produce these tables, we generated 2000 datasets using different seeds, computed 500 bootstrap estimates in each of these datasets, and calculated their 0.025 and 0.975 quantiles. The numbers in each cell are the proportion of times that quantiles straddled the true value.

\begin{table}[H]
  \singlespacing
  \caption{Bootstrap coverage for shock moments}
  \input{chapter1/tables/bs_coverage_shocks.tex}
  \label{tab:bs_coverage_shocks}
\end{table}

\begin{table}[H]
  \singlespacing
  \caption{Bootstrap coverage for random coefficient conditional moments}
  \input{chapter1/tables/bs_conditional_coverage.tex}
  \label{tab:bs_conditional_coverage}
\end{table}

\begin{table}[H]
  \singlespacing
  \caption{Bootstrap coverage for random coefficient unconditional moments}
  \input{chapter1/tables/bs_coverage_random_coefficients.tex}
  \label{tab:bs_coverage_random_coefficients}
\end{table}

\section{Empirical application}

Now we provide an illustration of our methods by estimating a Cobb-Douglas production function with random elasticity for a large number of Indian plants. We will be using the Annual Survey of Industries (ASI) dataset, released by the Central Statistical Organization of India for the years 2008 and 2009. This yearly survey collects data from sampled Indian economic units of production (individual factories, workshops, and establishments, hereafter called ``plants") that employ more than 10 regular workers. From this data we extract four variables: \emph{gross sale value} (value of the products sold by the plant, as purchased by their clients) $S_{t}$; \emph{capital} (fixed assets with a productive life of more than one year) $K_{t}$; \emph{wages} $W_{t}$; and production \emph{materials} $M_{t}$. All variables are in 2005 rupees. We keep only observations that we present in both years of analysis, and we are left with 13298 observations. Summary statistics are in subsection \ref{sec:summary_statistics} in the Appendix.

In addition to the variables above, we generate our model variables:
\begin{align}
  Y_{t} = \log\left( \frac{S_{t} - M_{t}}{W_{t}}\right) \qquad X_{t} = \log \left( \frac{K_{t}}{W_{t}} \right)
\end{align}
\noindent where $Y_{t}$ represents production value-added after sales normalized by wages, and $X_t$ represents normalized capital.

[[Interpretation]]

\begin{table}[H]
  \singlespacing
  \caption{Empirical application: Shock Estimates}
  \input{chapter1/tables/bootstrap_shock_empirical.tex}
\end{table}

\begin{table}[H]
  \singlespacing
  \caption{Empirical application: Random Coefficient Estimates (first period)}
  \input{chapter1/tables/bootstrap_random_coefficients_1_empirical.tex}
\end{table}

\begin{table}[H]
  \singlespacing
  \caption{Empirical application: Random Coefficient Estimates (second period)}
  \input{chapter1/tables/bootstrap_random_coefficients_2_empirical.tex}
\end{table}

\begin{figure}
  \centering
  \includegraphics[height=0.8\textheight]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter1/figures/simulation_shocks.pdf}
  \label{fig:simulation_shocks}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[height=0.8\textheight]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter1/figures/simulation_random_coefficients.pdf}
  \label{fig:simulation_random_coefficients}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[height=0.8\textheight]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter1/figures/rc1_empirical.pdf}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[height=0.8\textheight]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter1/figures/rc2_empirical.pdf}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[height=0.8\textheight]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter1/figures/shocks_empirical.pdf}
\end{figure}

\clearpage

\section{Appendix}

\subsection{Summary statistics} \label{sec:summary_statistics}

\begin{table}[H]
  \singlespacing
  \caption{ASI survey data: summary statistics}
  \input{chapter1/tables/summary_statistics_raw1.tex}
  \label{tab:summary_statistics_raw1}
\end{table}

\begin{table}[H]
  \singlespacing
  \caption{ASI survey data: correlations}
  \input{chapter1/tables/summary_statistics_raw2.tex}
  \label{tab:summary_statistics_raw2}
\end{table}

\begin{table}[H]
  \singlespacing
  \caption{Transformed variables: summary statistics}
  \input{chapter1/tables/summary_statistics_processed1.tex}
  \label{tab:summary_statistics_processed1}
\end{table}

\begin{table}[H]
  \singlespacing
  \caption{Transformed variables: correlations}
  \input{chapter1/tables/summary_statistics_processed2.tex}
  \label{tab:summary_statistics_processed2}
\end{table}

\subsection{Simulation results}

Here we show how our estimates of shock and random coefficient moments change as we increase the number of observations. Throughout, we used the RMSE-minimizing parameters shown in table \ref{tab:best_parameters}. To produce the tables below, we generated 8000 datasets for each number different number of observations, produced the relevant estimates and computed the performance measures.

\subsubsection*{Shock moments}

\begin{table}[H]
  \singlespacing
  \caption{Performance measures for estimates of E[U_2]}
  \input{chapter1/tables/simulation_statistics_EU.tex}
\end{table}

\begin{table}[H]
  \singlespacing
  \caption{Performance measures for estimates of E[V_2]}
  \input{chapter1/tables/simulation_statistics_EV.tex}
\end{table}

\begin{table}[H]
  \singlespacing
  \caption{Performance measures for estimates of Std[U_2]}
  \input{chapter1/tables/simulation_statistics_SU.tex}
\end{table}

\begin{table}[H]
  \singlespacing
  \caption{Performance measures for estimates of Std[B_2]}
  \input{chapter1/tables/simulation_statistics_SB.tex}
\end{table}

\begin{table}[H]
  \singlespacing
  \caption{Performance measures for estimates of Cov[U_2, V_2]}
  \input{chapter1/tables/simulation_statistics_CUV.tex}
\end{table}

\begin{table}[H]
  \singlespacing
  \caption{Performance measures for estimates of Corr[U_2, V_2]}
  \input{chapter1/tables/simulation_statistics_CorrUV.tex}
\end{table}

\subsection*{Random coefficient moments}

\begin{table}[H]
  \singlespacing
  \caption{Performance measures for estimates of E[A_1]}
  \input{chapter1/tables/simulation_statistics_EA.tex}
\end{table}

\begin{table}[H]
  \singlespacing
  \caption{Performance measures for estimates of E[B_1]}
  \input{chapter1/tables/simulation_statistics_EB.tex}
\end{table}

\begin{table}[H]
  \singlespacing
  \caption{Performance measures for estimates of Std[A_1]}
  \input{chapter1/tables/simulation_statistics_SA.tex}
\end{table}

\begin{table}[H]
  \singlespacing
  \caption{Performance measures for estimates of Std[B_1]}
  \input{chapter1/tables/simulation_statistics_SB.tex}
\end{table}

\begin{table}[H]
  \singlespacing
  \caption{Performance measures for estimates of Cov[A_1, B_1]}
  \input{chapter1/tables/simulation_statistics_CAB.tex}
\end{table}

\begin{table}[H]
  \singlespacing
  \caption{Performance measures for estimates of Corr[A_1, B_1]}
  \input{chapter1/tables/simulation_statistics_CorrAB.tex}
\end{table}

\subsection{Python module \texttt{FHHPS}}

In order to facilitate the adoption of the methods described in this paper by other researchers, I have developed the Python module \texttt{FHHPS}. Instructions for installations and usage can be found in the github repository\footnote{Url: https://github.com/halflearned/FHHPS}. Its API is inspired by the celebrated \texttt{scikit-learn} library\cite{sklearn_api}, and it should be familiar to statisticians and machine learning practitioners that use Python. The webpage linked above also contains one-line instructions for reproducing all of our figures and tables.

\clearpage
\begin{landscape}
\begin{table}[H]
  \singlespacing
  \caption{Panel regressions}
  \input{chapter1/tables/panel_lm_table.tex}
  \label{tab:panel_lm_table}
\end{table}
\end{landscape}

### 2018-04-04 chapter2/chapter2.tex
etc

\chapter{Two novel algorithms for dynamic kidney exchange}

\doublespacing

\section{Introduction}

Patients suffering from end-stage renal disease have two available treatments: dialysis and renal transplant. While transplant is associated with ``lower mortality rates and improved quality of life compared to chronic dialysis treatment" \citep{tonelli2011systematic}, severe kidney shortage prevents tens of thousands of patients every year from receiving a transplant. In the US alone, there are currently over 90,000 patients in the kidney transplant waiting list. These patient may wait several years before finding an available donor, and many will die -- often of co-morbidities -- before receiving a transplant.

In order to overcome this challenge, \emph{kidney paired donation} (KPD) has emerged as an alternative option for patients who have an incompatible but otherwise willing living donor. This approach, conceived about three decades ago by \cite{rapaport1986case}, involves pooling patient-donor pairs and swapping or exchanging donors so as to increase the overall number of matched patients.

[[[History -- Benefits]]]

Nevertheless, even today the amount of kidney exchanges remains smaller than its potential. As we will argue in this paper, one source of inefficiency in kidney exchange is that most transplant centers do not take into consideration the evolution of their kidney pool over time, which can lead to large losses in terms of the overall number of pairs matched over time.

To this end, this paper contributes to the \emph{dynamic} kidney matching literature by introducing two novel algorithms for kidney exchange. They are explained next.

\paragraph{\textbf{Main idea}} This paper proposes two novel approaches for increasing the cardinality of matched pairs in a discrete-time dynamic model of kidney exchange: the \emph{direction prediction} and the \emph{multi-armed-bandit} methods. We empirically evaluate these methods using simulations under a variety of data-generating processes that we call \emph{environments}.

Let's see how these methods work, and some of the challenges they overcome.

First, our \emph{direct prediction} method recasts the dynamic kidney exchange problem as a binary classification task: for each pair in the pool, we predict a binary label representing whether they should be matched today (one), or left for the future (zero), given their observable characteristics. This prediction is produced by an estimator trained in a large a number of simulations that were solved by an offline solver.

The main challenge with this approach is how to represent the data: while covariates such as ABO bloody type and current waiting times can be naturally represented by real numbers, it is not immediately clear how to represent their relationship to other nodes in the graph.  We attempt to overcome this issue by augmenting the data set with graph-theoretic notions such as measures of node centrality, size of graph, average degree, and so on.

Our second method, named the \emph{multi-armed bandit} (MAB) method, uses simulations to answer the question: ``if we commit to matching one particular set of pairs today, how likely are we to decrease the \emph{overall} number of matched pairs between now and a horizon $h$?". This probability can in principle be estimated via simulations. However, the main challenge here is that simulations are so computationally costly, and the number of available actions so vast, that we are cannot realistically produce accurate estimates of this probability for all points in the action space. To overcome this issue, we employ multi-armed bandit algorithms. Their role is to manage a relatively small computational budget and determine which actions are likely attractive (and should be explored further), and which are not (and thus not worth exploring much).

\paragraph{\textbf{Main results}} Our methods are evaluated entirely via simulations. The simulation setups, which we call \emph{environments}, are inspired by models used the kidney exchange literature, and differ by their rules regarding blood- and tissue-type compatibility.

We compare our method against two benchmarks: an algorithm that clears the maximal matching at each period, called \textsc{myopic}; an infeasible offline optimal algorithm that we call \textsc{OPT}. The former roughly represents what most organ clearinghouses are doing today, while the latter represents the maximal payoff we could have achieved, had we been able to know the future. Our metric for comparison is the \emph{per-period average number of matched pairs} over a long period of time.

In all environments, we observe that our implementation of the \emph{direct estimation} method improves upon \textsc{Myopic}, but only under very restricted conditions. However, the \emph{multi-armed bandit} (MAB) method is able to improve uniformly and substantially upon the \textsc{Myopic} benchmark. In particular, we observe that our MAB does particularly well in environments of moderate sparsity.

\subsection{Related literature} This paper pulls ideas from microeconomics, matching theory, and operations research, and uses methods drawn from computer science and machine learning. Let us take a brief look at these fields in turn.

\subsubsection{Kidney Exchange}

\paragraph{Early years} In microeconomics, the literature on kidney exchange began with \cite{roth2004kidney}, who provided a kidney exchange mechanism inspired by the \emph{housing} problem studied in the literature of mechanism design as in \cite{shapley1974cores} and later \cite{abdulkadirouglu1999house}. While their mechanism had significant virtues as being strategy-proof and Pareto-efficient, it also relied on large number of exchanges being conducted simultaneously, and drew criticism for making assumptions about heterogeneous preferences over kidneys. Subsequent work by \cite{roth2005pairwise} addressed some of the criticisms by focusing on logistically simple mechanisms that used only 2-way exchanges, and assumed also that patients were indifferent between all compatible kidneys. In a later paper, \cite{roth2007efficient} demonstrated via simulations that allowing for 3-way exchanges in addition to 2-way exchanges could increase the cardinality of matched pairs by a great deal, but larger cycles would only bring about modest improvements.

[[[More here]]]

\paragraph{Recent advances in static exchange} In the last ten years, both in academia and in medical practice the focus began to shift from exchanges via cycles to exchanges emanating from \emph{non-directed donors} (NDD) -- altruistic donors who are willing to donate their organs to anyone. Such exchanges yield a \emph{chain} of transplants that begins with the NDD and may either terminate with the last pair donating to a waitlist recipient in a kidney registry, or not terminate at all and have the last pair's donor await an opportunity to become a future living donor. In the former kind of exchange, called \emph{domino paired donation} by \cite{montgomery2006domino}, all transplants occur simultaneously. In the second kind, transplants may be spread over several months. For this reason, the latter kind of transplant is called \emph{non-simultaneous, extended, altruistic donor chain} by \cite{rees2009nonsimultaneous}.\footnote{On occasion, NEAD chains have also been called \emph{Never-ending altruistic donor chains}, as relayed by \cite[p.~235-6]{roth2015gets}.}

In recent years, a significant amount of attention has been devoted to these NEAD chains, including notably \cite{ashlagi2011nonsimultaneous} and \cite{ashlagi2012need}, whose results show that NEAD chains benefit highly sensitized patients in sparse pools of moderate size, because they decrease the need for simultaneous double-coincidence of wants in the exchange market. In fact, \cite{anderson2015finding} reports that ``NEAD chains are responsible for the majority of successful kidney transplants conducted via kidney exchange at both the [Allied for Paired Donation] and within other major exchange programs".

\paragraph{Dynamic kidney exchange} The literature concerning the \emph{dynamic kidney exchange} problem, where we take into account the evolution of the kidney exchange pool over time, is much smaller. It began with a seminal paper by \cite{unver2010dynamic}, who derives a dynamic mechanism that produces optimal n-way cyclic exchanges in the steady state of a continuous-time model with Poisson arrivals. The results in that paper rely on three simplifying assumptions. First, that the waiting cost is constant for all pairs; second, that pairs do not leave the pool unless they are matched; finally, that pairs within the pool are only blood-type incompatible. It can be shown that, under these assumptions, all pairs are rendered homogeneous. This dramatically decreases the dimension of the state space, and allows for an easily interpretable solution.

\cite{ashlagi2013kidney} note that the graphs of real-life pools tend to be sparse and filled with pairs that are either easy or extremely hard to match. In order to model this particular feature, they postulate a sparse heterogeneous random graph model containing only two such types, and propose a greedy algorithm that waits until the pool contains a certain number of pairs of each type, and then matches as many as pairs as possible. In a different vein, \cite{akbarpour2017thickness} studies the relationship between ``market thickness" (the number of available pairs in the exchange pool) and matching time in a simple model of stochastic arrival and departure. They show that, in their model, greedy algorithms that cleverly exploit the time to match can can perform close to infeasibly optimal benchmarks, even if these algorithms are ignorant about the global structure of the graph, and have no information about agents' departure times.

In the computer science and operations research literature, \cite{abraham2007clearing} and more recently \cite{anderson2015finding}, \cite{dickerson2016position} and \cite{dickerson2017small} have mostly focused on producing scalable combinatorial programming algorithms that can take on static matching problems with large graphs and allowing for both chains and cycles of moderately large length. However, a series of papers starting with \cite{awasthi2009online} and followed by \cite{dickerson2012dynamic} and \cite{dickerson2015futurematch} have dealt with the dynamic kidney problem by \emph{weighted myopia}. Their idea is to prevent wasteful matchings (e.g., an O-donor to an AB-patient) by artificially introducing negative weights to graph components containing them. In \cite{dickerson2015futurematch}, these optimal weights are computed from simulations involving historical data.

\paragraph{Computer science and machine learning} We were inspired by several computational alternatives to dynamic kidney exchange that have been proposed in the past few years. In particular, the \emph{multi-armed bandit method} described here is reminiscent Algorithm 1 in \citet{awasthi2009online}, inasmuch it also selects the best actions today by repeatedly simulating the future, solving an offline problem, computing a ``score" for each cycle, and selecting actions with maximum score. However, both our simulation and ``scoring" methods differ importantly because we leverage multi-armed bandit algorithms to decrease the computational burden of producing multiple simulations.

\citet{dickerson2012dynamic, dickerson2015futurematch} propose a related method that they call \emph{weighted myopia}. The idea is to use simulations for learning ``potentials" of specific elements of the graph. Much like the ``scores" in \citet{awasthi2009online}, the set of ``potentials" associated with a specific exchange encodes how desirable each exchange is in terms of its future value. The authors then use these pre-learned numbers to revise the weights in a myopic algorithm, forcing it to select more desirable exchanges where it would be otherwise indifferent. Our \emph{direct prediction method} works similarly, with three crucial differences. First, we do not directly predict how ``desirable" a matching will be, but instead we predict whether or not an offline, optimal algorithm would choose the node or not. Second, our method uses a more aggressive thresholding mechanism to select which nodes should be matched today, and which should be left for later. Third, we optionally make use of information about how the node fits inside the compatibility graph, so that the information about the node (e.g., the blood type and HLA profile of patient and donor) is augmented with graph-theoretic notions (e.g., measures of node centrality, degree, etc).

\paragraph{Sequential decision problems} Zooming out of our application, our problem lies within a larger class of sequential decision problems whose central feature is a concern with the \emph{exploration-exploitation trade-off}. That is, problems that can be described as the one faced by an agent who must spend a limited computational budget to \emph{explore} a certain action space so as to find and \emph{exploit} actions that will maximize her expected rewards. Crucially, the agent sequentially observes rewards for actions that she has taken, but does not observe rewards for other actions. This class also encompasses, for example, \emph{Markov decision process} (MDP) and its variants.

A \emph{multi-armed bandit}\footnote{The terminology comes from a turn-of-the-20th-century United States colloquialism, when slot machines were called ``one-armed bandits". While the name apparently suggests a loss in the long run, the ``bandit" problems studied in academic literature do not necessarily have negative expected payoff.} (MAB) problem is a particular version of an MDP where time is discrete, and at every period the agent faces a finite and fixed number of actions $K$ that produce stochastic rewards. The objective is usually to minimize the amount of \emph{regret} that accumulates over time, where \emph{regret} is loosely defined as the difference between the agent's actual accumulated rewards under her own strategy and an the average reward that the agent would have gotten had she chosen the optimal action at each period since the beginning. (We will make this definition more rigorous in a later section). Crucially, at every period the agent only observer rewards for actions that she has chosen, and not for other actions.

Multi-armed bandit problems were studied sporadically in the last century, with the earliest reference going as far back in time as \cite{thompson1933likelihood}. Interest was rekindled after a seminal paper by \cite{lai1985asymptotically}, who developed a strategy that provably attains an asymptotic lower bound of regret. Later, their analysis was simplified by \cite{agrawal1995sample} who also developed a finite-time analysis of regret. In recent years, other alternative algorithms have been proposed. In particular, in our paper we use the \emph{Thompson sampling} algorithm studied by \cite{agrawal2012analysis} and \cite{kaufmann2012thompson}, and the \emph{upper-confidence bound} (UCB1) algorithm developed by \cite{agrawal1995sample} and \cite{auer2002finite}. For an approachable review of multi-armed bandits, optimal strategies, and their variations, we refer the reader to the recent book by \cite{lattimore2018bandits}.

%%%%%%%%%
\section{Background and definitions}

\paragraph{Medical background} We say that a patient and a donor are \emph{compatible} if the donor's organ cells do not present \emph{antigens} that are capable of inducing an aggressive response by the patient's immune system. A successful transplant usually requires patients to be \emph{blood-type compatible} and \emph{tissue-type compatible} (also known as \emph{histocompatible}).

Blood-type compatibility refers to compatibility with respect to major ABO blood groups. For example, $O$-type patients can only receive from $O$-type donors, $AB$ patients may receive from any blood group, etc).

A patient is tissue-type compatible with a donor if the donor does not present alleles of gene complex called the \emph{human leukocyte antigens} (HLA) that are deemed unacceptable by the patient's immune system.

\paragraph{Technical definitions} We model the dynamic kidney exchange problem in a manner similar to \cite{unver2010dynamic} and \cite{akbarpour2017thickness}. A \emph{kidney exchange pool} is a directed random graph process $G_t = (V_t, E_t, X_t), t \in \mathbb{N}$ whose vertices $v \in V_t$ represent (patient, donor)-\emph{pairs}\footnote{Throughout, we will use the terms \emph{pair}, \emph{node} and \emph{vertex} interchangeably.}. A directed edge $e \in E_t$ between vertex $v$ and $v'$ means that the donor in pair $v$ can donate to the patient in pair $v'$. When such an edge exits, we say that \emph{$v$ is compatible with $v'$}.

Each pair is endowed with certain characteristics $x \in X_t$\footnote{In an slight abuse of notation, we will also denote by $X_t$ the matrix of pairs' observable characteristics.} such as patient and donor blood type (other characteristics will be discussed below).

We say that a pair \emph{enters} the kidney exchange pool when it first becomes available for exchange. After a certain number of periods, the pair \emph{leaves} the pool once and for all, or \emph{dies}. The difference between entry and death is called a pair's \emph{sojourn}.

An \emph{exchange} is an ordered tuple of nodes $m = (v_{i_1}, \cdots, v_{i_{k}})$ where each pair in the tuple is compatible with the next pair in the sequence. A \emph{matching} is a set of exchanges where no pair appears in more than one exchange.

An \emph{environment} is, informally speaking, a collection of rules governing which pairs are deemed compatible, the entry and death processes that govern the evolution of the pool, and which vertex characteristics are observable in and relevant to the problem. More formally, it can be defined as the conditional probability distribution between two kidney exchange pools, given the current pool $G_t$ and $G_t'$, given a matching $M_t$. We will expand on this below. Finally, a \emph{dynamic matching algorithm} is a procedure that selects matchings $M_t$ at every period. Once a matching $M_t$ is selected, all vertices and their edges are removed from the kidney exchange pool.

%It includes the rules for which pairs are deemed compatible, the entry and death processes that govern the evolution of the pool, which characteristics are observable to the algorithm at each period, and the reward observed by the agent as sequentially matches and clears pairs. More rigorously, at each period an environment is described bu tuple $(G, u)$, where:

% 
% \begin{itemize}
%   \item $G = (V, A, X, \pi)$ is a graph with vertex set $V$, adjacency matrix $A$, vertex characteristic matrix $X$, and $\pi: \mathbb{N} \times V \rightarrow \{0, 1\}$ is a function that maps a period $t \in \mathbb{N}$ and a vertex $v \in V$ to whether that vertex was alive on period $t$ (1) or not (0). \footnote{We decided to model dynamic graphs in this manner as it is closer to our simulation code. Note that all the information about entries and exits can be expressed by the function $\pi$.}
%   \item $u: 2^V \rightarrow \mathbb{R} \cup \{-\infty\} $ is the reward associated with matching subset of $V$. Infeasible exchanges receive reward $-\infty$.
% \end{itemize}

\subsection{Environments}

Our paper presents three environments inspired by previous works on dynamic kidney exchange. All three have the following assumptions in common.

First, the number of new incoming pairs in each period is drawn from the $Poisson(r)$ distribution, where $r \in \mathbb{N}$ denotes the \emph{entry rate}, and also equals the expected number of entrants per period. Second, each pair independently draws the length of their sojourn from the $Geometric(d)$ distribution. The parameter $d \in \mathbb{R}$ is the \emph{death rate}, and its reciprocal $\frac{1}{d}$ is the expected sojourn length. We note that, due to the memoryless property\footnote{If $X \sim Geometric(p)$, then $P(X > t+s | X > s) = P(X > t)$} of the Geometric distribution, the amount of time a pair has waited in the pool gives us no information about how much time they have until their death. Third, we assume that patients do not discriminate between compatible kidneys. This last assumption could also be understood as patients having binary preferences over kidneys (i.e., they receive utility 1 if they receive a transplant, and 0 otherwise). This assumption is consistent with other works in the literature, notably \cite{roth2005pairwise}.

In what follows we will explain the differences between each environment.

\subsection{ABO Environment}

In the \emph{ABO environment}, compatibility between two distinct pairs is based only on blood-type compatibility.

Blood types are drawn independently for patients and donors from the corresponding probability distribution in the US population\footnote{Roughly $O$:$49\%$, $A$:$36\%$, $B$:$11\%$, $AB$:$4\%$}, but are adjusted by the following assumption previously used in \citet{unver2010dynamic}: we allow for incompatibility between a donor and their own patient. The reason for this additional assumption is that, if there were truly no tissue type compatibilities, we would never observe pairs of type $(AB,\cdot)$, $(\cdot, O)$, or $(A,A)$, $(O,O)$, $(B,B)$, and $(AB,AB)$\footnote{For shorthand, we will sometimes write ``An (X, Z) pair" to mean ``any pairs where the patient has blood type X, and the donor has blood type Z".}., since their donors would be automatically compatible with their patients and they would never participate in an exchange. Following \citet{zenios2001primum}, we assume that for such patients the probability that a donor and their patient are incompatible is $p_c = 0.11$. Arrival rates are then adjusted accordingly, e.g., the arrival rate of $(A,O)$ pair is proportional to $0.48 \times 0.36 \times 0.11 \approx 0.019$. The exact probability distribution for all types can be found in the appendix.

\subsection{RSU Environment}

The \emph{RSU} environment is named after a classic simulation model in \citet*{roth2007efficient} and \cite{saidman2006increasing}. Each pair is characterized by patient and donor ABO blood types, current waiting time, and a \emph{calculated panel reactive antibody (cPRA)} level that represents the probability of a crossmatch with a random donor.\footnote{The original \citet{roth2007efficient} paper called this simply \emph{PRA}, and in real life there is an important distinction between the two measures. However, for the purposes of a simulation model this distinction is immaterial. We keep the name \emph{cPRA} for consistency with OPTN environment later.} The lower the cPRA, the higher the number of potentially compatible pairs.

The simulation process is as follows. First, we draw a pair in the same manner as in the ABO environment. Next, we draw if the patient is a female (with probability around 41\%), and if so we also draw whether her donor is her husband (spouses comprise about 49\% of donors). Finally, we draw a  cPRA level for the patient (Low: 70.1\%, Medium: 20\%, High: 9.9\%). This cPRA level determines the probability that they can receive a kidney from any donor, including their own: patients with low cPRA have a 5\% probability of positive crossmatch with a random donor; patients with medium cPRA have a 45\% chance, and patients with high cPRA have a 90\% chance of a crossmatch. If a patient is bloody or tissue-type incompatible with their own donor, they enter the pool. In addition, if the patient is female and her husband is the donor, the probability of positive crossmatch for low, medium and high cPRA patients goes up to 28.75\%, 58.75\% and 92.25\%. This last adjustment reflects the fact that women tend to produce antibodies against their husbands' antigens during pregnancy.

Once in the pool, the pair immediately forms directed edges with the existing pairs, again following the patient cPRA distribution. The resulting random graph is akin to a Erd\"{o}s-R\'{e}nyi $G(n,p)$ random graph where the probability of forming edges is heterogeneous across different pair types.

\subsection{OPTN Environment}

In the \emph{OPTN environment}, we use historical data collected by the United Network for Organ Sharing (UNOS) data provided in the Standard Research and Analysis (STAR) dataset. The STAR dataset contains information from all patients that were ever registered to the kidney waiting list in the United States for the past three decades, as well as from all living donors that actually participated in an transplant\footnote{To our knowledge, there is no centralized dataset containing information about registered donors that never went to transplant.}. From this original dataset, we excluded entries associated with the following:

\begin{itemize} 
  \item Patients that were registered for more than one organ (including those who were simultaneously waiting for kidney and pancreas)
  \item Patients that were not waiting for their first kidney transplant
  \item Donors and patients with incomplete tissue-type profile information.

The resulting dataset contained 117813 patients and 9337 living donors. We call this the ``historical dataset".

\paragraph{Patient and donor cPRA} We added two additional variables to the original dataset: a \emph{patient cPRA} and a \emph{donor cPRA}. While the patient cPRA is a measure of patient tissue-type incompatibility with a random donor \cite{cecka2010calculated}, our \meph{donor cPRA} is a novel measure of the opposite direction -- how frequently a donor is tissue-type incompatible with a random patient.\footnote{We thank Itai Ashlagi for the suggestion of a donor cPRA.} Both were computed empirically: for each patient our dataset, we checked the how many donors exhibited antigens that are unacceptable for the patient in any of the A, B, Bw, C, DR, DPB, DQ, and DQA loci, and assigned this positive crossmatch probability as their patient cPRA\footnote{A genetic \emph{locus} is a specific position on a chromossome. The loci above encodes a donor tissue type.}; for the donors, we worked in the opposite direction by calculating the frequency of patients who exhibited antibodies against their donor's antigens, and that became their donor cPRA. Figure \ref{fig:cpra} shows the distribution across the entire population.\footnote{We should remark that we found a very different patient cPRA distribution than the one in the OPTN dataset. This may have been because: in real life cPRA is computed using deceased donor data, while we used living donor data; we may have used different HLA equivalence tables; OPTN uses a more sophisticated model based on population genetics.\cite{optn2013cpra}}

\begin{figure}
\centering
\includegraphics[width=\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/computed_cPRA.pdf}
\caption{\textbf{Patient and donor cPRA} \\
    Computed from historical data. Our patient cPRA is the frequency of living donors that exhibit antigens that are unacceptable to the patient. We also define a donor cPRA by calculating the frequency of patients that have antibodies against the donors antigens.}
\label{fig:cpra}
\end{figure}

\paragraph{Artificial dataset} We created an artificial dataset by randomly drawing patients and living donors from the historical dataset and checking for blood-type compatibility, and tissue-type compatibility as explained above. Compatible pairs were discarded. We iterated in this manner to construct a dataset of about one million incompatible pairs. At every simulation period, a random number of pairs is drawn from this dataset.

%%%%%%%%
\section{Methods}

\subsection{Objective and benchmarks}

\paragraph{Objective} In this work, we focus on maximizing the undiscounted cardinality of matched pairs over $T$ periods. To make comparison to our benchmarks easier, we formulate the problem equivalently as maximizing the per-period average matching size over $T$ periods.

\paragraph{Benchmarks} Let $\mathcal{M}$ a set of available matchings at period $t$.
\begin{align}
  \max_{M \in \mathcal{M}}  &\sum_{m \in \mathcal{M}} w_{m} x_{m} \\
  \text{s.t.} \qquad &\sum_{v \in m} x_{m} \leq 1  \qquad   \forall v \in V\\ \label{eq:constraint}
        &x_{m} \in \{ 0, 1 \} \qquad   \forall m \in \mathcal{M}   
\end{align}
\noindent where $c \subset V$ is an exchange, $w_{c}$ is the cardinality of the exchange, and $x_c$ is a binary variable indicating whether or not the pair was selected. Constraint \ref{eq:constraint} ensures that each vertex is selected only once.

By an \emph{static matching problem at period $t$}, we mean a version the problem above where the only available matchings involve pairs $v \in V_{t}$. By an \emph{offline matching problem between $t$ and $s$}, we mean a version of the same problem where involving vertices $v \in \cup_{k=0}^{s} V_{t+k}$, and any matching $m = (v_{1}, \cdots, v_{k})$ has the property that the sojourns of the donating pair overlaps with the sojourn of the receiving pair.

We are interested in how our new methods perform relative to the following two benchmarks.

\paragraph{\textsc{Myopic}} At every period, the \textsc{Myopic} algorithm solves a static matching problem, finds the maximal matching and clears it immediately. In doing so, it disregards all observable characteristics of each pair, and in particular it ignores that some pairs might be useful to keep certain pairs may be easier or harder to match. Therefore, it may forgo the opportunity of matching a hard-to-match patient today, or postpone an easy-to-match pair for later. In essence, this is an approximation to what kidney exchanges currently do.

\paragraph{Optimal (\textsc{OPT})} This infeasible algorithm (henceforth OPT) solves the offline matching problem encompassing all periods between $1$ and $T$. This completely does away with the uncertainty arising from the temporal structure of the problem, hence we know that this is the maximum achievable utility

\paragraph{\textbf{Remark}} \textit{How much is there to be improved upon?} In Figure \ref{fig:greedy_opt_comparison}, we compare \textsc{Myopic} and \textsc{OPT} for a grid of different entry and death rates. These results show that the performance gap between the two can be fairly large, in particular in sparser environments like \emph{RSU} and \emph{OPTN}. We also note that the gap is narrower when: the death rate is high, because if most pairs will die soon, dynamic considerations play a smaller role; and when the entry rate is very large, because in a thicker market pairs are able to encounter a suitable match more easily.

\begin{figure}[htbp]
\centering
\hspace*{-3.5cm}
\includegraphics[width=1.4\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/greedy_opt_comparison_mcl2_2.pdf}
\caption{\textbf{Comparing \textsc{Myopic} and \textsc{OPT}}  \\
Ratio of average per-period matched pairs for different entry and death rates, over 3000 periods (darker hues are better). \texttt{Myopic} has better chances of achieving performances similar to \texttt{OPT} when the death rate is high (moving rightwards on the graphs), or when entry rate is high (moving downwards on the graphs).}
\label{fig:greedy_opt_comparison}
\end{figure}

% ---------------------------------------------------------------------------
% Algorithms
% ----------------------------------------------------------------------

\section{Algorithms}

We now present two novel methods to determine which patients should be matched.

\subsection{Direct prediction} \label{subsec:direct_prediction}

We can break down the dynamic kidney problem into two parts. The first is determining \emph{which} pairs should be matched today, and the second is deciding \emph{how} the selected pairs should be matched among themselves. Note also, that the second part of the problem can be solved immediately as an integer programming problem.

This is the insight that our \emph{direct prediction} method exploits. Essentially, we can reduce the problem to a classification task: at each period, we aim to produce a binary label for each node indicating whether is should be matched in this period (1) or left for later (0). Selected nodes are then passed to a static solver that finds the maximal matching among them. Once these nodes are cleared, time evolves to the next period.

The procedure is formalized in Algorithm \ref{alg:direct_prediction}, and also illustrated in Figure \ref{fig:direct_prediction}.

\begin{figure}[htbp]
\centering
\includegraphics[width=\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/direct_prediction.pdf}.
\caption{\textbf{Direct prediction method} \\ One period of simulation using direct prediction methods to choose cycles. See Algorithm \ref{alg:direct_prediction} for details.}
\label{fig:direct_prediction}
\end{figure}

\begin{algorithm}[htbp]
	\SetAlgoLined
  \DontPrintSemicolon

  \KwData{Environment simulator object \textsc{Env}; \\
          Integer programming solver object \textsc{Solver}; \\
          Statistical method \textsc{Classifier} \\
          Threshold $thres$; }

  \SetKwFunction{FChoose}{\textsc{direct\_prediction}}
  \SetKwProg{Fn}{Function}{:}{}
  \Fn{\FChoose{\textsc{Env}, \ \textsc{Solver}, \ \textsc{Classifier}, \ $thres$}}{
    \tcp{Retrieve node (and potentially also graph) data from current environment} \\
    $X, E_1, E_2$ \gets \textsc{Env.get\_data}()  \\
    \tcp{Classifier predicts matching probability for each pair using data} \\
    $prob$ \gets \textsc{Classifier}($X$, \ $E_1$, $E_2$)  \\
    \tcp{Get index of pairs whose probability is higher than threshold} \\
    $index$ $\gets$ \textsc{which}($prob > thres$) \\
    \tcp{Find maximal matching restricted to this subset} \\
    chosen\_cycles $\gets$ \textsc{Solver.solve(Env, subset=$index$)}  \\  
    \tcp{Return chosen cycles to be cleared} \\
    \KwRet chosen\_cycles \;
  }
	\caption{\textsc{direct\_prediction} Method}
	\label{alg:direct_prediction}
\end{algorithm}

\paragraph{Training the classifier} In order to produce data to feed into our ``classifier", we repeatedly created 1000-period simulation runs and solved them offline using \textsc{OPT}. Then, for every period $t$ in simulation $k$, we stored: a matrix $X_t^{k}$ whose rows represent each pairs observable characteristics; an additional matrices $E_{t}^{1k}$ containing additional graph-theoretical information such as several centrality measures (betweenness, in-degree, out-degree, harmonic, closeness), in- and out-degrees, and average neighbor degree; a conforming matrix $E_{t}^{2k}$ containing the entry and death rates used in that simulation; a binary vector $y_t^{k}$ indicating which nodes \textsc{OPT} chose to match in period $t$. Finally, we vertically concatenated each one of these matrices to create the final data set.

We repeated this procedure for each environment, and produced three artificial data sets, each containing approximately 2 million observations. These data sets were then fed to a series of predictive algorithms: penalized logistic regression \citep{wu2009genome}, random forest classifiers \citep{breiman2001random}, and gradient tree boosting classifiers \citep{friedman2001greedy}.

\subsection{Multi-armed bandit methods}

Since we are assuming access to the exact data-generating process, we can use simulations to choose the best action at each period $t$: choose a certain action; simulate the evolution of the kidney exchange pool until some time horizon $t + h$; solve the offline problem; evaluate the performance of the chosen action under some criterion; repeat until[[[[]]]]

However, the approach outlined above turns out to be naive because simulations are computationally expensive, and in practice we cannot repeat them enough times get reliable estimates of the average performance for each cycle choice, especially in large graphs. It is also incomplete because it does not specify exactly what is the best information we should extract from \texttt{OPT} results. In order to solve the first problem, we leverage theory and algorithms from the multi-armed bandit (MAB) literature. For the second, we propose a secondary objective based on what we call \emph{pseudo-rewards}.

Our procedure is illustrated in Figures \ref{fig:mab} and \ref{fig:mab_step}. At the beginning of the period $t$, the agent receives a set of cycles $C$ that are available to be cleared. If $C$ is empty, nothing happens and we move to the next period. Otherwise, the agent then picks a cycle $c \in C$ and simulates the future, including new entries and deaths, up to a horizon $h$. Next, \texttt{OPT} is run twice, once normally, and once with the additional constraint that $c$ be removed today. The size of the resulting matching in these two scenarios is compared. Naturally, the constrained version of \texttt{OPT} cannot achieve anything better than its unconstrained counterpart, but it might get to be equal. If it is, the agent receives a \emph{pseudo-reward} of one, otherwise it receives zero. This process is repeated: at each iteration $\ell$, a cycle $c_{\ell}$ is chosen and its pseudo-reward $r_{c, \ell}$ is revealed. When a preset computational budget of $L(|C|)$ iterations is hit, the agent then analyses the whole history of cycle choices and pseudo-rewards $H_{t} = \{ (c_{\ell}, r_{c\ell} ) \}_{\ell=1}^{L}$, and decides whether to match one of the cycles or move on to the next period. If a cycle $c$ is chosen it is immediately cleared, however the environment does not evolve to the next period yet. Instead, the history $H_t$ is discarded the procedure is repeated again with a reduced set of actions $C' \subset C$ that produces a new history $H_t'$ and so on, until either there are no more available choices or the agent decides to allow the environment to move on to time $t+1$. When that at last happens, entries and deaths are revealed, the agent receives a new set of cycles, and the process begin anew. All past information is ignored.

Two important details were left out of the explanation above. First, how does the agent chooses the next cycle to test? Second, how does it decide which cycle to choose (or no cycle at all)?

Let $c^{*}$ be a cycle that maximizes expected pseudo-rewards during one round of the algorithm:
$$c_{\ell}^{*} \in \arg\max_{c} E[r_{c,\ell}] $$
 Also, let \emph{regret} be defined as the difference between expected reward of the optimal choice $c^*$ and its own selected choice $c_{\ell}$.
 $$\Delta_\ell := E[r_{c^{*},\ell}] - E[r_{c,\ell}]$$

 A \emph{multi-armed bandit} (MAB) algorithm is an adaptive exploration procedure that seeks to minimize the cumulative regret over $L$ rounds. A good MAB algorithm will act so as to balance exploration (trying out different choices to get high-quality estimates of their rewards) and exploitation (using out better choices more often to increase total rewards), and produce regret that grows at an asymptotically slow rate. Here, we experiment with two common bandits algorithms, namely \emph{UCB1}, and \emph{Thompson sampling}. The literature on bandit algorithms is extensive and an in-depth explanation is outside the score of this paper. However, for context in the next paragraphs we provide some intuition for how and why they work.

\paragraph{UCB1} The \emph{Upper Confidence Bound 1} prescribes that at each period of repeat the following two steps.
\begin{enumerate}
    \item Construct a certain confidence interval around the average reward estimate for each arm. 
    \item Choose the action with the highest confidence upper bound.
\end{enumerate}

This heuristic is commonly named \emph{optimism in the face of uncertainty}\citet{kaelbling1996reinforcement}, because at every period we are choosing the action with the ``largest plausible"\cite[Ch. 3]{lattimore2018bandits} average reward estimate. The intuition is the following: if the agent is correct in choosing the optimistic action, then they receive zero regret; if the agent is incorrect, then they will thereafter review their estimates so as to decrease the upper bound for that action and avoid it in the future.

\cite{auer2002finite} proved that, in the absence of further information about the distribution of rewards, by appropriately constructing the confidence interval around each reward average we are able to attain an optimal logarithmic cumulative regret rate. Their proposed confidence interval for action $c$ in round $\ell$ has the form
\begin{align}
UCB[\ell, c] = \hat{\mu}_{c,\ell-1} + \sqrt{\frac{2\log(\ell)}{n_{c,\ell-1}} }   
\end{align}
\noindent where $\ell$ is the current round of the bandit algorithm, $\hat{\mu}_{c,\ell-1}$ is the running reward estimate for action $c$, and $n_{c,\ell-1}$ is the number of times that the action $c$ has been selected before round $\ell$. It is instructive to remark that the expression above is increasing in $\hat{\mu}_{c,\ell}$ -- so that actions with higher expected reward will be naturally chosen more often, leading to exploitation -- but it is also decreasing in $n_{c,\ell-1}$ -- making other actions more likely to be chosen in the future, leading to more exploration.

\paragraph{Thompson sampling} Also known as \emph{posterior sampling}, this algorithm reformulates the bandit problem in a Bayesian framework. The agent begins with a prior distribution $P(\mu_c)$ for each action $c$. This initial prior is updated as the history of actions and rewards accumulates. Like UCB1, the entire algorithm is explained in few steps:
\begin{enumerate}
  \item Compute and sample from the posterior distribution of average rewards 
  $$\tilde{\mu}_c \sim P(\mu_c | H_{\ell-1}), \forall c$$
  \item Choose the action with maximal reward among the samples.
  $$c_{\ell} = \arg\max_c \mu_c$$
\end{enumerate}

This simple algorithm can be shown to be optimal in terms of attaining a asymptotically logarithmic lower bound on Bayesian regret\cite{slivkins2018introduction}. However, more surprisingly, \cite{kaufmann2012thompson} showed that it also enjoys good frequentist properties, and under certain conditions performs no worse than UCB1 in terms of \emph{frequentist} regret. Moreover, Thompson sampling algorithm can be generalized in a number of ways (some of which we discuss in section \ref{sec:extensions}).

\begin{algorithm}[htbp]
 \SetAlgoLined
 \KwData{Environment simulator object \textsc{Env}; \\
         Integer programming solver object \textsc{Solver}; \\
         Threshold $thres$;
         Horizon $h$; }
  \DontPrintSemicolon
  \SetKwFunction{FChoose}{\textsc{multi\_armed\_bandit}}
  \SetKwProg{Fn}{Function}{:}{}
  \Fn{\FChoose{\textsc{Env}, \ \textsc{Solver}, \ \textsc{Bandit}, \ $thres$, \ $h$}}{

    $done$ \gets \texttt{False}
    \While{not at end of this document}{

      \tcp{Select a cycle or a null token} \\
      $c$ \gets \textsc{choose\_cycle}(\textsc{Env}, $h$, $thres$)  \\

      \tcp{Check if a cycle was indeed chosen} \\
      \eIf{$c$ \texttt{is not} \textsc{NULL}}{
        \tcp{Remove cycle and continue search} \\
        \textsc{Env.remove($c$)}
      }{
      \tcp{Just terminate search}\\
        $done$ \gets \texttt{True}
      }
    }

    \tcp{Return environment with removed cycles} \\
    \KwRet \textsc{Env} \;
  }
	\caption{\textsc{multi\_armed\_bandit} Method}
	\label{alg:mab_method}
\end{algorithm}

 % Algorithm
 \begin{algorithm}[htbp]
 	\SetAlgoLined
 	\KwData{Cycle $c$; Horizon $h$; \\
           Lists pseudo-reward statistics $Avg$, $Std$; \\
           Environment simulator object \textsc{Env}; \\
           Oracle solver object \textsc{OPT}; }
   \DontPrintSemicolon
   \SetKwFunction{FChoose}{\textsc{get\_pseudo\_reward}}
   \SetKwProg{Fn}{Function}{:}{}
   \Fn{\FChoose{\textsc{Env}, \ $c$, \ $Avg$, \ $Std$, \ $h$ }}{
     \tcp{Simulate up to horizon $h$ and find optimal matching} \\
     \textsc{Env.simulate}($h$)  \\
     $r_1$ \gets \textsc{OPT.solve}(\textsc{Env})  \\
     \tcp{Remove cycle $c$, find constrained optimal matching} \\
     \textsc{Env.remove}($c$)  \\
     $r_2$ \gets \textsc{OPT.solve}(\textsc{Env}) \\
     \tcp{Return 1 if rewards are equal, 0 otherwise} \\
     \KwRet $r_1 == r_2$ \;
   }
 	\caption{Function \textsc{get\_pseudo\_reward}}
 	\label{alg:get_pseudo_reward}
 \end{algorithm}

 \begin{algorithm}[htbp]

   \SetAlgoLined
   \DontPrintSemicolon

   \KwData{Horizon $h$; Number of iterations $L$; Threshold $thres$; \\
     Environment simulator object \textsc{Env};
     Multi-armed bandit algorithm object \textsc{MAB};
   }

   \SetKwFunction{FChoose}{\texttt{choose\_cycle}}
   \SetKwProg{Fn}{Function}{:}{}

   \Fn{\FChoose{\textsc{Env},\ \texttt{h}, \ $thres$ }}{
     \tcp{Initialize lists of current pseudo-reward statistics} \\
     $C$ \gets \textsc{Env.get\_available\_cycles()} \\
     $Avg$ \gets \textsc{zeros(length(C))} \\
     $Std$ \gets \textsc{zeros(length(C))} \\

     \tcp{Begin iterations} \\
     \For{$i\gets0$ \KwTo $L$}{
       \tcp{Bandit algorithm chooses next cycle to test given statistics} \\
       $c$ \gets \textsc{MAB.pull($C$, Avg, Std)}  \\
       \tcp{Compute pseudo reward for this cycle and update statistics} \\
       $r$ \gets \textsc{get\_pseudo\_reward}(\textsc{Env},\ $c$,\ $Avg$,\ $Std$,\ $h$) \\
       $Avg$ \gets \textsc{update\_running\_average($Avg$, $r$)} \\
       $Std$ \gets \textsc{update\_running\_std($Std$, $r$)} \\ 
     }  
     \tcp{Bandit algorithm chooses best cycle given statistics} \\
     $c\_best$ \gets \textsc{MAB.choose($C$, Avg, Std)} \\
     \tcp{Return best cycle, unless none of the pseudo-reward averages are above a certain threshold} \\
     \eIf{\textsc{All}($Avg \leq thres$)} {
       \KwRet \text{NULL} \\ 
       }{
       \KwRet $c$ \\
       }
     }
   }
 	\caption{Function \texttt{choose\_cycle}}
 	\label{alg:choose_cycle}
 \end{algorithm}

 \begin{figure}[htbp]
  \centering
  %\hspace{-1cm}
  \includegraphics[width=1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/mab.pdf}
  \caption{\textbf{Multi-armed bandit methods} \\ 
  One period of simulation using multi-armed bandit methods to choose cycles.}
  \label{fig:mab}
  \end{figure}

  \begin{figure}[htbp]
  \centering
  %\hspace*{-0.5cm}
  \includegraphics[width=1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/mab_step.pdf}
  \caption{\textbf{Function} \textsc{get\_pseudo\_reward} \\ Multi-armed bandits evaluate if a cycle should be cleared by checking if there is a high chance that the cycle will be used in the future. Such cycles get a \emph{lower} reward, and are left for later.}
  \label{fig:get_pseudo_reward}
  \end{figure}

\paragraph{What are pseudo-rewards?} As we match and clear out a cycle $c$ today, we forgo the opportunity of using any future cycles involving the nodes in $c$. However, because there may be multiple optimal matchings, sometimes the cycles that become unavailable due to the removal of the nodes in $c$ do not matter, and we can still find a matching of the same cardinality without them. In other words, we pay no price for removing these future cycles.

The pseudo-reward associated with cycle $c$ is a random variable whose expectation is the probability that removing a cycle today will \emph{not} negatively impact the optimal matching size between $t$ and $t+h$. The higher this number, the more confident we are that clearing $c$ out today will not give us trouble in the future. A concrete example of this idea in shown in Figure \ref{fig:pseudo_reward_intuition}.

Pseudo-rewards are a natural way for us to control which patients should be matched today. Pairs that are easy to match will likely belong to many cycles, so by removing them we will be incurring a large cost in terms of future cycles that will become unavailable. But that means that the pseudo-reward associated with cycles that involve them will be lower, making them less attractive. On the other hand, patients that are harder to match will not participate many future exchange, so the price we pay for matching them today is low, and their average pseudo-reward is high.

\begin{landscape}
\begin{figure}[h]
\centering
\includegraphics[width=600]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/pseudo_reward_intuition.pdf}
\caption[Intuition for pseudo-rewards]{\textbf{Intuition for pseudo-rewards} \\
  In the example above, at time $t$ we have three available cycles to choose from: $(1,2)$, $(1,3)$ and $(3,4)$, and we are contemplating choosing cycle $(1,2)$ at $t$. If we  constrain ourselves to do so, several other future cycles become unavailable (in red). Nevertheless, in three out of our four simulations we were able to clear the same number of cycles in the constrained scenario, so the running pseudo-reward average is $\frac{3}{4}$. 
}
\label{fig:pseudo_reward_intuition}
\end{figure}
\end{landscape}

% ---------------------------------------------------------------------------
% Results
% ---------------------------------------------------------------------------

\section{Results}

\subsection{Direct prediction methods}

Table \ref{tab:traditional_ml_classifier} shows the performance results for direct prediction methods as estimators, i.e., how accurately they are able to predict whether a node should be matched or not. We see that while overall accuracy can be relatively high at 70-80\%, precision (defined as the ratio between true positives and both true and false positives) is low even for simpler environments like $ABO$, indicating that the models are over-predicting matchings. Also, augmenting the data with information about the graph has little discernible effect under any of the performance criteria. This suggests that there might be gains from using other algorithms that make better use of information about the compatibility graph.

In order to empirically evaluate the performance of our proposed direct prediction method, we proceeded as follow. First, we simulated data using our environments, and applied Algorithm \ref{alg:direct_prediction} at each step, following the outline on Figure \ref{fig:direct_prediction}. Next, using the \emph{same} random seed, we simulated each environment and solved it again using \textsc{Myopic} in lieu of our method. Finally, we computed the average number of matched patients between periods 250 and 1000 (the first 250 periods were used as burn-in) for each algorithm and compared the two.

This process was repeated several times for: each classifier (logistic regression, gradient tree boosting, random forests), each environment ($ABO$, $RSU$, $OPTN$) and nine entry and death rate combinations. The threshold variable was fixed at $thres = \frac{1}{3}$ everywhere.

The result is shown on Figures \ref{fig:traditional_ml_mdp_algo_grb, fig:traditional_ml_mdp_algo_lr, fig:traditional_ml_mdp_algo_rf}, where we show the average performance ratio computed as above for each combination of environment, algorithm and entry/death rate configuration. \ref{Traditional tables are also available as part of the Supplementary Materials. They will be made available online later.}

Unfortunately, as we predicted in the beginning of the section, the poor predictive performance of the algorithms translates into poor performance as the classifier in our direct prediction method. Random forests, in particular, have the lowest rate of accuracy (Table \ref{tab:traditional_ml_classifier}), and also exhibit the lowest performance in the direct prediction method.

\begin{figure}
\centering
\includegraphics[width=\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/traditional_ml_mdp_algo_grb}
\caption{\textbf{Performance of gradient boosting in direct prediction method} \\
    Ratio of average matched patients against \textsc{Myopic}. Simulations ran over 750 periods (after a 250-period burn-in sequence). Lower row is when data was augmented with information about graph.}
\label{fig:traditional_ml_mdp_algo_grb}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/traditional_ml_mdp_algo_lr}
\caption{\textbf{Performance of logistic regression in direct prediction method} \\
    Similar to gradient boosting shown in Figure \ref{fig:traditional_ml_mdp_algo_grb}, logistic regression is only able to perform better than \textsc{Myopic} when the death rate is large.}
\label{fig:traditional_ml_mdp_algo_lr}
\end{figure}
bb
\begin{figure}
\centering
\includegraphics[width=\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/traditional_ml_mdp_algo_rf}
\caption{\textbf{Performance of random forests in direct prediction method} \\
    Random Forests's poor performance is likely due to its low accuracy, as we show on Table \ref{tab:tradional_ml_classifier}.}
\label{fig:traditional_ml_mdp_algo_rf}
\end{figure}

\subsection{Multi-armed bandit methods}

We evaluated the performance of our \emph{multi-armed bandit} in an analogous manner to the direct prediction method described in the previous section.

The average ratio between our method and \textsc{Myopic} is shown on Figure \ref{fig:mab} and elaborated upon on Tables \ref{tab:ucb1}-\ref{tab:exp3}.\footnote{For reasons of space, in the main paper we present only Table \ref{tab:ucb1}. The remaining tables will be made available in a supplementary only appendix.} The results suggest that, for the entry and death rate combinations we experimented on, the \emph{multi-armed bandit} method uniformly dominates \textsc{Myopic} in terms of average number of matched patients per period.

In sparser environments ($RSU$ and $OPTN$), for particular entry and death rate combinations, our method improve upon \textsc{Myopic} by up to about 4\%. This reflects an earlier observation we did when analyzing Figure \ref{fig:greedy_opt_comparison}: gains from taking dynamic consideration into account are larger in situations of moderate sparsity, because that is where \textsc{Myopic} forces the pool to be too thin, and drives the performance away from optimality.

\begin{figure}
\centering
\hspace*{-1.6cm}
\includegraphics[width=1.1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/mab_ABO.pdf}
\hspace*{-1.6cm}
\includegraphics[width=1.1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/mab_RSU.pdf}
\hspace*{-1.6cm}
\includegraphics[width=1.1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/mab_OPTN.pdf}
\caption{\textbf{Performance multi-armed bandit methods across environments} \\
Average percent improvement by the \emph{multi-armed bandit} method over \textsc{Myopic}. Averages were taken over 750 periods, after a 250-period burn-in sequence. See also Tables \ref{tab:ucb1}-\ref{tab:exp3}. \\
{\scriptsize Note: In $RSU$ row, white entries are missing (they are not zero).}}
\label{fig:mab}
\end{figure}

\section{Extensions and future work} \label{sec:extensions}

Our methods suggest a variety of extensions.

\paragraph{Improving the direct prediction method}

First, the relatively weak performance of our \emph{direct prediction} method could be improved in two ways. First, by improving our ability to use information about graph structure when predicting if a node should be matched or not. This could be done using new statistical methods that are applicable to non-euclidean spaces\citep{shuman2013emerging}. In particular, recent years have seen the emergence of interest in generalizing neural networks to model with structured datasets such as graphs. A promising approach is the one taken by \cite{kipf2016semi}, who build on earlier work by \cite{defferrard2016convolutional} to create a very simple recurrence relation for graph embedding:
\begin{align}
  H^{\ell + 1} = \sigma((A_{t} + I) H^{\ell}W^{\ell}) \qquad \text{with} \quad H^{\ell} = X_{t}
\end{align}

\noindent where $X_{t}$ is the matrix of observable characteristics, $W^{\ell}$ is a matrix of coefficients, $A_{t}$ is the adjacency matrix associated with the kidney exchange pool at $t$, $I$ is a comformable identity matrix, and $\sigma$ is a nonlinear function such as the inverse logistic CDF or ReLU\footnote{$ReLU(x) = \max(x, 0)$}. The authors prove that this can be interpreted as a differentiable version of the Weisfeiler-Lehman algorithm for graph isomorphism tests. For us, it is an especially convenient method because the input graphs need not be the same size throughout.

Second, the \emph{direct prediction} method produces a probability for every pair, but it does not consider correlations between multiple pairs in the same kidney exchange pool. This is unrealistic, as it should be often the case that pairs in the same exchange should have roughly equally high or low probability of being matched. In order to rectify this shortcoming of our approach, we would like to train our statistical estimator using the sequence-to-sequence models of recurrent neural networks studied by \cite{sutskever2014sequence} (see also  \citep{lecun2015deep} for an overview). The idea in those models

. We will be turning our attention to these methods next.

Second, while we see that the performance of our \emph{multi-armed bandit} methods was already satisfactory, we note that it uses only simulations and no data to decide which exchange to clear at each period. In that sense, it worked in a diametrically opposite way to the \emph{direct prediction} method that used only data and no simulations. In fact, an intermediate approach, using \emph{contextual bandits} \citep{lattimore2018bandits} could be used, blending information about data and simulations.

Thirdly, we would like to experiment with different objective functions, since, in reality, different exchanges may have different levels of desirability or priority. For example, it is common for pediatric patients and previous organ donors receive higher priority, as do \emph{ABDR0} exchanges involving ``perfectly matched" patients and donors.

Lastly, we remark that there still remains a moderately large gap between what could be optimally achieved if we could perfectly predict the evolution of the pool. It is conceivable that as better predictive methods arise, this gap will become smaller.

\section{Conclusion}

In this paper, we examined two novel algorithms for dynamic matching in a discrete-time model of kidney exchange: a \emph{direct prediction method} that tries to predict which pair should be matched at each period; and a \emph{multi-armed bandit method}, that uses simulations to score available exchanges in terms of their desirability. We evaluate these methods them using simulations under a variety of different settings, and compared them in terms of average number of matched pairs per period to a \textsc{Myopic} algorithm that finds and immediately clears the maximal matching at each period.

We find that our \emph{multi-armed bandit} method is able to uniformly dominate \textsc{Myopic} in all our simulation settings, including one where we draw pairs from the historical list of patients and donors that have undergone transplants in the United States.

\section*{Disclaimer}

This work was supported in part by Health Resources and Services Administration contract 234-2005-37011C. The content is the responsibility of the authors alone and does not necessarily reflect the views or policies of the Department of Health and Human Services, nor does mention of trade names, commercial products, or organizations imply endorsement by the U.S. Government.

\section{Tables}

\begin{table}[H]
  \singlespacing
  \caption{Blood type probabilities in the ABO environment}
  \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter2/tables/abo_probability.tex}
  \label{tab:abo_blood_type}
\end{table}

\begin{table}
  \centering
  \singlespacing
  \hspace*{-0.5cm}
  \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter2/tables/traditional_ml_classifier.tex}
   \caption{\textbf{Performance of statistical methods as classifiers} \\ 
   We experiment with three variations on the \emph{direct prediction} method by implementing it using different statistical. Here we see their performance as classifiers, that is, how well they are able to predict that the node was chosen to be matched by an offline algorithm. \\
  }
  \label{tab:traditional_ml_classifier}
\end{table}

\begin{table}
  \centering
  \singlespacing
  \hspace*{-0.5cm}
  \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter2/tables/traditional_mdp_grb_none.tex}
   \caption{\textbf{Gradient Boosting (no graph information)} \\ }
  \label{tab:grb_none}
\end{table}

\begin{table}
  \centering
  \singlespacing
  \hspace*{-0.5cm}
  \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter2/tables/traditional_mdp_grb_networkx.tex}
   \caption{\textbf{Gradient Boosting (augmented with graph information)} \\ }
  \label{tab:grb_networkx}
\end{table}

\begin{table}
  \centering
  \singlespacing
  \hspace*{-0.5cm}
  \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter2/tables/traditional_mdp_lr_none.tex}
   \caption{\textbf{Logistic Regression (no graph information)} \\ }
  \label{tab:lr_none}
\end{table}

\begin{table}
  \centering
  \singlespacing
  \hspace*{-0.5cm}
  \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter2/tables/traditional_mdp_lr_networkx.tex}
   \caption{\textbf{Logistic Regression (augmented with graph information)} \\ }
  \label{tab:lr_networkx}
\end{table}

\begin{table}
  \centering
  \singlespacing
  \hspace*{-0.5cm}
  \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter2/tables/traditional_mdp_rf_none.tex}
   \caption{\textbf{Random Forests (no graph information)} \\ }
  \label{tab:rf_none}
\end{table}

\begin{table}
  \centering
  \singlespacing
  \hspace*{-0.5cm}
  \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter2/tables/traditional_mdp_rf_networkx.tex}
   \caption{\textbf{Random Forests (augmented with graph information)} \\ }
  \label{tab:rf_networkx}
\end{table}

\begin{table}
  \centering
  \singlespacing
  \hspace*{-0.5cm}
  \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter2/tables/mab_UCB1.tex}
   \caption{\textbf{Multi-armed bandit algorithm UCB1} \\ 
  \textsc{Environ} is the environment used for simulation. \textsc{Entry} and \textsc{Death} are the Poisson entry rate of entry and the the Geometric rate of departure (times 100), respectively.  \textsc{Mean} and \textsc{Std} refers to the average number of matched patients over 750 periods (after 250 periods of burn-in). \textsc{Difference} and \textsc{Ratio} compare the average improvement between the algorithm and \textsc{Myopic}. \textsc{N} is the number of 1000-period simulations that used that particular configuration. Tables for other bandits are similar, and can be found in the online supplement that will be available online.}
  \label{tab:ucb1}
\end{table}

\begin{table}
  \centering
  \singlespacing
  \hspace*{-0.5cm}
  \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter2/tables/mab_Thompson.tex}
  \caption{\textbf{Multi-armed bandit algorithm Thompson Sampling} \\ 
 \textsc{Environ} is the environment used for simulation. \textsc{Entry} and \textsc{Death} are the Poisson entry rate of entry and the the Geometric rate of departure (times 100), respectively.  \textsc{Mean} and \textsc{Std} refers to the average number of matched patients over 750 periods (after 250 periods of burn-in). \textsc{Difference} and \textsc{Ratio} compare the average improvement between the algorithm and \textsc{Myopic}. \textsc{N} is the number of 1000-period simulations that used that particular configuration.}
  \label{tab:thompson}
\end{table}

### 2018-04-04 ec-18/kidney-ec-submission.tex
etc

@@
 
+\usepackage{textgreek}
+
 \usepackage[ruled]{algorithm2e} % For algorithms
@@
 \usepackage{ textcomp }
+\usepackage[demo]{graphicx}
 
@@
 % Title portion. Note the short title for running heads 
-\title[New strategies for dynamic kidney exchange]{New strategies for dynamic kidney exchange}  
+\title[Novel algorithms for dynamic kidney exchange]{Novel algorithms for dynamic kidney exchange}  
 \author{Submission 42}
@@
 \begin{abstract}
-[[Kidney exchange is great, provides a way to increase number of lives saved]] [[Most algorithms static, don't think about future]] [[There might be ways to improve, but literature still small]] [[This work adds two contributions to the literature]] [[Direct prediction method, MAB-based methods]] [[The latter can improve the average number of kidneys exchange by 3\% over the myopic algorithm]]
+In recent years, kidney paired donation (KPD) has emerged as an alternative for end-stage renal disease patients with incompatible living donors.
+However, we argue that a larger number of patients may be reached if kidney transplant centers take into consideration how their pool of patients and donors will evolve over time.
+To that end, we contribute to the matching literature in dynamic settings by proposing two novel kidney exchange methods, and empirically evaluate these methods using simulations.
+Our objective function is the average number of matched patients per period.
+We restrict our experiments to pairwise kidney exchange, but the methods described here are easily extensible, computational constraints permitting.
+We find that at least one of our methods, based on multi-armed bandit algorithms, is able to outperform the myopic method that is used by kidney transplants in practice, sometimes by as much as 4\%.
 \end{abstract}
@@
 
-[[Kidney exchange]]
+Patients suffering from end-stage renal disease have two available treatments: dialysis and renal transplant.
+While transplant is associated with ``lower mortality rates and improved quality of life compared to chronic dialysis treatment" \citep{tonelli2011systematic}, severe kidney shortage prevents tens of thousands of patients every year from receiving a transplant.
+In the US alone, there are currently over 90,000 patients in the kidney transplant waiting list, and patients often must wait several years before finding an available donor.
 
-[[Will interchangeably use the words 'node', 'vertex' and 'pair']]
+In order to overcome this challenge, in recent years \emph{kidney paired donation} (KPD) has emerged as an alternative option for patients who have an incompatible but otherwise willing living donor.
+The approach involves pooling patient-donor pairs and swapping or exchanging donors so as to increase the overall number of matched patients.
 
-\paragraph{\textbf{Main idea}} We propose and empirically evaluate two novel approaches for increasing the cardinality of matched pairs in a discrete time dynamic model of kidney exchange.
-Our two methods are named the \emph{direction estimation} and \emph{multi-armed-bandit} methods.
-The former recasts the dynamic kidney exchange problem as a binary classification task: for each pair in the pool, we predict a binary label representing whether they should be matched today (one), or left for the future (zero), given their observable characteristics.
-The latter uses simulations to select exchanges that are less likely to decrease the size of an optimal future matching.
-We restrict our analyses to pairwise exchanges (i.e., only two-cycles are allowed), but the ideas introduced here can be easily extended to more complex exchanges, computational constraints permitting.
+Kidney paired donation was proposed about three decades ago by \cite{rapaport1986case}, but its implementation in the United States began only in the early 2000s.
+At the same time, \cite{roth2004kidney} helped elevate KPD as an object of academic study, and since then it has evolved and expanded rapidly both in academic and outside of it.
+\citep{sonmez2013market}.
 
-Each of our two approaches faces their own challenges.
-In the \emph{direct estimation} method, one challenge is that an important part of our data -- the compatibility graph -- lives in a non-Euclidean space, and it is not obvious how to use it as an input to a classification algorithm.
-We attempt to apply a variety of new methods from the graphical signal processing literature to tackle this problem.
-[[[An additional challenge is that traditional classification algorithms are not equipped to produce dependent labels, but in our setting whether or not we decide that a certain pair is to be matched today ought to be related to our decisions about everyone else in the pool.]]]]
+Nevertheless, even today the amount of kidney exchanges remains smaller than its potential, due to several reasons.
+As we will argue in this paper, one source of inefficiency in kidney exchange is that most transplant centers do not take into consideration the evolution of their kidney pool over time and, as we will demonstrate shortly via simulations, this can lead to large losses in terms of the overall number of pairs matched over time.
 
-In the \emph{multi-armed bandit} method, we would like to predict how likely it is that a certain pair should be reserved for a future exchange, and match those less frequently.
-However, computationally costly simulations and the enormous branching factor of our problem precludes us from accurately estimating of this probability.
-In order to overcome this challenge, we employ multi-armed bandit (MAB) algorithms that optimally choose which exchange to collect statistics about under a relatively small computational budget.
+To this end we contribute to the \emph{dynamic} kidney matching literature by introducing two novel algorithms for kidney exchange. They are explained next.
+
+\paragraph{\textbf{Main idea}} We propose two novel approaches for increasing the cardinality of matched pairs in a discrete-time dynamic model of kidney exchange: the \emph{direction prediction} and the \emph{multi-armed-bandit} methods.
+We also empirically evaluated these methods using simulations, under a variety of data-generating processes that we call \emph{environments}.
+
+Let's see how these methods work, and some of the challenges they overcome.
+
+First, our \emph{direct prediction} method recasts the dynamic kidney exchange problem as a binary classification task: for each pair in the pool, we predict a binary label representing whether they should be matched today (one), or left for the future (zero), given their observable characteristics.
+The estimator used here was trained in data extracted from simulations under a variety of setups.
+
+The main challenge with this approach is how to represent the data: while covariates such as ABO bloody type and current waiting time can be naturally represented by real numbers, it is not immediately clear how to represent their relationship to other nodes in the graph.
+We attempt to overcome this issue by augmenting the data set with graph-theoretic notions such as measures of node centrality, size of graph, average degree, and so on.
+
+Our second method, named the \emph{multi-armed bandit} method, uses simulations to answer the question: ``if we take a particular action today, how likely are we to decrease the number of matched pairs between now and a horizon $h$?".
+Naturally, actions for which this probability is large should be left for the future, while actions for which this probability is low should be prioritized.
+The main challenge here is that simulations are so computationally costly, and the number of available actions so vast, that we are cannot realistically produce accurate estimates of this probability for all points in the action space.
+To overcome this issue, we employ multi-armed bandit (MAB) algorithms.
+Their role is to manage a relatively small computational budget and determine which actions are likely attractive (and should be explored further), and which are not (and thus not worth exploring much).
 
@@
 
-We apply our methods to each simulation setup and compare them to a \textsc{myopic} algorithm that clears the maximal matching at each period, and to an infeasible \textsc{OPT} algorithm that is able to observe future states of the world and act optimally.
+We have to benchmarks against which we compare our methods.
+First, a \textsc{myopic} algorithm that clears the maximal matching at each period.
+This benchmarks represents roughly what most organ clearinghouses are doing today.
+Second, in order to know what is the maximal payoff we could have achieve, had we been able to know the future, we also compare to an infeasible offline optimal algorithm \textsc{OPT}.
 Our metric for comparison is the per-period average number of matched pairs over a long period of time.
 
-In all environments, we observe that the \emph{direct estimation} unfortunately does equally well or worse than the myopic algorithm.
-However, the \emph{multi-armed bandit} method is able to improve substantially upon the \textsc{Myopic} benchmark, [[[suggesting that we could save more lives etc.
-]]]]
+In all environments, we observe that our implementation of the \emph{direct estimation} method improve upon \textsc{Myopic}, but only under restricted conditions.
+However, the \emph{multi-armed bandit} method is able to improve uniformly and substantially upon the \textsc{Myopic} benchmark, in particular in environments of moderate sparsity.
 
-\subsection{Related literature} This paper pulls ideas from microeconomics and matching theory, and uses methods drawn from computer science, machine learning, and operations research.
+\subsection{Related literature} This paper pulls ideas from microeconomics, matching theory, and operations research, and uses methods drawn from computer science and machine learning.
 Let us take a brief look at these fields in turn.
@@
 Two of the simulation setups in our work follows a modified, discrete-time version of the models in these seminal papers.
-
-In the computer science literature, \citet{awasthi2009online} and later \citet{dickerson2015futurematch} wrote computational alternatives to simple myopic kidney exchange.
-Our methods are similar to the \emph{weighted myopia} scheme seen in these papers.
 Other papers that influence ours are \citet{akbarpour2017thickness} and \citet{ashlagi2013kidney}.
 
-\paragraph{Computer science and machine learning}
+\paragraph{Computer science and machine learning} We were inspired by several computational alternatives to dynamic kidney exchange that have been proposed in the past few years.
+In particular, the \emph{multi-armed bandit method} described here is reminiscent Algorithm 1 in \citet{awasthi2009online}, inasmuch it also selects the best actions today by repeatedly simulating the future, solving the offline problem, producing a ``score" for each cycle, and selecting actions with maximum score.
+However, our simulation system and ``scoring" differ importantly in their details because we leverage multi-armed bandit algorithms to decrease the computational burden of producing multiple simulations.
+
+\citet{dickerson2012dynamic, dickerson2015futurematch} propose a related method that they call \emph{weighted myopia}.
+The idea is to use simulations for learning ``potentials" of specific elements of the graph.
+Much like the ``scores" in \citet{awasthi2009online}, the set of ``potentials" associated with a specific exchange encodes how desirable each exchange is in terms of its future value.
+The authors then use these pre-learned numbers to revise the weights in a myopic algorithm, forcing it to select more desirable exchanges where it would be otherwise indifferent.
+Our \emph{direct prediction method} works similarly, with three crucial differences.
+First, we do not directly predict how ``desirable" a matching will be, but instead we predict whether or not an offline, optimal algorithm would choose the node or not.
+Second, our method uses a more aggressive thresholding mechanism to select which nodes should be matched today, and which should be left for later.
+Third, we optionally make use of information about how the node fits inside the compatibility graph, so that the information about the node (e.g., the blood type and HLA profile of patient and donor) is augmented with graph-theoretic notions (e.g., measures of node centrality, degree, etc).
 
 %%%%%%%%%
-\section{Simulation environments}
+\section{Environments}
 
-\paragraph{Basic definitions and notation}
+Informally speaking, an \emph{environment} is a simulation setup.
+It includes the rules for which pairs are deemed compatible, the entry and death processes that govern the evolution of the pool, which characteristics are observable to the algorithm at each period, and what utility the agent gets as it matches the pairs.
 
-An \emph{agent} is an algorithm that decides who should be matched.
+In our paper we have three environments inspired by previous works on dynamic kidney exchange. In what follows we will explain how they differ.
 
@@
 \paragraph{Binary preferences} In this work we abstract from these concerns and consider every exchange to be equally desirable, so long as it is available.
-This is the same assumption as used in \cite{roth2005pairwise}.
+This assumption was previously used by \citet{roth2005pairwise}.
 
@@
 The \emph{RSU} environment is named after the simulation model in \citet*{roth2007efficient}.
-Each pair is characterized by patient and donor ABO blood types, current waiting time, and a \emph{calculated panel reactive antibody (cPRA)} level that governs the probability of a crossmatch with a random donor.\footnote{The original \citet{roth2007efficient} paper called this simply \emph{PRA}.
+Each pair is characterized by patient and donor ABO blood types, current waiting time, and a \emph{calculated panel reactive antibody (cPRA)} level that represents the probability of a crossmatch with a random donor.\footnote{The original \citet{roth2007efficient} paper called this simply \emph{PRA}.
 In real life there is an important distinction between the two measures, but for the purposes of a simulation model this distinction is immaterial.
@@
 The STAR dataset contains information from all patients that were ever registered to the kidney waiting list in the United States for the past three decades, as well as from living donors that participated in an transplant.
-From this original dataset, we excluded patients that were registered for more than one organ (including kidney+pancreas), patients that were not waiting for their first kidney transplant, and patients and donors with incomplete HLA profile information.
+From this original dataset, we excluded patients that were registered for more than one organ (including kidney+pancreas), patients that were not waiting for their first kidney transplant, and patients and donors with incomplete tissue-type profile information.
 The resulting dataset contained 117813 patients and 9337 living donors.
@@
 \paragraph{Patient and donor cPRA} We added two additional variables to the original dataset: a \emph{patient cPRA} and a \emph{donor cPRA}.
-While the patient cPRA is a measure of patient tissue-type incompatibility with a random donor \cite{cecka2010calculated}, our \meph{donor cPRA} is a novel measure of the opposite direction -- how frequently a donor is tissue-type incompatible with a random patient.\footnote{We thank Itai Ashlagi for the suggestion of a donor cPRA.} Both were computed empirically: for each patient our dataset, we checked the how many donors exhibited antigens that are unacceptable for the patient in any of the A, B, Bw, C, DR, DPB, DQ, and DQA loci, and assigned this positive crossmatch probability as their patient cPRA; for the donors, we worked in the opposite direction by calculating the frequency of patients who exhibited antibodies against their donor's antigens, and that became their donor cPRA.
+While the patient cPRA is a measure of patient tissue-type incompatibility with a random donor \cite{cecka2010calculated}, our \meph{donor cPRA} is a novel measure of the opposite direction -- how frequently a donor is tissue-type incompatible with a random patient.\footnote{We thank Itai Ashlagi for the suggestion of a donor cPRA.} Both were computed empirically: for each patient our dataset, we checked the how many donors exhibited antigens that are unacceptable for the patient in any of the A, B, Bw, C, DR, DPB, DQ, and DQA loci, and assigned this positive crossmatch probability as their patient cPRA\footnote{A genetic \emph{locus} is a specific position on a chromossome.
+The loci above encodes a donor tissue type.}; for the donors, we worked in the opposite direction by calculating the frequency of patients who exhibited antibodies against their donor's antigens, and that became their donor cPRA.
 Figure \ref{fig:cpra} shows the distribution across the entire population.\footnote{We should remark that we found a very different patient cPRA distribution than the one in the OPTN dataset.
@@
 
-\paragraph{Artificial dataset} Next, we created an artificial dataset by randomly drawing patients and living donors from the historical dataset and checking for blood-type compatibility, and tissue-type compatibility as explained above.
+\paragraph{Artificial dataset} We created an artificial dataset by randomly drawing patients and living donors from the historical dataset and checking for blood-type compatibility, and tissue-type compatibility as explained above.
 Compatible pairs were discarded.
 We iterated in this manner to construct a dataset of about one million incompatible pairs.
-Its columns are a set of dummies representing patient and donor blood times, a set of dummies for each allele in each HLA locus, and
-
-At every simulation period, a random number of pairs are drawn from this dataset.
-
-[[[[Algorhtm can still see original alleles. Patient and donor cPRA offer "long-run" measure of compatibility]]]]
+At every simulation period, a random number of pairs is drawn from this dataset.
 
@@
 
-\begin{figure}[H]
+\begin{figure}[htbp]
 \centering
@@
 
-\subsection{Direct prediction methods}
+\paragraph{Training the classifier} In order to produce data to feed into our ``classifier", we repeatedly created 1000-period simulation runs and solved them offline using \textsc{OPT}.
+Then, for every period $t$ in simulation $k$, we stored: a matrix $X_t^{k}$ whose rows represent each pairs observable characteristics; an additional matrices $E_{t}^{1k}$ containing additional graph-theoretical information such as several centrality measures (betweenness, in-degree, out-degree, harmonic, closeness), in- and out-degrees, and average neighbor degree; a conforming matrix $E_{t}^{2k}$ containing the entry and death rates used in that simulation; a binary vector $y_t^{k}$ indicating which nodes \textsc{OPT} chose to match in period $t$.
+Finally, we vertically concatenated each one of these matrices to create the final data set.
 
-In order to produce data to feed into our ``classifier", we repeatedly created 1000-period simulation runs and solved using them OPT using with each each node's observable characteristics as regressors, and a binary variable indicating whether they were matched or not as the regressand.
-The resulting artificial dataset of approximately 2 million observations was then fed to a series of predictive algorithms, including penalized logistic regression \citep{wu2009genome}, support vector machines with radial basis function kernels \citep{cortes1995support}, random forest classifiers \citep{breiman2001random}, gradient tree boosting classifiers \citep{friedman2001greedy} and deep feedforward networks \citep{goodfellow2016deep}.
-
-Moreover, in order to represent information about the graph and how the nodes fit within it, we augment the original dataset using one or both of the following.
-First, with vertex properties for each node such as the number of in and out-edges, average neighbor degree, some centrality measures (betweenness, in-degree, out-degree, harmonic, closeness) and more exotic indices such as the core number and pagerank of each node\cite{newman2010networks}.
-Second, we augment the data using the \emph{node2vec} graph embedding algorithm of \cite{grover2016node2vec}, a continuous feature representation algorithm inspired by the skip-gram model.[[[More on this?]]].
-These algorithms will produce auxiliary matrices $E_t^{1}$ and $E_t^2$ that are concatenated to the original data regressor matrix $X_t$, and the entire augmented data set can again be fed to the predictive algorithms above.
-Their performance is compared on the bottom half of Table \ref{tab:direct_prediction}.
+We repeated this procedure for each environment, and produced three artificial data sets, each containing approximately 2 million observations.
+These data sets were then fed to a series of predictive algorithms: penalized logistic regression \citep{wu2009genome}, random forest classifiers \citep{breiman2001random}, and gradient tree boosting classifiers \citep{friedman2001greedy}.
 
@@
 
-Our procedure is illustrated in Figures \ref{fig:mab} and \ref{fig:mab_step}.
+Our procedure is illustrated in Figures \ref{fig:mab} and \ref{fig:get_pseudo_reward}.
 At the beginning of the period $t$, the agent receives a set of cycles $C$ that are available to be cleared.
@@
 $$c_{\ell}^{*} \in \arg\max_{c} E[r_{c,\ell}] $$
-
  Also, let \emph{regret} be defined as the difference between expected reward of the optimal choice $c^*$ and its own selected choice $c_{\ell}$.
-
  $$\Delta_\ell := E[r_{c^{*},\ell}] - E[r_{c,\ell}]$$
@@
  A \emph{multi-armed bandit} (MAB) algorithm is a procedure to minimize the cumulative regret over $L$ rounds.
-A good MAB algorithm will act so as to balance exploration (trying out different choices to get high-quality estimates of their rewards) and exploitation (using out better choices more often to increase total rewards).
+A good MAB algorithm will act so as to balance exploration (trying out different choices to get high-quality estimates of their rewards) and exploitation (using out better choices more often to increase total rewards), and produce regret that grows asymptotically logarithmically with the horizon.
+Here, we will be experimenting with three common bandits algorithms, named \emph{UCB1}, \emph{EXP3} and \emph{Thompson sampling}.
 
- The literature on bandit algorithms is extensive and spans at least seventy years of research \citep{lattimore2018bandits}. Here we will focus on binary
+The literature on bandit algorithms is extensive and spans at least seventy years of research, and an in-depth review is outside the score of this paper.
+We refer the reader to \citep{lattimore2018bandits} for a survey.
 
@@
 
-\subsection{Direction prediction methods}
+\subsection{Direct prediction methods}
 
-Table \ref{tab:traditional_ml_classifier} shows the performance results for direct prediction methods as classification methods, i.e., how accurately they are able to predict whether a node should be matched or not.
-Results imply the following.
-
-First, while overall accuracy can be relatively high at 70-80\%, precision (defined as the ratio between true positives and both true and false positives) is low even for simpler environments like $ABO$, indicating that the models are over-predicting matchings.
-Second, augmenting the data with information about the graph has no discernible effect under any of the performance criteria.
+Table \ref{tab:traditional_ml_classifier} shows the performance results for direct prediction methods as estimators, i.e., how accurately they are able to predict whether a node should be matched or not.
+We see that while overall accuracy can be relatively high at 70-80\%, precision (defined as the ratio between true positives and both true and false positives) is low even for simpler environments like $ABO$, indicating that the models are over-predicting matchings.
+Also, augmenting the data with information about the graph has little discernible effect under any of the performance criteria.
 This suggests that there might be gains from using other algorithms that make better use of information about the compatibility graph.
 
-Unfortunately, the algorithms' substandard performance as classifiers translates into poor performance as policy as well.
+In order to empirically evaluate the performance of our proposed direct prediction method, we proceeded as follow.
+First, we simulated data using our environments, and applied Algorithm \ref{alg:direct_prediction} at each step, following the outline on Figure \ref{fig:direct_prediction}.
+Next, using the \emph{same} random seed, we simulated each environment and solved it again using \textsc{Myopic} in lieu of our method.
+Finally, we computed the average number of matched patients between periods 250 and 1000 (the first 250 periods were used as burn-in) for each algorithm and compared the two.
+
+This process was repeated several times for: each classifier (logistic regression, gradient tree boosting, random forests), each environment ($ABO$, $RSU$, $OPTN$) and nine entry and death rate combinations.
+The threshold variable was fixed at $thres = 3$ everywhere.
+
+The result is shown on Figures \ref{fig:traditional_ml_mdp_algo_grb}-\ref{fig:traditional_ml_mdp_algo_rf}, where we show the average performance ratio computed as above for each combination of environment, algorithm and entry/death rate configuration.
+\footnote{Traditional tables are also available as part of the Supplementary Materials.
+They will be made available online later.}
+
+Unfortunately, as we predicted in the beginning of the section, the poor predictive performance of the algorithms translates into poor performance as the classifier in our direct prediction method.
+Random forests, in particular, have the lowest rate of accuracy (Table \ref{tab:traditional_ml_classifier}), and also exhibit the lowest performance in the direct prediction method.
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/traditional_ml_mdp_algo_grb}
+\caption{\textbf{Performance of gradient boosting in direct prediction method} \\
+    Ratio of average matched patients against \textsc{Myopic}.
+Simulations ran over 750 periods (after a 250-period burn-in sequence).
+Lower row is when data was augmented with information about graph.}
+\label{fig:traditional_ml_mdp_algo_grb}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/traditional_ml_mdp_algo_lr}
+\caption{\textbf{Performance of logistic regression in direct prediction method} \\
+    Similar to gradient boosting shown in Figure \ref{fig:traditional_ml_mdp_algo_grb}, logistic regression is only able to perform better than \textsc{Myopic} when the death rate is large.}
+\label{fig:traditional_ml_mdp_algo_lr}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/traditional_ml_mdp_algo_rf}
+\caption{\textbf{Performance of random forests in direct prediction method} \\
+    Random Forests's poor performance is likely due to its low accuracy (shown on table in online supplement)}
+\label{fig:traditional_ml_mdp_algo_rf}
+\end{figure}
 
@@
 
+We evaluated the performance of our \emph{multi-armed bandit} in an analogous manner to the direct prediction method described in the previous section, except that we used Algorithms \ref{alg:choose_cycle} and \ref{alg:get_pseudo_reward} when selecting nodes to be matched at each period.
+
+The average ratio between our method and \textsc{Myopic} is shown on Figure \ref{fig:mab} and elaborated upon on Table \ref{tab:ucb1} in the appendix.\footnote{For reasons of space, in the main paper we present only Table \ref{tab:ucb1}.
+The remaining tables will be made available in a supplementary online appendix.} The results suggest that, for the entry and death rate combinations we experimented on, the \emph{multi-armed bandit} method uniformly dominates \textsc{Myopic} in terms of average number of matched patients per period.
+
+In sparser environments ($RSU$ and $OPTN$), for particular entry and death rate combinations, our method improve upon \textsc{Myopic} by up to about 4\%.
+This reflects an earlier observation we did when analyzing Figure \ref{fig:greedy_opt_comparison}: gains from taking dynamic consideration into account are larger in situations of moderate sparsity, because that is where \textsc{Myopic} forces the pool to be too thin, and drives the performance away from optimality.
+
+\begin{figure}
+\centering
+\hspace*{-1.6cm}
+\includegraphics[width=1.1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/mab_ABO.pdf}
+\hspace*{-1.6cm}
+\includegraphics[width=1.1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/mab_RSU.pdf}
+\hspace*{-1.6cm}
+\includegraphics[width=1.1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/mab_OPTN.pdf}
+\caption{\textbf{Performance multi-armed bandit methods across environments} \\
+Average percent improvement by the \emph{multi-armed bandit} method over \textsc{Myopic}.
+Averages were taken over 750 periods, after a 250-period burn-in sequence.
+See also Table \ref{tab:ucb1} and additional tables in online supplement.
+\\
+{\scriptsize Note: In $RSU$ row, white entries are missing (they are not zero).}}
+\label{fig:mab}
+\end{figure}
+
 \section{Conclusion}
 
-\paragraph{Summary}
+In this paper, we examined two novel algorithms for dynamic matching in a discrete-time model of kidney exchange: a \emph{direct prediction method} that tries to predict which pair should be matched at each period; and a \emph{multi-armed bandit method}, that uses simulations to score available exchanges in terms of their desirability.
+We evaluate these methods them using simulations under a variety of different settings, and compared them in terms of average number of matched pairs per period to a \textsc{Myopic} algorithm that finds and immediately clears the maximal matching at each period.
 
-\paragraph{Extensions}
-Another source of error is the fact that we are predicting a label for each node independently from the labels we have predicted for other nodes.
+We find that our \emph{multi-armed bandit} method is able to uniformly dominate \textsc{Myopic} in all our simulation settings, including one where we draw pairs from the historical list of patients and donors that have undergone transplants in the United States.
 
- In reality, some exchanges are more or less desirable than others, either for ethical concerns or because of predicted health benefit.
-For example, it is common for pediatric patients and previous organ donors receive higher priority, as do exchanges involving patients with no HLA mismatch.
+Our methods suggest a variety of extensions.
+First, the relatively weak performance of our \emph{direct prediction} method could be improved in two ways: by improving our ability to use information about graph structure when predicting if a node should be matched or not; and, by making sure that what we predict for one pair should be dependent on what we predict for the other pairs in the same graph.
+The former could be done using new statistical methods on non-euclidean spaces \citep{shuman2013emerging}, and the latter using algorithms that output joint distributions of arbitrary length such as recurrent neural networks \citep{lecun2015deep}.
+We will be turning our attention to these methods next.
 
-\section{Conclusions}
+Second, while we see that the performance of our \emph{multi-armed bandit} methods was already satisfactory, we note that it uses only simulations and no data to decide which exchange to clear at each period.
+In that sense, it worked in a diametrically opposite way to the \emph{direct prediction} method that used only data and no simulations.
+In fact, an intermediate approach, using \emph{contextual bandits} \citep{lattimore2018bandits} could be used, blending information about data and simulations.
 
-% Appendix
-\appendix
-\section{Sections}
+Thirdly, we would like to experiment with different objective functions, since, in reality, different exchanges may have different levels of desirability or priority.
+For example, it is common for pediatric patients and previous organ donors receive higher priority, as do \emph{ABDR0} exchanges involving ``perfectly matched" patients and donors.
 
-\section{Supplementary materials}
-
-\begin{table}
-  \centering
-  \singlespacing
-  \hspace*{-1cm}
-  \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/tables/traditional_ml_classifier.tex}
-   \caption{\textbf{Traditional methods} \\ Performance of traditional econometric and classification methods when predicting whether a node will be matched (1) or not (0).
-\emph{Additional} refers to extra information about graph structure passed to the classifier: \emph{Graph Stats} are graph-theoretical measures of node characteristics such as centrality; \emph{Embedding} is a 10-dimensional real embedding of the node provided by node2vec \citep{grover2016node2vec}.
-Recall is true positive rate $\frac{tp}{tp+fn}$; Specificity is true negative rate $\frac{tp}{tn+fp}$; Precision is $\frac{tp}{tp+fp}$; Accuracy is $\frac{tp+tn}{tp+tn+fp+fn}$.
-See section \ref{sec:[[[]]]} for methodology and section \ref{sec:[[[]]]} for discussion of results.
-}
-  \label{tab:tradional_ml_classifier}
-\end{table}
+Lastly, we remark that there still remains a moderately large gap between what could be optimally achieved if we could perfectly predict the evolution of the pool.
+It is conceivable that as better predictive methods arise, this gap will become smaller.
 
@@
 
-The author would like to thank Itai Ashlagi for the idea of including a donor cPRA in the OPTN environment.
+The author would like to thank Itai Ashlagi for the idea of a donor cPRA.
 
@@
 
+\appendix
+
+\section{Tables}
+
+\begin{table}
+  \centering
+  \singlespacing
+  \hspace*{-0.5cm}
+  \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/tables/traditional_ml_classifier.tex}
+   \caption{\textbf{Performance of statistical methods as classifiers} \\ 
+   We experiment with three variations on the \emph{direct prediction} method by implementing it using different statistical.
+Here we see their performance as classifiers, that is, how well they are able to predict that the node was chosen to be matched by an offline algorithm.
+\\
+  }
+  \label{tab:traditional_ml_classifier}
+\end{table}
+
+\begin{table}
+  \centering
+  \singlespacing
+  \hspace*{-0.5cm}
+  \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/tables/mab_UCB1.tex}
+   \caption{\textbf{Multi-armed bandit algorithm UCB1} \\ 
+  \textsc{Environ} is the environment used for simulation.
+\textsc{Entry} and \textsc{Death} are the Poisson entry rate of entry and the the Geometric rate of departure (times 100), respectively.
+\textsc{Mean} and \textsc{Std} refers to the average number of matched patients over 750 periods (after 250 periods of burn-in).
+\textsc{Difference} and \textsc{Ratio} compare the average improvement between the algorithm and \textsc{Myopic}.
+\textsc{N} is the number of 1000-period simulations that used that particular configuration.
+Tables for other bandits are similar, and can be found in the online supplement that will be available online.}
+  \label{tab:ucb1}
+\end{table}
+
+% 
+% 
+% \section{Supplementary materials}
+% 
+% 
+% 
+% \begin{table}
+%   \centering
+%   \singlespacing
+%   \hspace*{-0.5cm}
+%   \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/tables/mab_Thompson.tex}
+%   \caption{\textbf{Multi-armed bandit algorithm Thompson Sampling} \\ 
+%  \textsc{Environ} is the environment used for simulation.
+\textsc{Entry} and \textsc{Death} are the Poisson entry rate of entry and the the Geometric rate of departure (times 100), respectively.
+\textsc{Mean} and \textsc{Std} refers to the average number of matched patients over 750 periods (after 250 periods of burn-in).
+\textsc{Difference} and \textsc{Ratio} compare the average improvement between the algorithm and \textsc{Myopic}.
+\textsc{N} is the number of 1000-period simulations that used that particular configuration.}
+%   \label{tab:thompson}
+% \end{table}
+% 
+% \begin{table}
+%   \centering
+%   \singlespacing
+%   \hspace*{-0.5cm}
+%   \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/tables/mab_EXP3.tex}
+%   \caption{\textbf{Multi-armed bandit algorithm EXP3} \\ 
+%  \textsc{Environ} is the environment used for simulation.
+\textsc{Entry} and \textsc{Death} are the Poisson entry rate of entry and the the Geometric rate of departure (times 100), respectively.
+\textsc{Mean} and \textsc{Std} refers to the average number of matched patients over 750 periods (after 250 periods of burn-in).
+\textsc{Difference} and \textsc{Ratio} compare the average improvement between the algorithm and \textsc{Myopic}.
+\textsc{N} is the number of 1000-period simulations that used that particular configuration.}
+%   \label{tab:exp3}
+% \end{table}
+% 
+% 
+% 
+% 
+% \begin{table}
+%   \centering
+%   \singlespacing
+%   \hspace*{-0.5cm}
+%   \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/tables/traditional_mdp_grb_none.tex}
+%    \caption{\textbf{Gradient Boosting (no graph information)} \\ }
+%   \label{tab:grb_none}
+% \end{table}
+% 
+% 
+% \begin{table}
+%   \centering
+%   \singlespacing
+%   \hspace*{-0.5cm}
+%   \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/tables/traditional_mdp_grb_networkx.tex}
+%    \caption{\textbf{Gradient Boosting (augmented with graph information)} \\ }
+%   \label{tab:grb_networkx}
+% \end{table}
+% 
+% 
+% \begin{table}
+%   \centering
+%   \singlespacing
+%   \hspace*{-0.5cm}
+%   \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/tables/traditional_mdp_lr_none.tex}
+%    \caption{\textbf{Logistic Regression (no graph information)} \\ }
+%   \label{tab:lr_none}
+% \end{table}
+% 
+% 
+% \begin{table}
+%   \centering
+%   \singlespacing
+%   \hspace*{-0.5cm}
+%   \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/tables/traditional_mdp_lr_networkx.tex}
+%    \caption{\textbf{Logistic Regression (augmented with graph information)} \\ }
+%   \label{tab:lr_networkx}
+% \end{table}
+% 
+% 
+% \begin{table}
+%   \centering
+%   \singlespacing
+%   \hspace*{-0.5cm}
+%   \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/tables/traditional_mdp_rf_none.tex}
+%    \caption{\textbf{Random Forests (no graph information)} \\ }
+%   \label{tab:rf_none}
+% \end{table}
+% 
+% 
+% \begin{table}
+%   \centering
+%   \singlespacing
+%   \hspace*{-0.5cm}
+%   \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/tables/traditional_mdp_rf_networkx.tex}
+%    \caption{\textbf{Random Forests (augmented with graph information)} \\ }
+%   \label{tab:rf_networkx}
+% \end{table}
+%
+
 \end{document}

### 2018-04-04 legacy/paper.tex
etc

--- 2018-02-11 paper.tex
@@
 \usepackage[font=footnotesize]{caption}
-\usepackage{lineno}
 
@@
 
-\linenumbers
-
 \section{Introduction} \label{sec:introduction}
 
-\paragraph{Main idea} We study new dynamic kidney matching algorithm that takes as input a pool of patient-donor pairs and computes, for each cycle that is currently available for matching, a probability that it should be chosen to be cleared today, as well as another probability indicating whether we should not match any more cycles.
+\paragraph{Main idea} We study a new dynamic kidney matching algorithm that takes as input a pool of patient-donor pairs and computes, for each cycle that is currently available for matching, a probability that it should be chosen to be cleared today, as well as another probability indicating whether we should not match any more cycles.
 In order to get these probabilities, our algorithm first produces a prior distribution estimated by observing past choices made by an infeasible optimal algorithm, and then updates these priors with simulations.
@@
 
+\paragraph{What are pseudo-rewards?} Figure
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=1.1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/pseudo_reward_intuition.pdf}
+\caption{Pseudo-reward intuition}
+\label{fig:pseudo_reward_intuition}
+\end{figure}
+
 \section{Results} \label{sec:results}
@@
 \section{Conclusion} \label{sec:conclusion}
-
-\section{Acknowledgements}
-
-This work was supported in part by Health Resources and Services Administration contract 234-2005-37011C.
-The content is the responsibility of the authors alone and does not necessarily reflect the views or policies of the Department of Health and Human Services, nor does mention of trade names, commercial products, or organizations imply endorsement by the U.S.
-Government.
 

### 2018-04-04 Untitled.r

@@ -1,26 +1,0 @@
-
-# Helper functions
-make_heterogeneity_table <- function(data, covariates, group, prefix="Group") {
-
-  # Compute mean mean and std. error
-  grp <- data %>% dplyr::group_by_(group)
-  se <- grp %>% dplyr::summarise_all(~sqrt(var(.)/length(.)))
-  m <- grp %>% dplyr::summarise_all(mean) 
-  
-  # Choose covariate order based on Vaar(E[Y|X])/E[Y|X]
-  cov_order <- data %>%
-               select(covariates, group) %>% 
-               group_by_(group) %>%
-               summarize_all(mean) %>%
-               summarize_all(~var(.)/(mean(.) + 1)) %>%
-               select(covariates) %>%
-               as.numeric %>% order() %>% rev()
-
-  # Make a table of characters
-  tab <- map2_df(m, se, ~str_c(round(.x,2), " (", round(.y,2), ") "))  %>% 
-                  select(covariates[cov_order]) %>%
-                  as.data.frame() %>% t()
-  colnames(tab) <- map(seq(dim(tab)[2]), ~str_c(prefix, .))
-  
-  return(tab)
-}

### 2018-04-06 chapter1/chapter1.tex
82 pages

@@
 \chapter{Heterogenous Production Functions, Panel Data, and Productivity Dispersion}
+\begin{center}
+\large
+\textsc{Jeremy Fox, Vitor Hadad, Stefan Hoderlein} \\ \textsc{Amil Petrin, Robert Sherman}
+\end{center}
 
@@
 
-[Primary motivations of panel data: controlling for unobserved heterogeneity]
+% [One or two sentences providing basic introduction to the field, comprehensible to any economist]
+Econometricians often need control for unobserved heterogeneity,
 
-[Correlated random coefficients]
+%[Two to three sentences of a more detailed background, comprehensible to econometricians]
 
-[]
-Random coefficients
- Consumer panels with heterogeneous consumers
- Production function parameters (A,B) vary across firms in
-same industry
+%[One sentence clearly stating the general problem being addressed by this particular study]
+In the context of panel data, it has become standard to controlling for unobserved heterogeneity via a \emph{fixed effects} approach, where we do not impose restrictions the distribution of latent heterogeneity parameters conditional on observable characteristics \citet{wooldridge2010econometric}.
+However, this flexibility often comes at a cost in terms of statistical modeling: the most ubiquitous fixed effects strategy in panel data is via the introduction of a time-invariant random intercept.[\ref{}, \ref{}, \ref{}], but there are many instances of economic problems where this is might not be appropriate (e.g.
+[QUICK EXAMPLE]).
 
-Random coefficients (A,B) can be correlated with inputs (X1,X2)
-  Estimator allows arbitrary correlations of production function parameters (A1,B1) with inputs (X1,X2)
-  Timing assumptions with random coefficients (A,B)
-  Input decisions made in period t − 1 with knowledge of t − 1
-production function
-  Estimation, consistency: censoring for unconditional means
-  India: log TFP dispersion increases in random coefficients vs OLS
+%[One sentence summarizing main result]
+Here, we present constructive identification and estimation results for the moments and marginal distribution of random coefficients in a linear panel data model with \emph{two} random coefficients
+\begin{align}
+  Y_{it} = A_{it} + X_{it}B_{it} + Z_{it}^{T}\beta_{t}
+\end{align}
+
+\noindent where the intercept $A_{it}$ and slope $B_{it}$ follow a specific autoregressive process but are otherwise allowed to be arbitrarily correlated with all regressors, and $Z_{it}$ is a (potentially large) vector of covariates associated with individual-invariant coefficients $\beta$.
+
+% [Two or three sentences explaining what the main results reveals in direct comparison to what was thought to be case previously, or how the main result adds to previous knowledge]
+% [One sentence to put the results into more general context]
+Our results develop upon and extend the results of earlier work by \cite{graham2012identification}, who only focus on the first moments of the random coefficients and have stricter timing assumptions than ours.
+Another paper is \cite{arellano2011identifying}, who, like us, provide a method for identification and estimation of higher moments and distributional characteristics of random coefficients, but they maintain certain regularity assumptions that we do not impose.
+Our model also permits that lagged dependent variables to be regressors.
+More broadly, this paper contributes to the larger literature of panel data models with unobserved heterogeneity, as we discuss below.
+
+% [Two to three sentences to provide a broader perspective, readily comprehensible to a scientist in any discipline..]
+In our particular application, we take $Y_{it}$ to be value-added log-production, while the two random coefficients represent total factor productivity (TFP) and the elasticity of one of the inputs in a Cobb-Douglas model of production.
+There is strong evidence for heterogeneity in this setting, where it is known as \emph{productivity dispersion}.
+In the United States, \cite{syverson2004product} finds large plant-level discrepancies in TFP, noting that within four-digit SIC industries ``the plant at the 90th percentile of the productivity distribution makes almost twice as much output with the same measured inputs as the 10th percentile plant", and a later survey by \cite{syverson2011determines} confirms this to be a typical finding.
+In fact, other work by \citep{hsieh2009misallocation} indicates that this gap may be small compared to other countries.
+In our work we note, however, that the large observed variation in TFP might be a product of restrictive statistical models: when only the intercept is able to vary, it may end up absorbing most the heterogeneity in the data, leading to inflated estimates of the standard deviation.
+On the other hand, we remark that our specific model and data differ from the ones above so we cannot directly comparison our results, but leave it as a suggestion for future research.
+
+% ----------------------
+% LITERATURE REVIEW 
+% ----------------------
 
@@
 
-In its most general form, the basic linear panel model used in econometrics can be described as
-
+In full generality, panel data models can be represented as
 \begin{align}
-  Y_{it} = A_{it} + \sum_{k=1}^{K} X_{it}^{(k)}B_{it}^{(k)}
+  Y_{it} = m(H_{it}, X_{it}, U_{it}) \label{eq:full_generality_panel}
 \end{align}
 
-\noindent where $Y_{it}, X_{it}, \{A_{it}, B_{it}\}_{k=1}^{K}$ are random variables associated with individual $i$ at period $t$.
-The parameters $B_{it}^{(k)}$ represent the causal effect of an increase in their associated regressor, while $A_{it}$ is a disturbance term whose interpretation depends on the specific application.
-When they have a nondegerate probability distribution, they receive various names in the literature, such as ``\emph{unobserved component}, \emph{latent variable}, or \emph{unobserved heterogeneity}"\citep[~p.251]{wooldridge2010econometric}.
+\noindent where $Y_{it}$ is the dependent variable of interest for unit $i$ at period $t$, $H_{it}$ is an individual-specific, possibly infinite-dimensional object often called the \emph{unobserved heterogeneity} component, $X_{it}$ are observable characteristics, and $U_{it}$ is an idiosyncratic error term.
+For example, $Y_{it}$ might represent an individual's wages over time, $X_{it}$ a vector of characteristics that includes educational attainment, $H_{it}$ a measure of their cognitive or manual ability, and $U_{it}$ measurement error.
+The econometrician is usually interested in some functional of this object, such as the causal average effect of on $Y_{t}$ of a change in $X_{it}$ (known as the \emph{average partial effect} of education on wages).
 
- but here we call them \emph{random coefficients}.
+Particular care must be taken with the \emph{unobserved heterogeneity} parameter $H_{it}$.
+\citet[~p.2]{browning2007heterogeneity} define \emph{heterogeneity} as ``the dispersion in factors that are relevant and known to individual agents when making a particular decision"\footnote{Browning \textit{et al} were, in turn, paraphrasing \cite{cunha2005separating}}.
+The qualifier \emph{unobserved} serves to emphasize that the econometrician does not have access to information about such factors.
+This is problematic because, even when $H_{it}$ itself is not of interest, it may preclude the identification of causal objects To continue our running example: even if we observe that workers with higher educational attainment receive higher wages, we cannot a priori claim that we have found a causal effect of $X_{it}$, as it may be that high-ability ($H_{it}$) workers are able to self-select into higher-paying jobs.
 
-Without further restrictions, nothing can be said about the objects $A_{it}$ and $B_{it}$, or even its moments.
-In the linear panel data literature, common identifying assumptions include the following three.
+One of the foremost advantages of dealing with panel data is that, under specific assumptions, we are able to control for the unobserved heterogeneity component.
+However, without further assumptions, however, nothing can be said.
+Therefore, econometricians have posited a number of restrictions on the model above, each producing different identification strategies.
 
-\begin{enumerate}
-  \item $A_{it} = a + \epsilon_{it}$ and $B_{it} = b$, where $a, b$ are constants and $E[\epsilon_{it}|X_{it}] = 0 \forall t$.
-  \item $A_{it} = A_{i} + \epsilon_{it}$ and $B_{it} = b$, where $b$ is a constant and $E[\epsilon_{it}|X_{is}] = 0 \forall t, s$.
-  \item $A_{it} = A_{i} + \epsilon_{it}$ and $B_{it} = b$, where $b$ is a constant and $E[\epsilon_{it}|X_{is}] \neq 0 \forall t, s$.
-\end{enumerate}
+Econometric applications (such as \cite{???}, \cite{???} often to cite a few) assume that $m$ is a linear function.
+\begin{align}
+  Y_{it} = A_{it} + X_{it}^{T}B_{it} + U_{it}
+\end{align}
 
-The first one is the standard linear model that can be consistently estimated by simply running a standard linear regression of $Y_{it}$ on $X_{it}$.
-The second and third are called \emph{random} and \emph{fixed} panel data models, and have received a great deal of attention in the literature (see [???] for a review).
+\noindent where $H_{it} = [A_{it}, B_{it}]$, a vector of \emph{random coefficients}. Furthermore, they will often impose additionally that 
+\begin{align}
+  Y_{it} = A_{i} + X_{it}^{T}\beta + U_{it}
+\end{align}
+
+and that unobserved heterogeneity appears only as an intercept.
+
+% One of the main objectives is to model heterogeneity
+
+% Importance of controlling for heterogeneity
+
+Imposing different restrictions will give rise to different identification strategies.
+As recounted by \cite{graham2012identification}, the \emph{fixed effects} method implies restrictions on the functional form $m$ and on the conditional distribution of the idiosyncratic shocks $F(U_{it}|X_{\cdot}, A_{\cdot})$, but leaves the conditional distribution of unobserved heterogeneity $F(A|X)$ unrestricted.
+A related method, named \emph{random effects}, entails imposing an independence assumption between $A_{\cdot}$ and $X_{\cdot}$, but that usually leads to models that are harder to motivate economically.
+
+Further restrictions can be imposed for simplicity.
+
 Figure \ref{fig:panel_data_models} illustrates the difference between them.
+
+[[[These APEs are essentially derivatives of the average structural function of Blundell and Powell (2004).]]]
+
+In this paper we focus on the more general case where we allow for more than one random coefficient, and also allow them to be correlated with the regressor.
+The literature dealing with this more general case is substantially smaller.
+\citet{hoderlein2010analyzing} analyzed a for any number of random coefficients, but required the stronger assumption that the coefficients are independent from the regressors.
+
+% Closest works are Chamberlain (1982), Arellano and Bonhomme (2011), Graham nd Powell (2012), Evdokimov (2011)]
+
+% Production function estimation
+> Recent papers on heterogeneous production functions Kasahara, Shrimpf and Suzuki (2016), Balat, Brambilla and Sazaki (2016)
+  Extend proxy variable approaches but require variables to be chosen in period t − 1 with only static considerations, as in FOCs. Ackerberg and Hahn (2016)
+  Scalar unobservable enters into nonparametric production function
 
@@
 
-In this paper we focus on the more general case where we allow for more than one random coefficient, and also allow them to be correlated with the regressor.
-The literature dealing with this more general case is substantially smaller.
-\citet{hoderlein2010analyzing} analyzed a for any number of random coefficients, but required the stronger assumption that the coefficients are independent from the regressors.
-
-> [Chamberlain (1982), Arellano and Bonhomme (2011), Graham
-and Powell (2012), Evdokimov (2011)]
-
-> Recent papers on heterogeneous production functions Kasahara, Shrimpf and Suzuki (2016), Balat, Brambilla and Sazaki (2016)
-  Extend proxy variable approaches but require variables to be chosen in period t − 1 with only static considerations, as in FOCs. Ackerberg and Hahn (2016)
-  Scalar unobservable enters into nonparametric production function
-
-  Cobb-Douglas
-  Yi,t =Ai,t +BKKi,t +BLLi,t
-    Productivity dispersion: standard deviation of Ai,t across plants or firms in same industry
-    Syverson (2011) surveys empirical findings on dispersion
-    Typical finding: some plants produce more than twice as much
-  output for same inputs
-    Assuming  BK,BL  is the same across plants in a 4- or 2-digit industry classification may be too restrictive
-    If BK,BL  varybyplanti andtimet andabove i,t i,t
-  specification is estimated, much of this heterogeneity enters into Ai,t
-
-  Estimate BK as expenditure share on capital input i,t
-  Based on static profit maximizing FOC for Cobb-Douglas
-  Treats input as having no adjustment costs, opposite of our
-approach
-  Based on strong conduct assumption: static profit maximization
-  We do not impose profit maximization, important for studying unproductive firms
-  We allow unobservables outside production function to affect input choice
-  Firm-specific input prices
-  Adjustment costs
-  Product demand
-
 \section{Model and assumptions}
@@
 \end{align*}
-%$Z_{t}$ (a $k \times 1$ vector)
-%and $\beta_t$  is a $k \times 1$ vector of fixed coefficients, 
 where $X_{t}$ is a scalar, and $A_{t}, B_{t}$ are scalar random coefficients that evolve according to an AR(1) Markov process.
@@
 \end{align}
-% 
-% Let $r$ be a positive integer and for $t \geq 2$ define $W_t = (X_{t-1}, X_{t})$. We will say that $(U_t, V_t)$ are \emph{r\textsuperscript{th}-moment independent} of $W_t$ if 
-%   \begin{align}
-%     \text{For }1 \leq j \leq r,
-%     \quad     
-%     E[U_{t}^{r-j} V_{t}^{j} | W_t] = E[U_{t}^{r-j} V_{t}^{j}]
-%   \end{align}
-% \end{definition}
 
@@
   Shocks are sequentially independent from all covariates
-  % \begin{align}
-  %   \text{For }1 \leq j \leq r,
-  %   \quad     
-  %   E[U_{t}^{r-j} V_{t}^{j} | W_t] = E[U_{t}^{r-j} V_{t}^{j}]
-  % \end{align}
   \begin{align}
@@
   \begin{cases}
-    %Y_{1} = A_{1} + B_{1}X_{1} + Z_1^{T}\beta_1 \\ 
-    %Y_{2} = A_{2} + B_{2}X_{2} + Z_2^{T}\beta_2
     Y_{1} = A_{1} + B_{1}X_{1} \\ 
@@
   \begin{cases}
-    %Y_{1} = A_{1} + B_{1}X_{1}  + Z_1^{T}\beta_1 \\ 
-    %Y_{2} = A_{1} + U_{2} + B_{1}X_{2} + V_{2}X_{t} + Z_2^{T}\beta_2 \label{eq:expanded}
     Y_{1} = A_{1} + B_{1}X_{1} \\ 
@@
 \begin{align}
-  %E[Y_{2} - Y_{1} \ | \  W_{2} = w_{2}] = 
-  %E[U_{2}] + E[V_{2}]x_2 + z_2^{T}\beta_2 - z_1^{T}\beta_{1}
   E[Y_{2} - Y_{1} \ | \  W_{2} = w_{2}] = 
@@
 \begin{align}
-%\min_{a, b, \beta_1, \beta_2}
   \min_{\theta_u, \theta_v}  
-  %E[(Y_{2} - Y_{1} - a - bx_2 - z_2^{T}\beta_2 + z_1^{T}\beta_{1})^2 | W_2 = w_2]
   E[(Y_{2} - Y_{1} - \theta_u -  x\theta_v)^2 \  | \ W_2 = w_2]
@@
   \begin{cases}
-    % E[Y_{1}|W_{2}]
-    %   = E[A_{1}|W_{2}] + E[B_{1}|W_{2}]x_{1} + z_1^{T}\beta_1 \\ 
-    % E[Y_{2}|W_{2}]
-    %   = E[A_{1}|W_{2}] + E[U_{2}] + E[B_{1}|W_{2}]x_{2} + E[V_{2}]x_{t} + z_2^{T}\beta_2
       E[Y_{1}|W_{2}]
@@
   \begin{bmatrix}
-   %  E[Y_{1}|W_{2}] - z_1^{T}\beta_1 \\  
-  %  E[Y_{2}|W_{2}] - E[U_{2}] -  E[V_{2}]x_{t} - z_2^{T}\beta_1
     E[Y_{1}|W_{2}] \\  
@@
   \begin{bmatrix}
-    % E[Y_{1}|W_{2}]  - z_1^{T}\beta_1  \\  
-    % E[Y_{2}|W_{2}] - E[U_{2}] -  E[V_{2}]x_{t} - z_2^{T}\beta_1
     E[Y_{1}|W_{2}] \\  
@@
 \begin{align}
-  %E[Y_{2} - Y_{1} \ | \  W_{2} = w_{2}] = 
-  %E[U_{2}] + E[V_{2}]x_2 + z_2^{T}\beta_2 - z_1^{T}\beta_{1}
   E[(Y_{2} - Y_{1})^2 \ | \  W_{2} = w_{2}] = 
@@
 \begin{align}
-%\min_{a, b, \beta_1, \beta_2}
 E[U_2^2], E[V_2^2], E[U_2V_2] =
   \min_{\theta_{uu}, \theta_{vv}, \theta_{uv}}  
-  %E[(Y_{2} - Y_{1} - a - bx_2 - z_2^{T}\beta_2 + z_1^{T}\beta_{1})^2 | W_2 = w_2]
   E[(Y_{2} - Y_{1} - \theta_{uu} -  x^2\theta_{vv} - 2x\theta_{uv})^2 \  | \ W_2 = w_2]
@@
 
-[[Interpretation]]
+% Interpretation
+
+;;Comparison to static FOCS:
+;; Estimate BK as expenditure share on capital input i,t
+;; Based on static profit maximizing FOC for Cobb-Douglas
+;;Treats input as having no adjustment costs, opposite of our
+approach
+;;; Based on strong conduct assumption: static profit maximization
+
+We do not impose profit maximization, important for studying unproductive firms
+We allow unobservables outside production function to affect input choice
+Firm-specific input prices
+Adjustment costs
+Product demand
 
@@
   \label{fig:simulation_shocks}
+  \caption[Shock estimates in simulated data]{Shock estimates in simulated data}
 \end{figure}
@@
   \label{fig:simulation_random_coefficients}
+  \caption[Random coefficient estimates in simulated data]{Random coefficient estimates in simulated data}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[height=0.8\textheight]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter1/figures/shocks_empirical.pdf}
+  \caption[Shock estimates using ASI data]{Shock estimates using ASI data}
 \end{figure}
@@
   \includegraphics[height=0.8\textheight]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter1/figures/rc1_empirical.pdf}
+  \caption[Random coefficient estimates using ASI data (2008)]{Random coefficient estimates using ASI data (2008)}
 \end{figure}
@@
   \includegraphics[height=0.8\textheight]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter1/figures/rc2_empirical.pdf}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[height=0.8\textheight]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter1/figures/shocks_empirical.pdf}
+    \caption[Random coefficient estimates using ASI data (2009)]{Random coefficient estimates using ASI data (2009)}
 \end{figure}
@@
 \clearpage
+
+\section{Conclusion and future work}
+
+What if T > 2? The approach generalizes but seems to become more complicated (as for example U2,V2 are correlated with X3,X4,...).
+• It would be useful to see how strict the timing assumptions are (in standard linear panel models different timing assumptions can be dealt with by changing the set of instruments).
+For example: what if U2 independent of X1 but not of X2,X3,...?
+• The authors’ setup could be used to learn about the firms’ ob- jectives, by estimating the determinants of input choice: Xit given
+Ai,t−1, Bi,t−1, Xi,t−1, ....
+• Looking ahead: nonlinear models? Example: CES.
+
+% ------------
+% APPENDIX
+% ----------
 

### 2018-04-06 chapter2/chapter2.tex
82 pages

@@
 
-%It includes the rules for which pairs are deemed compatible, the entry and death processes that govern the evolution of the pool, which characteristics are observable to the algorithm at each period, and the reward observed by the agent as sequentially matches and clears pairs.
-More rigorously, at each period an environment is described bu tuple $(G, u)$, where:
-
-% 
-% \begin{itemize}
-%   \item $G = (V, A, X, \pi)$ is a graph with vertex set $V$, adjacency matrix $A$, vertex characteristic matrix $X$, and $\pi: \mathbb{N} \times V \rightarrow \{0, 1\}$ is a function that maps a period $t \in \mathbb{N}$ and a vertex $v \in V$ to whether that vertex was alive on period $t$ (1) or not (0).
-\footnote{We decided to model dynamic graphs in this manner as it is closer to our simulation code.
-Note that all the information about entries and exits can be expressed by the function $\pi$.}
-%   \item $u: 2^V \rightarrow \mathbb{R} \cup \{-\infty\} $ is the reward associated with matching subset of $V$. Infeasible exchanges receive reward $-\infty$.
-% \end{itemize}
-
 \subsection{Environments}
@@
   \item Donors and patients with incomplete tissue-type profile information.
+\end{itemize}
 
@@
 \includegraphics[width=\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/computed_cPRA.pdf}
-\caption{\textbf{Patient and donor cPRA} \\
+\caption[Patient and donor cPRA]{\textbf{Patient and donor cPRA} \\
     Computed from historical data.
@@
 \includegraphics[width=1.4\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/greedy_opt_comparison_mcl2_2.pdf}
-\caption{\textbf{Comparing \textsc{Myopic} and \textsc{OPT}}  \\
+\caption[Comparing \textsc{Myopic} and \textsc{OPT}]{\textbf{Comparing \textsc{Myopic} and \textsc{OPT}}  \\
 Ratio of average per-period matched pairs for different entry and death rates, over 3000 periods (darker hues are better).
@@
 \includegraphics[width=\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/direct_prediction.pdf}.
-\caption{\textbf{Direct prediction method} \\ One period of simulation using direct prediction methods to choose cycles. See Algorithm \ref{alg:direct_prediction} for details.}
+\caption[Direct prediction method]{\textbf{Direct prediction method} \\ One period of simulation using direct prediction methods to choose cycles.
+See Algorithm \ref{alg:direct_prediction} for details.}
 \label{fig:direct_prediction}
@@
   \includegraphics[width=1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/mab.pdf}
-  \caption{\textbf{Multi-armed bandit methods} \\ 
+  \caption[Multi-armed bandit methods (one-period)]{\textbf{Multi-armed bandit methods} \\ 
   One period of simulation using multi-armed bandit methods to choose cycles.}
@@
   \includegraphics[width=1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/mab_step.pdf}
-  \caption{\textbf{Function} \textsc{get\_pseudo\_reward} \\ Multi-armed bandits evaluate if a cycle should be cleared by checking if there is a high chance that the cycle will be used in the future.
+  \caption[Function \textsc{get\_pseudo\_reward}]{\textbf{Function} \textsc{get\_pseudo\_reward} \\ Multi-armed bandits evaluate if a cycle should be cleared by checking if there is a high chance that the cycle will be used in the future.
 Such cycles get a \emph{lower} reward, and are left for later.}
@@
 \includegraphics[width=\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/traditional_ml_mdp_algo_grb}
-\caption{\textbf{Performance of gradient boosting in direct prediction method} \\
+\caption[Performance of gradient boosting in direct prediction method]{\textbf{Performance of gradient boosting in direct prediction method} \\
     Ratio of average matched patients against \textsc{Myopic}.
@@
 \includegraphics[width=\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/traditional_ml_mdp_algo_lr}
-\caption{\textbf{Performance of logistic regression in direct prediction method} \\
+\caption[Performance of logistic regression in direct prediction method]{\textbf{Performance of logistic regression in direct prediction method} \\
     Similar to gradient boosting shown in Figure \ref{fig:traditional_ml_mdp_algo_grb}, logistic regression is only able to perform better than \textsc{Myopic} when the death rate is large.}
@@
 \includegraphics[width=\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/traditional_ml_mdp_algo_rf}
-\caption{\textbf{Performance of random forests in direct prediction method} \\
+\caption[Performance of random forests in direct prediction method]{\textbf{Performance of random forests in direct prediction method} \\
     Random Forests's poor performance is likely due to its low accuracy, as we show on Table \ref{tab:tradional_ml_classifier}.}
@@
 \includegraphics[width=1.1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/mab_OPTN.pdf}
-\caption{\textbf{Performance multi-armed bandit methods across environments} \\
+\caption[Performance of multi-armed bandit across environments]{\textbf{Performance multi-armed bandit methods across environments} \\
 Average percent improvement by the \emph{multi-armed bandit} method over \textsc{Myopic}.
@@
 The idea in those models
+
+[[[Other loss functions]]]
 
@@
   \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter2/tables/traditional_ml_classifier.tex}
-   \caption{\textbf{Performance of statistical methods as classifiers} \\ 
+   \caption[Performance of statistical methods as classifiers]{
+   \textbf{Performance of statistical methods as classifiers} \\ 
    We experiment with three variations on the \emph{direct prediction} method by implementing it using different statistical.
@@
   \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter2/tables/traditional_mdp_grb_none.tex}
-   \caption{\textbf{Gradient Boosting (no graph information)} \\ }
+  \caption[Performance of direct prediction method with gradient boosting (1)]{\textbf{Performance of direct prediction method with gradient boosting} Not using information about graph.}
   \label{tab:grb_none}
@@
   \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter2/tables/traditional_mdp_grb_networkx.tex}
-   \caption{\textbf{Gradient Boosting (augmented with graph information)} \\ }
+  \caption[Performance of direct prediction method with gradient boosting (2)]{\textbf{Performance of direct prediction method with gradient boosting} Using information about graph.}
   \label{tab:grb_networkx}
@@
   \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter2/tables/traditional_mdp_lr_none.tex}
-   \caption{\textbf{Logistic Regression (no graph information)} \\ }
+  \caption[Performance of direct prediction method with logistic regression (1)]{\textbf{Performance of direct prediction method with logistic regression} Not using information about graph.}
   \label{tab:lr_none}
@@
   \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter2/tables/traditional_mdp_lr_networkx.tex}
-   \caption{\textbf{Logistic Regression (augmented with graph information)} \\ }
+  \caption[Performance of direct prediction method with logistic regression (2)]{\textbf{Performance of direct prediction method with logistic regression} Using information about graph.}
   \label{tab:lr_networkx}
@@
   \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter2/tables/traditional_mdp_rf_none.tex}
-   \caption{\textbf{Random Forests (no graph information)} \\ }
+  \caption[Performance of direct prediction method with random forests (1)]{\textbf{Performance of direct prediction method with random forests} Not using information about graph.}
   \label{tab:rf_none}
@@
   \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter2/tables/traditional_mdp_rf_networkx.tex}
-   \caption{\textbf{Random Forests (augmented with graph information)} \\ }
+  \caption[Performance of direct prediction method with random forests (2)]{\textbf{Performance of direct prediction method with random forests} Using information about graph.}
   \label{tab:rf_networkx}
@@
   \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter2/tables/mab_UCB1.tex}
-   \caption{\textbf{Multi-armed bandit algorithm UCB1} \\ 
+   \caption[Performance of multi-armed bandit method using UCB1]{\textbf{Multi-armed bandit algorithm UCB1} \\ 
   \textsc{Environ} is the environment used for simulation.
@@
   \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter2/tables/mab_Thompson.tex}
-  \caption{\textbf{Multi-armed bandit algorithm Thompson Sampling} \\ 
+   \caption[Performance of multi-armed bandit method using Thompson sampling]{\textbf{Multi-armed bandit algorithm Thompson Sampling} \\ 
  \textsc{Environ} is the environment used for simulation.

### 2018-04-09 chapter1/chapter1.tex
89 pages

@@
-\chapter{Heterogenous Production Functions, Panel Data, and Productivity Dispersion}
+\chapter*{Heterogenous Production Functions, Panel Data, and Productivity Dispersion}
+\addcontentsline{toc}{chapter}{Heterogenous Production Functions, Panel Data, and Productivity Dispersion}
 \begin{center}
@@
 % [One or two sentences providing basic introduction to the field, comprehensible to any economist]
-Econometricians often need control for unobserved heterogeneity,
-
-%[Two to three sentences of a more detailed background, comprehensible to econometricians]
-
-%[One sentence clearly stating the general problem being addressed by this particular study]
+Appropriately modelling unobserved heterogeneity is one of the foremost challenges faced by econometricians.
+%[Two to three sentences of a more detailed background, comprehensible to econometricians]  %[One sentence clearly stating the general problem being addressed by this particular study]
 In the context of panel data, it has become standard to controlling for unobserved heterogeneity via a \emph{fixed effects} approach, where we do not impose restrictions the distribution of latent heterogeneity parameters conditional on observable characteristics \citet{wooldridge2010econometric}.
-However, this flexibility often comes at a cost in terms of statistical modeling: the most ubiquitous fixed effects strategy in panel data is via the introduction of a time-invariant random intercept.[\ref{}, \ref{}, \ref{}], but there are many instances of economic problems where this is might not be appropriate (e.g.
-[QUICK EXAMPLE]).
+However, this flexibility often comes at a cost in terms of statistical modeling: the most ubiquitous fixed effects strategy in panel data is via the introduction of a time-invariant random intercept \citep{arellano2003panel}, but there are many instances of economic problems where this is might not be appropriate \citep{browning2007heterogeneity}.
 
@@
 In our work we note, however, that the large observed variation in TFP might be a product of restrictive statistical models: when only the intercept is able to vary, it may end up absorbing most the heterogeneity in the data, leading to inflated estimates of the standard deviation.
-On the other hand, we remark that our specific model and data differ from the ones above so we cannot directly comparison our results, but leave it as a suggestion for future research.
+On the other hand, we remark that our specific model and data differ from the ones above so we cannot directly compare our results, but leave it as a suggestion for future research.
 
@@
 
-\section{Literature review}
+\section{Background and literature review}
 
-In full generality, panel data models can be represented as
+% The class of problems we're interested in
+\paragraph{Panel data and heterogeneity} In full generality, panel data models can be represented as
 \begin{align}
@@
 
-\noindent where $Y_{it}$ is the dependent variable of interest for unit $i$ at period $t$, $H_{it}$ is an individual-specific, possibly infinite-dimensional object often called the \emph{unobserved heterogeneity} component, $X_{it}$ are observable characteristics, and $U_{it}$ is an idiosyncratic error term.
+\noindent where $Y_{it}$ is the dependent variable of interest for unit $i$ at period $t$, $H_{it}$ is an individual-specific, possibly infinite-dimensional object often called the \emph{unobserved heterogeneity} component, $X_{it}$ are observed characteristics, and $U_{it}$ is an idiosyncratic error term.
 For example, $Y_{it}$ might represent an individual's wages over time, $X_{it}$ a vector of characteristics that includes educational attainment, $H_{it}$ a measure of their cognitive or manual ability, and $U_{it}$ measurement error.
-The econometrician is usually interested in some functional of this object, such as the causal average effect of on $Y_{t}$ of a change in $X_{it}$ (known as the \emph{average partial effect} of education on wages).
+The econometrician is usually interested in some functional of this object, such as the average effect of a change in $X_{it}$ on $Y_{t}$ while keeping everything else fixed -- known as the \emph{average partial effect} (APE) \citet{blundell2003endogeneity}.
 
+% What is heterogeneity
 Particular care must be taken with the \emph{unobserved heterogeneity} parameter $H_{it}$.
@@
 The qualifier \emph{unobserved} serves to emphasize that the econometrician does not have access to information about such factors.
-This is problematic because, even when $H_{it}$ itself is not of interest, it may preclude the identification of causal objects To continue our running example: even if we observe that workers with higher educational attainment receive higher wages, we cannot a priori claim that we have found a causal effect of $X_{it}$, as it may be that high-ability ($H_{it}$) workers are able to self-select into higher-paying jobs.
+This is problematic because, even when $H_{it}$ itself is not of interest, it may preclude the identification of causal parameters.
+To continue our running example: even if we observe that workers with higher educational attainment receive higher wages, we cannot a priori claim that we have found a causal effect of $X_{it}$, as it may be that high-ability ($H_{it}$) workers are able to self-select into higher-paying jobs.
 
-One of the foremost advantages of dealing with panel data is that, under specific assumptions, we are able to control for the unobserved heterogeneity component.
-However, without further assumptions, however, nothing can be said.
-Therefore, econometricians have posited a number of restrictions on the model above, each producing different identification strategies.
-
-Econometric applications (such as \cite{???}, \cite{???} often to cite a few) assume that $m$ is a linear function.
-\begin{align}
-  Y_{it} = A_{it} + X_{it}^{T}B_{it} + U_{it}
-\end{align}
-
-\noindent where $H_{it} = [A_{it}, B_{it}]$, a vector of \emph{random coefficients}. Furthermore, they will often impose additionally that 
+% Panel data models help with heterogeneity
+One of the foremost advantages of dealing with panel data is that, under certain conditions, the econometrician may be able to leverage the existence of multiple observations per unit of analysis to control for this unobserved heterogeneity.
+This is possible, for example, if the researcher additionally imposes that $m$ is linear, and that all the unobserved heterogeneity can be explained by an individual but time-invariant shift, as in:
 \begin{align}
@@
 
-and that unobserved heterogeneity appears only as an intercept.
+\noindent where all the heterogeneity $H_{it}$ is entirely subsumed in the random scalar intercept coefficient $A_{i}$.
+Such linear panel data models are the econometric ``workhorse in empirical studies" \citet{wooldridge2005fixed}.
 
-% One of the main objectives is to model heterogeneity
+% Fixed effects vs random effects
+The average partial effect in this case is represented by components of the shared coefficient $\beta$, which can be identified under additional restrictions on the stochastic process governing $U_{it}$.
+In common econometric parlance, identification strategies that allow for $Cov(X_{it}, A_{it}) \neq 0$ are named \emph{fixed effects} models, while those that additionally impose that this covariance must be zero are called \emph{random effect} models.\citet{wooldridge2010econometric, arellano2003panel}.
 
-% Importance of controlling for heterogeneity
+% Intercept not enough
+More generally, \cite{graham2012identification} define the class of \emph{fixed effects} methods as one that implies restrictions on the functional form of $m$ and on the conditional distribution of the idiosyncratic shocks $F(U_{it}|X_{\cdot}, A_{\cdot})$, but leaves the conditional distribution of unobserved heterogeneity $F(A_{it}|X_{it})$ unrestricted.
+A stronger class of \emph{random effects} methods additionally imposes an independence assumption between $A_{\cdot}$ and $X_{\cdot}$.
+While facilitating identification and estimation, the latter is harder to motivate economically -- it would mean, in our example, that workers' wages were unrelated to their unobserved ability (see \cite{griliches1977estimating} for a discussion).
 
-Imposing different restrictions will give rise to different identification strategies.
-As recounted by \cite{graham2012identification}, the \emph{fixed effects} method implies restrictions on the functional form $m$ and on the conditional distribution of the idiosyncratic shocks $F(U_{it}|X_{\cdot}, A_{\cdot})$, but leaves the conditional distribution of unobserved heterogeneity $F(A|X)$ unrestricted.
-A related method, named \emph{random effects}, entails imposing an independence assumption between $A_{\cdot}$ and $X_{\cdot}$, but that usually leads to models that are harder to motivate economically.
+\paragraph{Random coefficients} In the model above, all the heterogeneity is subsumed into one random scalar that entering additively.
+However, the ubiquitousness of such models may be explained more for their convenience than for their economic realism.
+As put by \citet[~p.11]{browning2007heterogeneity},
 
-Further restrictions can be imposed for simplicity.
+\begin{quote}
+\singlespacing
+Almost always decisions on how to include allowance for heterogeneity are made using conventional schemes that have been designed by statisticians to put in the heterogeneity in such a way that we can immediately take it out again.
+\end{quote}
 
-Figure \ref{fig:panel_data_models} illustrates the difference between them.
+% Therefore random coeffs
+Therefore, in this paper we focus on partially rolling back some of the assumptions above and allow for a more general case with more than one correlated random coefficient.
+We will be working with a specific version of this general \emph{random coefficients} model
 
-[[[These APEs are essentially derivatives of the average structural function of Blundell and Powell (2004).]]]
+\begin{align}
+  Y_{it} = A_{it} + X_{it}^T B_{it} + Z_{it}^T \beta
+\end{align}
 
-In this paper we focus on the more general case where we allow for more than one random coefficient, and also allow them to be correlated with the regressor.
-The literature dealing with this more general case is substantially smaller.
-\citet{hoderlein2010analyzing} analyzed a for any number of random coefficients, but required the stronger assumption that the coefficients are independent from the regressors.
+\noindent where now  $H_{it} = [A_{it}, B_{it}^T]^T$, so that now we have \emph{vector-valued heterogeneity}.
+Moreover, we will be principally interest with the case when $A_{it}, B_{it}$ are \emph{endogenous}, that is, not independent of $X_{it}$ (in the opposite case we say that the random coefficients are \emph{exogenous}).
 
-% Closest works are Chamberlain (1982), Arellano and Bonhomme (2011), Graham nd Powell (2012), Evdokimov (2011)]
+The econometric literature dealing with this more general case is substantially smaller.
+Early work by \cite{swamy1970efficient} studied a panel data model with exogenous, stationary random coefficients and produced a feasible GLS estimator for random coefficient averages.
+\cite{mundlak1978models} and \cite{chamberlain1984panel} were one of the first to study panel data models with correlated random coefficients and derived efficient estimators under parametric assumptions about the distribution of random coefficients and errors.
+\cite{chamberlain1992efficiency} derived an efficient semiparametric estimator of the first moment of random coefficient under regularity assumptions.
+Later, \cite{arellano2011identifying} produced identification results for higher moments of the distribution of random coefficients under similar conditions, while \cite{graham2012identification} extended his approach to obtain the first moment in the \emph{irregular} case when such assumptions are not satisfied.
+As we will see, the main feature of irregular identification is convergence rates smaller than root-n \citep{lewbel2016identification}, which in turn happens due to the dependence on aspects of ``small" subpopulations defined on measure-zero sets.
+Our paper can be seen as a further development of these last two papers, as we identify all moments of the distribution of random coefficients under irregularity.
 
-% Production function estimation
-> Recent papers on heterogeneous production functions Kasahara, Shrimpf and Suzuki (2016), Balat, Brambilla and Sazaki (2016)
-  Extend proxy variable approaches but require variables to be chosen in period t − 1 with only static considerations, as in FOCs. Ackerberg and Hahn (2016)
-  Scalar unobservable enters into nonparametric production function
+It should also be noted that even without panel data, there may be instances where aspects of random coefficients may be identified by if the variation in $X_{it}$ can be explained by variation in a set exogenous excluded variables.
+Examples of this \emph{triangular} model include \cite{heckman1998instrumental}, \cite{hoderlein2010analyzing}, and \cite{hoderlein2017triangular}.
 
@@
 
+\paragraph{Production function estimation} Our illustrative empirical application will involve the identification and estimation of a production function, so that $Y_{it}$ will represent output and $X_{it}$ inputs to production.
+This is a vast and comprehensive field on its own, so here we will simply note that two issues are pervasive throughout the literature and relevant to our paper.
+
+First, there is a need to take heterogeneity into account when estimating production functions.
+The recognition of this fact goes at least as far back as \cite{marschak1944random} who note that [p.
+145]
+\begin{quote} 
+\singlespacing
+[T]he production function will change, even within the same industry, from firm to firm and from year to year, depending on the technical knowledge, the will, effort, and luck of a given entrepreneur: these factors (...) may be represented by one or more random parameters.
+\end{quote}
+
+Second, that this heterogeneity is plausibly endogenous, since a firm's choice of inputs is likely correlated with their own (unobserved) productivity \citep{gandhi2011identification}.
+Empirical applications usually address this issue via a dynamic panel approach (as in \cite{arellano1991some}) or via proxy variable approaches (as in \cite{levinsohn2003estimating}).
+
+In allowing for correlated random coefficients, our paper will incorporate a flexible form of vector-valued heterogeneity that takes the endogeneity problem into account.
+
 \section{Model and assumptions}
 
-In order to reduce clutter we will drop the $i$ subscript. For $1 \leq t \leq T$, define
+From now on, in order to reduce clutter we will drop the $i$ subscript. For $1 \leq t \leq T$, define
 \begin{align*}
@@
     \quad     
-    E[W_tW_t^{T}] \text{is nonsingular}
+    E[W_tW_t^{T} | X_1 = X_2] \qquad \text{is nonsingular}
   \end{align}
@@
 In the context of firm profit maximization, this means that a firm is allowed to choose future inputs based on their previous productivity.
-In a different context, they would also allow for a dynamic model with $X_t = Y_{t-1}$.
-Crucially, we have not made any assumptions about contemporaneous independence between random coefficients and regressors, thus allowing for a unobserved heterogeneity of a general type.
+In a different context, this model would also allow for a dynamic model with $X_t = Y_{t-1}$.
+Crucially, we do not make any assumptions of contemporaneous independence between random coefficients and regressors.
 
@@
 
-In this section, we show how to identify the $r^{th}$ moments of the random coefficients $(A_{t}, B_{t})$. Broadly, our identification strategy will follow the next steps.
+In this section, we show how to identify the $r^{th}$ moments of the random coefficients $(A_{t}, B_{t})$. Broadly, our identification strategy will be constructive, and follow these four steps.
 
@@
 \paragraph{Assumption 3.1'} \label{a:sequentialindependence}
-  \textot{Shocks are r\textsuperscript{th}-moment independent from all covariates}
+  \textit{Shocks are r\textsuperscript{th}-moment independent from all covariates}
   \begin{align}
@@
 
-In this section we will prove that we can identify the marginal density of random coefficients $f_{A_tB_t}$. We begin by strengthening the
+In this section we will prove that we can identify the marginal density of random coefficients $f_{A_tB_t}$. We begin by strengthening some of the assumption above.
+
+\paragraph{Assumption 1.1'} \label{a:sequentialindependence}
+\textit{Shocks are jointly independent of past shocks and current and past regressors}
+\begin{align}
+  \text{For }1 \leq j \leq r,
+  \quad     
+  (U_{t}, V_{t}) \perp (U_{1}, V_{1}, \cdots, U_{t-1}, V_{t-1}, X_{1}, \cdots, X_{t-1})
+\end{align}
+
+\paragraph{Assumption 1.5} \label{a:}
+\textit{The joint distribution of shocks is positive everywhere}
+\begin{align}
+  \text{For }1 \leq j \leq r,
+  \quad     
+  \phi_{U_t, V_t}(s_1, s_2) := E[\exp(i[s_1 U_{t_1} + s_{2} V_{t_2}])] > 0 \qquad \forall t \in \mathbb{R}^2
+\end{align}
+
+\begin{theorem}[Nonparametric identification] If Assumptions 1.1', 1.3, 1.4 and 1.5 are satisfied, then the joint distribution of random coefficients $f_{A_t, B_t}$ is identified for all $t$.
+\end{theorem}
+
+\paragraph{Proof sketch} Consider two adjacent periods, here denoted by $t = 1,2$.
+Let $\phi_{Y_{1}, Y_{2}}(s_1, s_2| x_{1},x_{2}) := E[\exp(i[s_1 Y_{1} + s_{2} Y_{2}])| X_{1} = x_{1}, X_{2} = x_{2}]$ be the joint characteristic function of the dependent variable conditional on observable covariates.
+We can identify the characteristic function of shocks by evaluating this characteristic function at the ``diagonal" $X_1 = X_2 = x$ with $s_{2} = -s_{1} = s$.
+\begin{align}
+\phi_{Y_{1}, Y_{2}}(-s, s| x, x) &= E[\exp(i[sY_{2} - sY_{1}])| X_{1} = X_{2} = x] \\
+                          &= E[\exp(i[sU_{2} - sxV_{2}])| X_{1} = X_{2} = x] \\ 
+                          &= E[\exp(i[sU_{2} - sxV_{2}])] \qquad \because \text{Assumption 1.1'}    \\
+                          &=: \phi_{U_{2},V_{2}}(s, s') \qquad s' := sx          
+\end{align}
+
+Applying the inverse Fourier transform to the joint characteristic function of the shocks $\phi_{U_{2},V_{2}}$, we retrieve their distribution $f_{U_{2}, V_{2}}$.
+Note that the full support Assumption 1.3 was used in the inversion step.
+
+Next, as in the moment identification discussion, write
+\begin{align}
+  \begin{bmatrix} Y_{1} \\ Y_{2} \end{bmatrix}
+  =
+  \Gamma_1(x_1, x_2) \begin{bmatrix} A_{1} \\ B_{1} \end{bmatrix} +
+  \Gamma_2(x_1, x_2) \begin{bmatrix} U_{2} \\ V_{2} \end{bmatrix}  
+\end{align}
+\noindent for appropriate matrices $\Gamma_1$ and $\Gamma_2$.
+Now choose $t_1 = s^{T}\Gamma_{1}$ and $t_2 = t\Gamma_{1}^{-1}(x_1, x_2)\Gamma_2(x_1, x_2)$.
+Note that this exists whenever $x_1 \neq x_2$.
+The conditional characteristic function of $Y_{1}, Y_{2}$ then decomposes multiplicatively:
+\begin{align}
+\phi_{Y_{1}, Y_{2}}(t_1, t_2| x_1, x_2) &= E[\exp(it_1^{T}Y_{1} + it_2^{T}Y_{2}])  | X_{1} = x_1, X_{2} = x_2] \\
+          &= E[\exp(is\cdot [A_1, B_1] - it_2 \cdot [U_2, V_2] ) | X_{1} = x_1, X_{2} = x_2] \\    
+          &= E[\exp(is\cdot [A_1, B_1] | X_{1} = x_1, X_{2} = x_2]\cdot E[\exp(it_2 \cdot [U_2, V_2] )] \\
+          &=: \phi_{A_1, B_1}(s_1, s_2 |x_1, x_2)\cdot  \phi_{U_2, V_2}(t_2) \label{eq:decomposition}
+\end{align}
+
+Since we have already identified $\phi_{U_2, V_2}(s)$ for all $s \in \mathbb{R}^2$, we can retrieve $\phi_{Y_1, Y_2}(t_2|x_1, x_2)$ by dividing both sides of Equation \ref{eq:decomposition} by $\phi_{U_2, V_2}(t_2)$ whenever $t_2$ exists.
+This implies that we can retrieve the conditional density function $f{A_1, B_1}(a_1, b_1 | x_1, x_2)$ by applying the inverse Fourier transform to the ratio $\frac{\phi_{Y_1, Y_2}(t_2|x_1, x_2)}{\phi_{U_2, V_2}(t_2)}$.
+Finally, averaging over $X_1, X_2$, we can retrieve the marginal density $f{A_1, B_1}$.\footnote{We are glossing over the fact that we have not identified the characteristic function over the diagonal $x_1 = x_2$.
+For the complete argument please see a supplementary note.}
 
@@
 In this final step, we must exclude observations near the diagonal $X_1 ≈ X_2$, and then average the conditional moments associated with the remaining observations.
-According to our asymptotic results, included observations must satisfy the following conditions:
+For the first moment of the intercept, we compute
+
 \begin{align}
-  |\Delta X_{2i}| &> c_{cens1} \hat{\sigma}_{X_2} n^{-\alpha_{cens1}} \quad \text{ for first moments}\\
-  |\Delta X_{2i}| &> c_{cens1} \hat{\sigma}_{X_2} n^{-\alpha_{cens2}} \quad \text{ for second moments}
+  \hat{E}[A_1] = \frac{\sum_{i=1}^{n} \hat{E}[A_1] | X_{1i}, X_{2i} \cdot 1\{ |X_{2i} - X_{2i}| > t_{1}(n) \} }
+                      {\sum_{i=1}^{n} 1\{ |X_{2i} - X_{2i}| > t_1(n) \} }
 \end{align}
 
-As shown in section \ref{\asymptotic_results}, the optimal thresholds can be shown to be associated with exponents $\alpha_{cens1} = \frac{1}{4}$ and $\alpha_{cens2} = \frac{1}{8}$.
-However, theory gives us no guidance regarding the choice of constants $c_{cens1}$ and $c_{cens2}$, so we resort to experimentation in the next section.
+\noindent and analogously for the slope and for higher moments.
+Here, $t_{1}(n)$ and $t_2(n)$ (for second moments) are \emph{censoring threshold} functions that vanish asymptotically.
+Consistency and asymptotic normality of the resulting estimates are ensured provided that
+\begin{align}
+ t_1(n) \equiv c_{cens1} \hat{\sigma}_{X_2} n^{-\frac{1}{4}} \quad \text{ for first moments}\\
+ t_2(n) \equiv c_{cens1} \hat{\sigma}_{X_2} n^{-\frac{1}{4}} \quad \text{ for second moments}
+\end{align}
+
+\noindent where $c_{cens1}$, $c_{cens2}$ are constants that are not pinned down by theory. In order to choose these constants, we resort to experimentation as shown in section \ref{sec:simulation}.
+
+%-------------------
+% Asymptotic results
+%--------------------
+
+\section{Asymptotic results}
+
+In this section we present an abridged version of our asymptotic results without proof.\footnote{For an elaborated discussion, please see supplementary material available upon request.)}
+
+\begin{definition} Define the following objective functions
+  \begin{align}
+    \hat{G}_{n}(\alpha) &= \frac{1}{n} \sum_{i=1}^{n} \left [\left( \hat{E}(A_1|X_1, X_2) - \alpha \right)^2 1\{ | X_2 - X1 | > h_n \} \right] \label{eq:hat_g} \\ 
+    G_{n}(\alpha) &= E\left [\left( \hat{E}(A_1|X_1, X_2) - \alpha \right)^2 1\{ | X_2 - X1 | > h_n \} \right]  \\ 
+    \Gamma(\alpha) &= E\left [\left( \hat{E}(A_1|X_1, X_2) - \alpha \right)^2 \right]  \label{eq:gamma}
+  \end{align}
+  and let $\hat{\alpha}$ and $\alpha_{0}$ be the minimizers of \ref{eq:hat_g} and \ref{eq:gamma}. Explicitly:
+  \begin{align}
+    \hat{\alpha} &= \frac{ \sum_{i=1}^{n} \hat{E}[A_{1} | X_{i1}, X_{2i}] 1\{| X_{2i} - X_{1i}| > h_{n}\}  } { \sum_{i=1}^{n} 1\{ |X_{2i} - X_{1i}| > h_{n} \} } \\
+    \alpha_{0} &= E[E[(A_1 | X_{1}, X_{2})]] = E[A_{1}]
+  \end{align}
+\end{definition}
+
+\begin{theorem}[Consistency] If the following two conditions hold 
+  \begin{enumerate}
+    \item $\sup_{\alpha} \left| \hat{G}_{n}(\alpha) - \Gamma(\alpha) \right| \in o_{p}(1)$
+      \item For each $\delta > 0$, $\inf_{\alpha: |\alpha - \alpha_{0}| \Gamma(\alpha) > \Gamma(\alpha_0)}$ 
+  \end{enumerate}
+  Then $|\hat{\alpha} - \alpha_{0}| \in o_{p}(1)$
+\end{theorem}
+
+\begin{theorem}[Asymptotic Normality for conditional moments] Suppose that $x_1 \neq x_2$ and $f(x_1, x_2) > 0$. If the following conditions hold
+  \begin{enumerate}
+    \item For $k \in \{0, 1\}$, the function $r_{k}(x_1, x_2, \theta)$ defined below has continuous second partial derivative with respect to $x_1$.
+    \begin{align}
+      r_{k}(x_1, x_2, \theta) = E[(Y_2  - Y_1 - EU_2 - EV_2X_2)X_2^{k} | X_1 = x_1, X_2 = x_2]f(X_1 = x=1, X_2 = x_2)
+    \end{align}
+    \item The function $s_t(x_1, x_2) = E[Y_1 | X_1 = x_1, X_2 = x_2]f(X_2 = x_2 | X_1 = x_1)$ is twice continuously differentiable
+    \item The marginal density of the first regressor $f(x_1)$ is twice continuously differentiable
+    \item The conditional means of $Y_{1}$ and $Y_{2}$ are estimated with uniform kernels endowed with bandwidth $h_{nw}(n) \propto n^{-\alpha_{nw}}$, where $\frac{1}{5} < \alpha_{nw} < \frac{1}{2}$
+    \item The censoring threshold functions decay as the following rates $t_1(n) \propto n^{-\frac{1}{4}}$ and $t_2 \propto n^{-\frac{1}{8}}$
+  \end{enumerate}
+\noindent 
+Then the estimates of conditional first and second moments are asymptotically normal.
+% \begin{align}
+%   \sqrt{nh_n^2}
+%   \begin{bmatrix}
+%     \widehat{E}[A_{1} | X_1 = x_1, X_2 = x_2] - E[A_1 | X_1 = x_1, X_2 = x_2] \\
+%     \widehat{E}[A_{1} | X_1 = x_1, X_2 = x_2] - E[B_1 | X_1 = x_1, X_2 = x_2] 
+%   \end{bmatrix}
+%   \rightarrow^{D}
+%   \mathcal{N}
+%     \left(
+%     \begin{bmatrix} 0 \\ \end{bmatrix},
+%     V(x_1, x_2)
+%     \right)
+% \end{align}
+% \noindent for a given matrix $V(x_1, x_2)$. A similar result holds for 
+\end{theorem}
 
@@
 The performance of our estimators is illustrated on Figures \ref{fig:simulation_shocks} and \ref{fig:simulation_random_coefficients} and on the accompanying tables on the Appendix.
-We observe that indeed the distribution tends to collapse to its true value as the number of observations increases.
+We observe that indeed the distribution tends to concentrate around its true value as the number of observations increases.
 
@@
 
-Now we provide an illustration of our methods by estimating a Cobb-Douglas production function with random elasticity for a large number of Indian plants.
+Now we provide an illustration of our methods by estimating a Cobb-Douglas production function with random coefficients for a large number of Indian plants.
 We will be using the Annual Survey of Industries (ASI) dataset, released by the Central Statistical Organization of India for the years 2008 and 2009.
 This yearly survey collects data from sampled Indian economic units of production (individual factories, workshops, and establishments, hereafter called ``plants") that employ more than 10 regular workers.
-From this data we extract four variables: \emph{gross sale value} (value of the products sold by the plant, as purchased by their clients) $S_{t}$; \emph{capital} (fixed assets with a productive life of more than one year) $K_{t}$; \emph{wages} $W_{t}$; and production \emph{materials} $M_{t}$.
+
+From the ASI dataset we extract four variables: \emph{gross sale value} (value of the products sold by the plant, as purchased by their clients) $S_{t}$; \emph{capital} (fixed assets with a productive life of more than one year) $K_{t}$; \emph{wages} $W_{t}$; and production \emph{materials} $M_{t}$.
 All variables are in 2005 rupees.
-We keep only observations that we present in both years of analysis, and we are left with 13298 observations.
-Summary statistics are in subsection \ref{sec:summary_statistics} in the Appendix.
+We keep only observations that we present in both years of analysis, and we are left with 13298 observations.\footnote{Additional information about data cleaning and manipulation is available as Jupyter Notebook at \url{https://github.com/halflearned/FHHPS/blob/master/empirical/FHHPS_data_cleaning.html}}
 
@@
 \noindent where $Y_{t}$ represents production value-added after sales normalized by wages, and $X_t$ represents normalized capital.
+Summary statistics for all variance can be found in subsection \ref{sec:summary_statistics} in the Appendix.
 
-% Interpretation
+\paragraph{Interpretation} Each plant's output is modeled as a Hicks-neutral Cobb-Douglas function.
+Therefore, the random variable $A_{t}$ represents \emph{total factor productivity}, while $B_{t}$ represents the expenditure share on capital inputs.å
 
-;;Comparison to static FOCS:
-;; Estimate BK as expenditure share on capital input i,t
-;; Based on static profit maximizing FOC for Cobb-Douglas
-;;Treats input as having no adjustment costs, opposite of our
-approach
-;;; Based on strong conduct assumption: static profit maximization
+Included in the modeling decision above there are several assumptions.
+First, timing: we are assuming that each firm observes the entire history of past inputs and outputs $(Y_{1}, X_{1}, \cdots, Y_{t-1}, X_{t-1})$, as well as the values of their own TFP and capital share $(A_{1}, B_{1}, \cdots, A_{t-1}, B_{t-1})$.
+Based on this information, they choose this period's inputs $X_{t}$.
+Once the decision is taken, this period's shocks $U_{t}, V_{t}$ are revealed, and production $Y_{t}$ is realized.
 
-We do not impose profit maximization, important for studying unproductive firms
-We allow unobservables outside production function to affect input choice
-Firm-specific input prices
-Adjustment costs
-Product demand
+Second, we have left the stochastic process governing the evolution of $X_{t}$ unrestricted.
+In particular, we do not impose the assumption that firms are profit-maximizers -- although they may be.
+This is important for studying unproductive firms as well.
 

### 2018-04-09 chapter2/chapter2.tex
89 pages

@@
-\chapter{Two novel algorithms for dynamic kidney exchange}
+\chapter*{Two novel algorithms for dynamic kidney exchange}
+\addcontentsline{toc}{chapter}{Two novel algorithms for dynamic kidney exchange}
+\begin{center}
+\large
+\textsc{Vitor Hadad}
+\end{center}
 
@@
 \begin{landscape}
+\vspace*{2cm}
 \begin{figure}[h]
@@
 \caption[Intuition for pseudo-rewards]{\textbf{Intuition for pseudo-rewards} \\
-  In the example above, at time $t$ we have three available cycles to choose from: $(1,2)$, $(1,3)$ and $(3,4)$, and we are contemplating choosing cycle $(1,2)$ at $t$.
-If we  constrain ourselves to do so, several other future cycles become unavailable (in red).
-Nevertheless, in three out of our four simulations we were able to clear the same number of cycles in the constrained scenario, so the running pseudo-reward average is $\frac{3}{4}$.
+  Each row in the example above, is supposed to represent one simulation.
+The three tuples on the left, $(1,2)$, $(1,3)$ and $(3,4)$, represent the three cycles that are available to be matched today.
+The tuples to the right indicate cycles that appear once we simulate $h$ periods into the future.
+In this example, we are contemplating selecting and clearing cycle $(1,2)$ at $t$.
+\\
+  As explained in Figure \ref{fig:get_pseudo_reward}, at each simulation we solve the problem twice: once where all cycles between $t$ and $t+h$ are available, and a second time where we are constrained to match cycle $(1,2)$ at $t$.
+In the constrained case, all other cycles that involve pair $1$ or pair $2$ become unavailable (in red).
+\\
+  In three out of our four simulations we were able to clear the same number of cycles in the constrained scenario, so the running pseudo-reward average is $\frac{3}{4}$.
+This indicates that, more often than not, removing the cycle $(1,2)$ today will not decrease the overall long-run payoff, so matching it today is desirable.
 }
@@
 \end{figure}
+\vspace*{\fill}
 \end{landscape}
@@
 \end{figure}
-bb
+
 \begin{figure}
@@
 In particular, recent years have seen the emergence of interest in generalizing neural networks to model with structured datasets such as graphs.
-A promising approach is the one taken by \cite{kipf2016semi}, who build on earlier work by \cite{defferrard2016convolutional} to create a very simple recurrence relation for graph embedding:
+A promising approach is the one taken by \cite{kipf2016semi}, who build on earlier work by \cite{defferrard2016convolutional} to derive a very simple recurrence relation for graph embedding:
 \begin{align}
-  H^{\ell + 1} = \sigma((A_{t} + I) H^{\ell}W^{\ell}) \qquad \text{with} \quad H^{\ell} = X_{t}
+  H^{\ell + 1} = \sigma((A_{t} + I) H^{\ell}W^{\ell}) \qquad \text{with} \quad H^{0} = X_{t}
 \end{align}
@@
 The authors prove that this can be interpreted as a differentiable version of the Weisfeiler-Lehman algorithm for graph isomorphism tests.
-For us, it is an especially convenient method because the input graphs need not be the same size throughout.
+For us, it is an especially convenient method because the input graphs need not be the same size throughout the analysis.
 
-Second, the \emph{direct prediction} method produces a probability for every pair, but it does not consider correlations between multiple pairs in the same kidney exchange pool.
-This is unrealistic, as it should be often the case that pairs in the same exchange should have roughly equally high or low probability of being matched.
-In order to rectify this shortcoming of our approach, we would like to train our statistical estimator using the sequence-to-sequence models of recurrent neural networks studied by \cite{sutskever2014sequence} (see also  \citep{lecun2015deep} for an overview).
-The idea in those models
+Second, when the \emph{direct prediction} method computes a selection probability for each pair, it does not consider correlations between multiple pairs in the same kidney exchange pool.
+However, intuition suggests that it should be often the case that pairs in the same exchange have similar probabilities of being matched.
+In order to rectify this shortcoming of our approach, we would like to train our statistical estimator using the sequence-to-sequence (\emph{seq2seq}) models of recurrent neural networks studied by \cite{sutskever2014sequence}, \cite{kalchbrenner2013recurrent} and \cite{cho2014learning}.
+Sequence-to-sequence models are a relatively new paradigm that allows for variable length inputs (needed here because different pools have different sizes), and correlated outcomes of variable length.
+(See also  \citep{lecun2015deep} for an overview)
 
-[[[Other loss functions]]]
+\paragraph{Improving the MAB algorithm}
 
-. We will be turning our attention to these methods next.
+Our \emph{multi-armed bandit} method can be quite computationally intensive, requiring multiple simulations per exchange.
+When the exchange pool or the set of available actions is large, it may be difficult to scale.
+It is important to consider extensions to deal with this issue.
 
-Second, while we see that the performance of our \emph{multi-armed bandit} methods was already satisfactory, we note that it uses only simulations and no data to decide which exchange to clear at each period.
-In that sense, it worked in a diametrically opposite way to the \emph{direct prediction} method that used only data and no simulations.
-In fact, an intermediate approach, using \emph{contextual bandits} \citep{lattimore2018bandits} could be used, blending information about data and simulations.
+First, note how our MAB algorithm only makes use of simulations to select exchange that will be clear, and it makes no use explicit use of information about pairs characteristics, or about the structure of the graph.
+However, it is intuitive that pairs endowed with similar characteristics should have similar probabilities of being matched.
+In terms of the model discussed in this paper, this would translate into correlated pseudo-rewards among pairs, where the strength of the correlation should be proportional to how similar two pairs are deemed to be.
+The generalized versions of the Thompson Sampling algorithm (See \citet{russo2017tutorial}) are able to deal with this, and could be employed.
 
-Thirdly, we would like to experiment with different objective functions, since, in reality, different exchanges may have different levels of desirability or priority.
+% A principled way of going about this extension would be to 
+% 
+% 
+% 
+% \begin{algorithm}[htbp]
+%   \SetAlgoLined
+%   \DontPrintSemicolon
+%   \KwData{Horizon $h$; Number of iterations $L$; Threshold $thres$; \\
+%     Environment simulator object \textsc{Env};
+%     Multi-armed bandit algorithm object \textsc{MAB};
+%   }
+% 
+%   \SetKwFunction{FChoose}{\texttt{extended\_mab}}
+%   \SetKwProg{Fn}{Function}{:}{}
+% 
+%   \Fn{\FChoose{\textsc{Env},\ \texttt{h}, \ $thres$ }}{
+%     \tcp{Begin iterations} \\
+%     \For{$i\gets0$ \KwTo $L$}{
+%       \tcp{} \\
+%       $S_{t} = \arg\max_{C \subset \mathcal{C}} \sum_{c} {w_c}$
+%       \tcp{a} \\
+%       $b$
+%     }
+%   }
+%  \caption{Function \texttt{extended\_mab}}
+%  \label{alg:extended_mab}
+% \end{algorithm}
+
+\paragraph{Other extensions} We would like to experiment with different objective functions, since, in reality, different exchanges may have different levels of desirability or priority.
 For example, it is common for pediatric patients and previous organ donors receive higher priority, as do \emph{ABDR0} exchanges involving ``perfectly matched" patients and donors.
-
-Lastly, we remark that there still remains a moderately large gap between what could be optimally achieved if we could perfectly predict the evolution of the pool.
-It is conceivable that as better predictive methods arise, this gap will become smaller.
 
@@
   \singlespacing
-  \caption{Blood type probabilities in the ABO environment}
+  \caption{Blood type probabilities in the ABO and RSU environments}
   \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter2/tables/abo_probability.tex}

### 2018-04-11 abstract.tex
Almost done: 97 pages

\chapter{Abstract}

\doublespacing
This dissertation is divided into two parts.

\paragraph{Part I - Dynamic Kidney Exchange} In recent years, kidney paired donation (KPD) has an emerged as an attractive alternative for end-stage renal disease patients with incompatible living donors. However, we argue that the matching algorithm currently used by organ clearinghouses is inefficient, in the sense that a larger number of patients may be reached if kidney transplant centers take into consideration how their pool of patients and donors will evolve over time. In our work \emph{Two Novel Algorithms for Dynamic Kidney Exchange}, we explore this claim and propose new computational algorithms to increase the cardinality of matchings in a discrete-time dynamic kidney exchange model with Poisson entries and Geometric deaths.

Our algorithms are classified into \emph{direct prediction} methods and \emph{multi-armed} bandit methods. In the \emph{direct prediction method}, we use machine learning estimator to produce a probability that each patient-donor pair should be matched today, as opposed to being left for a future matching. The estimators are trained on offline optimal solutions. In contrast, in \emph{multi-armed bandit} methods, we use simulations to evaluate the desirability of different matchings. Since the amount of different matchings is enormous, multi-armed bandits (MAB) are employed to decrease order to decrease the computational burden.

Our methods are evaluated using simulations in a variety of simulation configurations. We find that the performance of at least one of our methods, based on multi-armed bandit algorithms, is able to uniformly dominate the myopic method that is used by kidney transplants in practice.

We restrict our experiments to pairwise kidney exchange, but the methods described here are easily extensible, computational constraints permitting.

\paragraph{Part II - Econometrics}  In our econometric paper \emph{Heterogenous Production Functions, Panel Data, and Productivity}, we present methods for identification of moments and nonparametric marginal distributions of endogenous random coefficient models in fixed-$T$ linear panel data models.

Our identification strategy is constructive, immediately leading to relatively simple estimators that can be shown to be consistent and asymptotically normal. Because our strategy makes use of special properties of ``small" (measure-zero) subpopulations, our estimators are \emph{irregularly} identified: they can be shown to be consistent and asymptotically Normal, but converge at rates slower than root-n.

We provide an illustration of our methods by estimating first and second moments of random Cobb-Douglas coefficients in production functions, using Indian plant-level microdata.

### 2019-09-13 ec-18/sample-ec-submission.tex

@@
 		Natural Science Foundation of
-		China}{http://dx.doi.org/10.13039/501100001809} under Grant
+		China}{https://doi.org/10.13039/501100001809} under Grant
 	No.:~\grantnum{GS501100001809}{61273304\_a}
